### 2015-03-05 Paper/paper.tex
Initial commit

% File: paper.tex 
%  Cre: 2013-09-12 
%  $Id: paper.tex 4261 2015-01-27 23:49:04Z jhennawi $

\documentclass[iop]{emulateapj}
\bibliographystyle{apj}
\usepackage{amsmath} 
\usepackage{aas_macros,amssymb}
\usepackage{natbib,graphicx,layouts}
\usepackage{color,hyperref}
\usepackage{bm}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\definecolor{notecolor}{rgb}{0.8,0,0}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
  filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing}

\shorttitle{Pressure smoothing in nonlinear IGM}
\shortauthors{Kulkarni et al.}

\begin{document}

\title{Flagging gravitational wave sources from radio jet morphology}
\author{Girish Kulkarni\altaffilmark{1} and
  Abraham Loeb\altaffilmark{2}}
\altaffiltext{1}{Institute of Astronomy, University of Cambridge,
  Madingley Road, Cambridge CB3 0HA, UK;
  \email{kulkarni@ast.cam.ac.uk}}
\altaffiltext{2}{Harvard}

\begin{abstract}
Apart from a stochastic gravitational wave (GW) background, LISA will
also be able to detect GW emission from binary black holes. The number
of such binary sources depends on the major merger rates of gas-rich
galaxies and subsequent evolution of their central black holes.  The
latter of these is uncertain.  Therefore, an alternate method of
flagging GW sources is useful.  Here we show that if one of the binary
BHs is accreting and has a jet, then the dynamical evolution of the
binary is imprinted on the jet morphology.  Using this dependance, it
is possible to flag binary BHs that are entering their GW inspiral
phase in the present epoch.  We further show that this relationship
between the binary dynamics and the jet morphology has two other
applications: (1) at large binary separations the Mpc-scale jet
morphology can probe the acts as a probe of the pc-scale gas supply to
the binary, and (2) at smaller separations the jet morphology acts as
a test of GR.
\end{abstract}

\keywords{cosmology: IGM}

\section{Introduction}
\label{sec:intro}

Consider two black holes with masses $M_1$ and $M_2$ in a circular
orbit of radius $a$ around each other \citep{2010PhRvD..81d7503L}.
Let $M_1=M_2$ (assume equal masses for simplicity).  The orbital speed
of each black hole is given by
\begin{equation}
  v_\mathrm{orb} = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2} = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
  \label{eqn:vorb}
\end{equation}
where $a_{16}\equiv a/10^{16}$ cm and $M_8\equiv M/10^8$
M$_\odot$. Now imagine a jet coming out of one of the BHs at an angle
$\chi$ to the direction of the orbital angular momentum.  Then because
of the orbital motion, the jet will sweep out a cone.  This is shown
in Figure \ref{fig:BH_orbit}.  The half opening angle of the cone is
given by
\begin{equation}
  \sin\psi = \frac{v_\mathrm{orb}\cos\chi}{v_\mathrm{jet}},
\end{equation}
where $v_\mathrm{jet}$ is the ballistic speed of the jet.  Figure
\ref{fig:jet} shows how this looks.

For observable predictions we have to project this in observer plane.
Let the angle between the jet direction and the observer line of sight
be $i$.  Then
\begin{equation}
  \psi^\mathrm{obs}\approx\frac{\psi}{\sin i}.
\end{equation}

\section{Jet structure when binary enters GW-induced inspiral}

Now when the binary is at the separation at which gravitational
wave-induced inspiral begins, it has a velocity given by Equation (5)
of your 2009 paper:
\begin{equation}
  v_\mathrm{orb} = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
  \label{eqn:v}
\end{equation}
where all symbols on the right hand side have the same meaning as in
your paper.  The separation between the two black at this instant is
given by Equation (7) of your paper:
\begin{equation}
  a = 8.3\times 10^{-3}\, \mathrm{pc}\, M_8^{3/4}\dot{\mathcal{M}}^{-1/4}.
\end{equation}
This gives
\begin{equation}
  \psi^\mathrm{obs}=2.42^\circ \left(\frac{v_\mathrm{jet}}{0.95 c}\right)^{-1}\cos\left(\frac{\chi}{30^{\circ}}\right)\left[\sin\left(\frac{i}{15^{\circ}}\right)\right]^{-1},
\end{equation}
This value in itself is probably not very interesting as it depends on
$v_\mathrm{jet}$ and $\chi$, but it gets interesting when we consider
the evolution of $\psi^\mathrm{obs}$ below.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw (0,0) ellipse (3 and 0.5);
      \fill[black] (0,0) circle (1pt);
      %\draw [->,densely dashed] (0,0) -- (0,2.0);
      \fill[black] (2,-0.39) circle (3pt);
      \fill[black] (-2,0.39) circle (3pt);
      \draw [->] (2,-0.39) -- (3.5,-0.1);
      \draw [->,densely dashed] (2,-0.39) -- (2,1.0);
      \draw [->] (2,-0.39) -- (2.5,0.8);
      \draw [->,red] (2,-0.39) -- (3.5,0.8);
      \draw (4.1,-0.1) node {${\mathbf{v}}_\mathrm{orbital}$};
      \draw (2.7,1.0) node {${\mathbf{v}}_\mathrm{jet}$};
      \draw (4.0,1.0) node {$\tilde{\mathbf{v}}_\mathrm{jet}$};
    \end{tikzpicture}
  \end{center}
  \caption{Jet and orbital configuration. The effect of
    $\mathbf{v}_\mathrm{orbital}$ is to precess the jet as the BH
    moves along the orbit. Note that
    $\tilde{\mathbf{v}}_\mathrm{jet}={\mathbf{v}}_\mathrm{orbital}+{\mathbf{v}}_\mathrm{jet}$.}
  \label{fig:BH_orbit}
\end{figure}

\section{Evolution of jet structure}

Binary separation, jet velocity, and the two angles, tell us about the
instantaneous configuration of the jet.  (Recall that we are only
considering an equal mass binary.)  The jet evolution is governed by
the dynamical evolution of the binary separation.  During the
gas-dominated phase, this is given by
\begin{equation}
\frac{a}{\dot a} = \mu/\dot M = 1.1\times 10^7\, \mathrm{yr}\,\dot{\mathcal{M}}^{-1}.   
\end{equation}
This increases by two orders of magnitude during the GW phase to
\begin{equation}
\frac{a}{\dot a} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
\end{equation}
From Equation (\ref{eqn:vorb}), the orbital velocity of each BH evolves as
\begin{equation}
  \frac{\dot v_\mathrm{orb}}{v_\mathrm{orb}} = -\frac{1}{2}\frac{\dot a}{a}.
\end{equation}
which gives the evolution of the opening angle of the cone as 
\begin{equation}
  \dot\psi^\mathrm{obs}=\dot\psi = -\frac{1}{2\cot\psi}\frac{\dot a}{a}.
\end{equation}
This can be solved analytically to get
\begin{equation}
  \sin\psi = \sin\psi_0\left(\frac{a_0}{a}\right),
\end{equation}
where $\psi_0$ and $a_0$ are the initial values. Thus as the binary
continues to shrink the BH orbital velocity grows and the opening
angle of the conical jet increases (cone gets wider and wider).  The
rate at which this happens tells everything about the binary's
dynamical evolution.

Figures \ref{fig:a_gw} show this evolution for the numbers assumed
above.

\section{Use jets to flag GW binaries}

Figure \ref{fig:a_gas} shows the evolution of the binary when it is in
the gas accretion phase.  We have chosen steady gas accretion with
$\dot{\mathcal{M}}=1$ here for simplicity.  Note that the starting
separation here is about 100 pc.

What we find is that in the beginning the binary orbital period is
about $10^7$ yr. So the jet will show no signs of precision at this
point.  However as the binary separation reduces to about 10 pc, the
period reduces to $\sim 100$ yr and the jet will start showing
precision.  At this point the time scale for binary merger is $\sim
10^7$ yr.

In other words \emph{since the gas accretion time scale is of the
  order of the typical jet lifetime} the jet structure will show a
transition from a straight jet to a precessing jet at a separation of
\begin{equation}
  v_\mathrm{apparant}\left(\frac{t_0}{10^7}\right).
\end{equation}
from the binary.  \textbf{[Convert this to arcseconds on the sky.]}
The location at which the jet morphology shows this precision and the
opening angle evolution give a direct constraint on the epoch of
graviational wave emission of the binary.

Thus the new insight here is that the signature of the jet precision
due to GW-inspiral is observationally accessible.

\section{Jet morphology as probe of GR}

Variability of lensed quasars constraints the size of accretion disks
to about $10^{15}$ cm ($\sim 10^{-3}$ pc) \citep{2010ApJ...712.1129M}.
This is smaller than the $a_\mathrm{GW}$ by a factor of less than 10.

This means that the GW-induced precision will be visible in the jet
precision for a very short distance near the source before the
accretion disk is disrupted.

Last plots in this note show some example jets.  In all these plots
$\psi$ is the half-opening angle of the conical jet, and $i$ is the
angle between the jet and the line of sight.

A rough argument about observability: For our example, the period is
0.1 yr at $t = 63967.3816055$ yr.  We can assume that the chirp begins
at $t = 10^4$ yr when period is $\sim 2$ yr.  So the total duration of
the chirp is $5.3\times 10^4$ yr.

A 2 yr period gives $\sim 1$ milli-arcsec wiggles in the jet.  So the
idea is that the wiggles will change from milli-arcsec to micro-arcsec
over $\sim 10^5$ yr.  That is over a length scale of $\sim 30$ kpc.

Now the question is: how big an angle would a 30 kpc jet subtend on
the sky in arcsecs?

Let us assume $z = 0.302$ which is the redshift of 1928+738 jet that
Roos (1993) have studied.  At this redshift 30 kpc will subtend about
6.748195445702 arcsec.

So we are talking about mciro- to milli-arcssec scale wiggles in tens
of arcsec scale jets.  Jets are routinely seen over hundreds of
arcsec.

\section{Conclusion}

From Figures \ref{fig:a_gw} we can make these four points:

\begin{itemize}
\item From the jet structure, we can tell how long until the binary enters GW-induced inspiral.
\item In particular, we can flag binaries that are entering GW-inspiral now.
\item We could locate binaries that are now in GW inspiral phase but
  have helical jet structures that are impossible to resolve.
\item From the above jet structure, we get an idea of the history of gas inflow on the binary.
\end{itemize}

There will also probably be some spectral evolution along the jet.
This spectral evolution will act as a cross check on the evolution
deduced from the jet morphology.  An important advantage of this
technique is that it allows us to locate binaries that may be
impossible to locate even with VLBI.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw [->,densely dashed,thick] (0,0) -- (3.0,0);
      \draw [->,densely dashed,thick] (0,0) -- (0,4.0);
      \draw [->,thick] (0,0) -- (2.0,4.0);
      \draw [->,thick,red] (0,0) -- (5.0,4.0);
      \draw (1.0,2.0) arc (63:90:2.2);
      \draw (2.0,1.6) arc (38.6:63:2.56);
      \draw (0.5,2.5) node {$\chi$};
      \draw (1.90,2.3) node {$\psi$};
      \draw [densely dashed,thick] (0,0) -- (3.0,6.0);
      \draw (0.0,4.5) node {$\mathbf{L}_\mathrm{orbital}$};
      \draw (4.0,0.0) node {$\mathbf{v}_\mathrm{orbital}$};
      \draw (1.0,3.0) node {$\mathbf{v}_\mathrm{jet}$};
    \end{tikzpicture}
  \end{center}
  \caption{The orbital velocity $\mathbf{v}_\mathrm{orbital}$
    introduces the jet precision angle $\psi$.  Without the orbital
    motion, $\psi=0$ and the jet traces a cylindrical surface.  With
    the orbital motion, as the BH moves along its circular orbit, the
    red vector oscillates around the solid black vector and produces a
    conical surface with $\psi$ as its half-opening angle.  Note that
    $v_\mathrm{orbital}$ is exagerated here.  In practice,
    $v_\mathrm{orbital}\ll v_\mathrm{jet}\sim c$.  This figure then
    shows that
    $\sin{\psi}\sim\tan{\psi}=v_\mathrm{orbital}\cos{\chi}/v_\mathrm{jet}$.}
  \label{fig:jet}
\end{figure}

\begin{figure*}
\begin{center}
  \begin{tabular}{ccc}
    \includegraphics*[width=2.2in]{a_gw.pdf} &
    \includegraphics*[width=2.2in]{v_orb.pdf} &
    \includegraphics*[width=2.2in]{psi.pdf}
  \end{tabular}
\end{center}
\caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
\label{fig:a_gw}
\end{figure*}

\begin{figure*}
\begin{center}
  \begin{tabular}{ccc}
    \includegraphics*[width=2.2in]{a_gas.pdf} &
    \includegraphics*[width=2.2in]{v_orb_gas.pdf} &
    \includegraphics*[width=2.2in]{psi_gas.pdf}
  \end{tabular}
\end{center}
\caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
\label{fig:a_gas}
\end{figure*}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics*[width=\columnwidth]{a_gw.pdf}
%%   \end{center}
%%   \caption{Evolution of the binary separation.  Time $t$ is time since
%%     the gravitational wave induced phase began.}
%%   \label{fig:a_gw}
%% \end{figure}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics*[width=\columnwidth]{v_orb.pdf}
%%   \end{center}
%%   \caption{Evolution of the orbital velocity}
%%   \label{fig:v_orb}
%% \end{figure}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics*[width=\columnwidth]{psi.pdf}
%%   \end{center}
%%   \caption{Evolution of the intrinsic half-opening angle of the
%%     conical jet.  it is important to note that the jet lifetime is
%%     about $10^7$ yr.}
%%   \label{fig:psi}
%% \end{figure}

\pagebreak
\pagebreak

\clearpage

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{jet_i10_psi20.pdf}
  \end{center}
  \caption{$i=10^\circ$, $\psi=20^\circ$, $\beta=0.90$}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{jet_i40_psi20.pdf}
  \end{center}
  \caption{$i=40^\circ$, $\psi=20^\circ$, $\beta=0.90$}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{jet_i80_psi20.pdf}
  \end{center}
  \caption{$i=80^\circ$, $\psi=20^\circ$, $\beta=0.90$}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{jet_i40_psi40.pdf}
  \end{center}
  \caption{$i=40^\circ$, $\psi=40^\circ$, $\beta=0.90$}
\end{figure}

\bibliography{refs}
\end{document}

### 2015-03-05 Paper/scratch.tex
Initial commit

    \draw [->,decorate,decoration={snake, post length=1mm}] (-3.0,0.0)
    -- (-0.1,0.0) node[midway,above] {$\epsilon=h\nu$} ; 
    \draw [->,decorate,decoration={snake, post length=1mm}] (0.1,0.1)
    -- (2.5,1.7) node[midway,sloped,above] {$\epsilon_1=h\nu_1$};
    \draw (1.0,0.0) arc (0:37:1cm);
    \draw (1.2,0.35) node {$\theta$};
    \draw (1.3,-0.7) node {$\mathbf{p},E$};
    \fill (0,0) circle (2pt);
    \draw [->] (0,0) -- (1.5,-1.5);
    \draw (1.75,-1.75) node {$\vec{P}_{ef}$};
    \draw (-3.35,0.0) node {$\vec{P}_{\gamma i}$};
    \draw (2.8,1.85) node {$\vec{P}_{\gamma f}$};

, and imgine further than
the angle between the jet direction and the observer line of sight is
$i$.

What are the advantages of this method of flagging LISA sources?  Can
you go down to binaries that cannot be resolved with VLBI? Can you go
down in mass or mass ratio or something?

  = 2.42\, \mathrm{degrees}.

### 2015-03-05 gas.py

@@ -1,0 +1,99 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+
+M = 1.0 # 1.0e8 Msun
+coeff = -2.56e5 
+
+t0 = 1.0e-3 # yr
+a0 = 1.0e5 # 1.0e16 cm
+
+t = np.logspace(-3.0,8.0,num=100)
+
+def binary_separation_gas(t):
+    k = -1.0e7 # yr 
+    return a0*np.exp((t-t0)/k)
+
+a_gas = binary_separation_gas(t)
+
+def orbital_speed(a_16):
+    return 5.8e3 / np.sqrt(a_16) # km/s 
+
+v = orbital_speed(a_gas)
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5) # yr 
+    return t
+
+P = binary_orbital_period(a_gas)
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    return psi*180.0/np.pi
+
+psi = half_opening_angle_intrinsic(a_gas)
+
+def half_opening_angle_observed(psi):
+    inclination = 15.0*np.pi/180.0
+    return psi/np.sin(inclination)
+
+#psi_observed = half_opening_angle_observed(psi)
+
+debug = True
+if debug: 
+    data = np.vstack((t,a_gas))
+    i = 1
+    while i < t.size:
+        print t[i], a_gas[i], P[i]
+        i += 1
+
+# Plot showing evolution of a_gas.
+    
+fig = plt.figure(figsize=(7, 7), dpi=100)
+ax = fig.add_subplot(1, 1, 1)
+ax.tick_params('both', which='major', length=7, width=1)
+ax.tick_params('both', which='minor', length=3, width=1)
+ax.set_xscale('log')
+ax.set_yscale('log')
+a_gas *= 0.003241 # convert to pc
+ax.plot(t,a_gas,c='k',lw=2)
+ax.set_xlim(1.0e4,1.0e8)
+ax.set_xlabel('$t$ [yr]') 
[…]

### 2015-03-05 gas_shrink.py

@@ -1,0 +1,12 @@
+
+import numpy as np
+from scipy.integrate import ode 
+
+def f(t, y):
+    return np.exp(t)
+
+t0, y0 = 0.0, 1.0
+r = ode(f).set_integrator('vode', method='bdf', with_jacobian=True)
+r.set_initial_value(y0, t0)
+print r.integrate(1.0)
+

### 2015-03-05 gw.py

@@ -1,0 +1,138 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+
+M = 1.0 # 1.0e8 Msun
+coeff = -2.56e5 
+
+t0 = 1.0e-2 # yr
+a0 = 1 # 1.0e16 cm
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a 
+    
+# t1 = np.logspace(-2.0,4.80617,num=100)
+# a1 = binary_separation_gw(t1) 
+
+# t2 = np.logspace(4.80617,4.80619,num=300)
+# a2 = binary_separation_gw(t2) 
+
+t1 = np.logspace(-2.0,3.8,num=100)
+a1 = binary_separation_gw(t1) 
+
+t2 = np.logspace(3.8,4.80619,num=100000)
+a2 = binary_separation_gw(t2) 
+
+t = np.concatenate((t1,t2))
+a = np.concatenate((a1,a2))
+
+def binary_separation_gas(t):
+    k = 1.0e7 # yr 
+    return a0*np.exp((t-t0)/k)
+
+a_gas = binary_separation_gas(t)
+
+def orbital_speed(a_16):
+    return 5.8e3 / np.sqrt(a_16) # km/s 
+
+v = orbital_speed(a) 
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    return psi*180.0/np.pi
+
+psi = half_opening_angle_intrinsic(a)
+
+def half_opening_angle_observed(psi):
+    inclination = 15.0*np.pi/180.0
+    return psi/np.sin(inclination)
+
+psi_observed = half_opening_angle_observed(psi)
+
+debug = False
+if debug: 
+    data = np.vstack((t,a_gas))
+    i = 250
+    while i < t.size:
+        print t[i], a_gas[i]
+        i += 1
+
+# Plot showing evolution of a.
+    
+fig = plt.figure(figsize=(7, 7), dpi=100)
[…]

### 2015-03-05 jet.py

@@ -1,0 +1,226 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+case = int(sys.argv[1])
+
+if case == 0:
+
+    """ Just some example jets>"""
+
+    fig = plt.figure(figsize=(7, 7), dpi=80)
+
+    ax = fig.add_subplot(1, 1, 1) 
+    
+    x = np.arange(0, 20*np.pi, 0.1)
+    y = np.sin(x)*x*np.tan(0.1)
+    plt.plot(x,y)
+    
+    y = np.sin(x)*x*np.tan(0.4)
+    plt.plot(x,y)
+
+    plt.show()
+
+elif case == 1:
+
+    """ cone angle narrowing down"""
+
+    fig = plt.figure(figsize=(7, 7), dpi=80)
+    ax = fig.add_subplot(1, 1, 1)
+
+    x = np.arange(0, 20*np.pi, 0.1)
+
+    y = np.sin(x)*x
+    for i, s in enumerate(y):
+        rate = 0.02
+        angle = rate*x[i]
+        y[i] = s*np.tan(angle)
+    plt.plot(x,y)
+    
+    plt.show()
+
+elif case == 2:
+
+    """ cone angle narrowing down with differing rates"""
+
+    fig = plt.figure(figsize=(7, 7), dpi=80)
+    ax = fig.add_subplot(1, 1, 1)
+
+    x = np.arange(0, 20*np.pi, 0.1)
+
+    y = np.sin(x)*x
+    for i, s in enumerate(y):
+        if x[i] > 45.0: 
+            rate = 0.02
+        else:
+            rate = 0.01
+        angle = rate*x[i]
+        y[i] = s*np.tan(angle)
+            
+    plt.plot(x,y)
+    
+    plt.show()
+    
+elif case == 3:
+
+    """ different periods """ 
+
+    fig = plt.figure(figsize=(7, 7), dpi=80)
+    ax = fig.add_subplot(1, 1, 1)
+
+    x = np.arange(0, 20*np.pi, 0.1)
+    y = np.zeros(x.size)
+
+    for i, s in enumerate(y):
+        angle = 0.1
+        p = 0.5
[…]

### 2015-03-05 jet3d.py

@@ -1,0 +1,105 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+psi = 20.0 # degrees
+i = 40.0 # degrees
+theta = 0.0 # degrees
+
+output_filename = 'jet_i%2d_psi%2d.pdf' % (i, psi)
+
+psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+theta *= np.pi/180.0 # radians
+
+beta = 0.90
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+d = 100.0 # Mpc 
+
+t = np.logspace(-1.0,2.0,1000) # yr 
+
+def Omega(t):
+    period = 10.0 # yr
+    t = t%period
+    return 2.0*np.pi*t/period # radians
+
+azimuth = Omega(t) 
+
+def vel(t):
+    # We are following geometry from Gower et al. Figure 1. 
+    vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(t)) + np.cos(psi)*np.cos(i))
+    vy = beta*c*np.sin(psi)*np.sin(Omega(t))
+    vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(t)))
+    return vx, vy, vz # km/s
+
+velx, vely, velz = vel(t)
+
+x = np.zeros(t.size)
+y = np.zeros(t.size)
+z = np.zeros(t.size)
+time_prev = t[0]
+vx_prev, vy_prev, vz_prev = vel(time_prev)
+pos_x = 0.0
+pos_y = 0.0
+pos_z = 0.0
+x[0] = pos_x
+y[0] = pos_y
+z[0] = pos_z
+t_counter = 0 
+prevtime = t[t_counter]
+
+phi_z = np.zeros(t.size) 
+phi_y = np.zeros(t.size)
+phi_z[t_counter] = 0.0
+phi_y[t_counter] = 0.0
+angpos_z = 0.0
+angpos_y = 0.0
+
+for time in t[1:]:
[…]

### 2015-03-05 jet3d.py
Documentation.

@@ -11,27 +11,27 @@ import sys
 psi = 20.0 # degrees
 i = 40.0 # degrees
 theta = 0.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
 
 output_filename = 'jet_i%2d_psi%2d.pdf' % (i, psi)
+save_pdf = False 
 
 psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
 theta *= np.pi/180.0 # radians
 
-beta = 0.90
 c = 3.0e5 # km/s; speed of light
 yrbys = 3.154e7
 kpcbykm = 3.086e16
-d = 100.0 # Mpc 
 
 t = np.logspace(-1.0,2.0,1000) # yr 
 
 def Omega(t):
+    # Angular velocity of jet precession 
     period = 10.0 # yr
     t = t%period
     return 2.0*np.pi*t/period # radians
-
-azimuth = Omega(t) 
 
 def vel(t):
     # We are following geometry from Gower et al. Figure 1. 
@@ -45,61 +45,18 @@ velx, vely, velz = vel(t)
 x = np.zeros(t.size)
 y = np.zeros(t.size)
 z = np.zeros(t.size)
-time_prev = t[0]
-vx_prev, vy_prev, vz_prev = vel(time_prev)
-pos_x = 0.0
-pos_y = 0.0
-pos_z = 0.0
-x[0] = pos_x
-y[0] = pos_y
-z[0] = pos_z
-t_counter = 0 
-prevtime = t[t_counter]
 
-phi_z = np.zeros(t.size) 
-phi_y = np.zeros(t.size)
-phi_z[t_counter] = 0.0
-phi_y[t_counter] = 0.0
-angpos_z = 0.0
-angpos_y = 0.0
-
-for time in t[1:]:
-    t_counter += 1 
-    vx_curr, vy_curr, vz_curr = vel(time)
-
-    pos_x += (time-prevtime)*(vx_curr+vx_prev)*0.5*yrbys/kpcbykm # kpc
-    pos_y += (time-prevtime)*(vy_curr+vy_prev)*0.5*yrbys/kpcbykm # kpc
-    pos_z += (time-prevtime)*(vz_curr+vz_prev)*0.5*yrbys/kpcbykm # kpc 
-
-    x[t_counter] = pos_x
-    y[t_counter] = pos_y
[…]

### 2015-03-06 jet3d.py
Change to observer frame

@@ -14,8 +14,10 @@ theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
-output_filename = 'jet_i%2d_psi%2d.pdf' % (i, psi)
-save_pdf = False 
+d *= 1.0e3 # kpc
+
+output_filename = 'jet_i%2d_psi%2d' % (i, psi)
+save_pdf = False
 
 psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
@@ -25,20 +27,22 @@ c = 3.0e5 # km/s; speed of light
 yrbys = 3.154e7
 kpcbykm = 3.086e16
 
-t = np.logspace(-1.0,2.0,1000) # yr 
+t = np.logspace(-1.0,2.0,10**4) # yr
 
 def Omega(t):
     # Angular velocity of jet precession 
     period = 10.0 # yr
-    t = t%period
     return 2.0*np.pi*t/period # radians
 
 def vel(t):
-    # We are following geometry from Gower et al. Figure 1. 
+    # We are following geometry from Gower et al. Figure 1.
     vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(t)) + np.cos(psi)*np.cos(i))
     vy = beta*c*np.sin(psi)*np.sin(Omega(t))
     vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(t)))
-    return vx, vy, vz # km/s
+    return sign*vx, sign*vy, sign*vz # km/s
+
+# sign sets Forward or backward jet
+sign = 1
 
 velx, vely, velz = vel(t)
 
@@ -49,14 +53,64 @@ z = np.zeros(t.size)
 y = vely*t*yrbys/kpcbykm # kpc
 z = velz*t*yrbys/kpcbykm # kpc
 
+y_obs = y/(1.0-velx/c)
+z_obs = z/(1.0-velx/c)
+
+phi_y_obs = y_obs/d * 180.0/np.pi * 3600.0 # arcsec
+phi_z_obs = z_obs/d * 180.0/np.pi * 3600.0 # arcsec 
+
+sign = -1
+
+velx, vely, velz = vel(t)
+
[…]

### 2015-03-06 scratch.py
Scratch code

@@ -1,0 +1,112 @@
+time_prev = t[0]
+vx_prev, vy_prev, vz_prev = vel(time_prev)
+pos_x = 0.0
+pos_y = 0.0
+pos_z = 0.0
+x[0] = pos_x
+y[0] = pos_y
+z[0] = pos_z
+t_counter = 0 
+prevtime = t[t_counter]
+
+phi_z = np.zeros(t.size) 
+phi_y = np.zeros(t.size)
+phi_z[t_counter] = 0.0
+phi_y[t_counter] = 0.0
+angpos_z = 0.0
+angpos_y = 0.0
+
+for time in t[1:]:
+    t_counter += 1 
+    vx_curr, vy_curr, vz_curr = vel(time)
+
+    pos_x += (time-prevtime)*(vx_curr+vx_prev)*0.5*yrbys/kpcbykm # kpc
+    pos_y += (time-prevtime)*(vy_curr+vy_prev)*0.5*yrbys/kpcbykm # kpc
+    pos_z += (time-prevtime)*(vz_curr+vz_prev)*0.5*yrbys/kpcbykm # kpc 
+
+    x[t_counter] = pos_x
+    y[t_counter] = pos_y
+    z[t_counter] = pos_z
+
+    prevtime = time
+    vx_prev = vx_curr
+    vy_prev = vy_curr
+    vz_prev = vz_curr
+
+t_observer = t/(1.0-velx/c) # yr
+
+
+# now calculate anglar motion on the sky 
+phi_z = np.zeros(t.size) 
+phi_y = np.zeros(t.size)
+time_prev = t[0]
+vx_prev, vy_prev, vz_prev = vel(time_prev)
+
+def binary_separation_gw(t):
+    coeff = -2.56e5 
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a 
+
+def half_opening_angle_intrinsic_old(a_16):
+    psi0 = 1.0*np.pi/180.0 # radians 
+    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    return psi*180.0/np.pi # degrees
+
+def half_opening_angle_intrinsic(t):
+    y = 0.680681*t+1.93193 # degrees
+    return y 
+
+
+t0 = t[0]
+a0 = 1 # 1.0e16 cm
+
+a = binary_separation_gw(t)
+psi_intrinsic = half_opening_angle_intrinsic(a)
+
+def psi_of_t(time):
[…]

### 2015-03-06 gw.py
Plot the chirp

@@ -56,6 +56,13 @@ def half_opening_angle_observed(psi):
 
 psi_observed = half_opening_angle_observed(psi)
 
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5) # yr 
+    return t
+
+P = binary_orbital_period(a)
+
+
 debug = False
 if debug: 
     data = np.vstack((t,a_gas))
@@ -64,75 +71,100 @@ if debug:
         print t[i], a_gas[i]
         i += 1
 
-# Plot showing evolution of a.
+all = False
+if all: 
+        
+    # Plot showing evolution of a.
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    ax = fig.add_subplot(1, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    a *= 0.003241 
+    ax.plot(t,a,c='k',lw=2)
+    ax.set_xlim(1.0e1,1.0e5)
+    ax.set_xlabel('$t$ [yr]',labelpad=15) 
+    ax.set_ylabel('$a$ [pc]')
+    plt.savefig("a_gw.pdf",bbox_inches='tight')
+    #plt.show()
+
+    # Plot showing evolution of a_gas.
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    ax = fig.add_subplot(1, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    ax.plot(t,a_gas)
+    ax.set_xlabel('$t$ [yr]') 
+    ax.set_ylabel('$a$ [10^{16} cm]')
+
+    # Plot showing evolution of v_orb.
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    ax = fig.add_subplot(1, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    ax.plot(t,v,c='k',lw=2)
+    ax.set_xlim(1.0e1,1.0e5)
[…]

### 2015-03-06 jet3d.py

@@ -8,18 +8,21 @@ import numpy as np
 import math 
 import sys
 
-psi = 20.0 # degrees
+#psi = 20.0 # degrees
 i = 40.0 # degrees
 theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
+a0 = 1 # 1.0e16 cm
+coeff = -2.56e5
+
 d *= 1.0e3 # kpc
 
-output_filename = 'jet_i%2d_psi%2d' % (i, psi)
+#output_filename = 'jet_i%2d_psi%2d' % (i, psi)
 save_pdf = False
 
-psi *= np.pi/180.0 # radians 
+#psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
 theta *= np.pi/180.0 # radians
 
@@ -27,18 +30,56 @@ c = 3.0e5 # km/s; speed of light
 yrbys = 3.154e7
 kpcbykm = 3.086e16
 
-t = np.logspace(-1.0,2.0,10**4) # yr
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
 
-def Omega(t):
-    # Angular velocity of jet precession 
-    period = 10.0 # yr
-    return 2.0*np.pi*t/period # radians
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5) # yr 
+    return t
 
-def vel(t):
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 30.0*np.pi/180.0 # radians 
+    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    return psi#*180.0/np.pi # degrees 
+
+case = 1
+
+if case == 0: 
+    t = np.logspace(-1.0,2.0,10**4) # yr
+    t0 = t[0]
+    print t.size
+
+elif case == 1: 
+    t1 = np.logspace(-2.0,3.8,num=100)
+    t2 = np.logspace(3.8,4.80619,num=100000)
+    tb = np.concatenate((t1,t2))
+    t0 = tb[0]
+    a = binary_separation_gw(tb)
+    pb = binary_orbital_period(a)
[…]

### 2015-03-06 scratch.py

@@ -110,3 +110,6 @@ ax.plot(t,psi_intrinsic,c='k',lw=1)
 ax.set_xlabel('kpc',labelpad=15)
 ax.set_ylabel('kpc',labelpad=15)
 
+    t1 = np.logspace(-2.0,3.8,num=100)
+    t2 = np.logspace(3.8,4.80619,num=100000)
+    t = np.concatenate((t1,t2))

### 2015-03-06 jet3d.py
Change psi

@@ -8,7 +8,7 @@ import numpy as np
 import math 
 import sys
 
-psi = 20.0 # degrees
+psi = 63.0 # degrees
 i = 40.0 # degrees
 theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 

### 2015-03-06 jet3d.py
Debug

@@ -40,7 +40,7 @@ def binary_orbital_period(a_16):
     return t
 
 def half_opening_angle_intrinsic(a_16):
-    psi0 = 30.0*np.pi/180.0 # radians 
+    psi0 = 20.0*np.pi/180.0 # radians 
     psi = np.arcsin(np.sin(psi0)*a0/a_16)
     return psi#*180.0/np.pi # degrees 
 
@@ -76,6 +76,7 @@ def Omega(time):
 def vel(time):
     idx = np.abs(t-time).argmin()
     psi = psic[-idx]
+    print time, psi 
     # We are following geometry from Gower et al. Figure 1.
     vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(time-t0)) + np.cos(psi)*np.cos(i))
     vy = beta*c*np.sin(psi)*np.sin(Omega(time-t0))
@@ -84,6 +85,9 @@ def vel(time):
 
 # sign sets Forward or backward jet
 sign = 1
+
+for x in t:
+    v1, v2, v3 = vel(x) 
 
 velx, vely, velz = vel(t)
 

### 2015-03-11 jet3d.py
Change parameter values

@@ -8,16 +8,16 @@ import numpy as np
 import math 
 import sys
 
-psi = 63.0 # degrees
+psi = 20.0 # degrees
 i = 40.0 # degrees
 theta = 0.0 # degrees
-beta = 0.90 # jet speed in units of c 
+beta = 0.20 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
 d *= 1.0e3 # kpc
 
-output_filename = 'jet_i%2d_psi%2d' % (i, psi)
-save_pdf = False
+output_filename = 'jet_i%2d_psi%2d_beta_%3.2f' % (i, psi, beta)
+save_pdf = True
 
 psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 

### 2015-03-11 Paper/paper.tex
Add new plots to paper

@@
 
+\clearpage
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet structure in the source frame. $\beta=0.9, 0.8, 0.4$ from left to right.}
+\label{fig:srcframe}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_obs.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_obs.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_obs.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet structure in the observer frame. $\beta=0.9, 0.8, 0.4$ from left to right.}
+\label{fig:obsframe}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_ang.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_ang.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_ang.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet structure in angular coordinates. $\beta=0.9, 0.8, 0.4$ from left to right.}
+\label{fig:ang}
+\end{figure*}
+
 \bibliography{refs}

### 2015-03-12 jet3d_check.py
Just a checking copy

@@ -1,0 +1,102 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '11'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+psi = 20.0 # degrees
+i = 40.0 # degrees
+theta = 0.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+a0 = 1 # 1.0e16 cm
+coeff = -2.56e5
+
+d *= 1.0e3 # kpc
+
+output_filename = 'jet_i%2d_psi%2d' % (i, psi)
+save_pdf = False
+
+psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+theta *= np.pi/180.0 # radians
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 20.0*np.pi/180.0 # radians 
+    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    return psi#*180.0/np.pi # degrees 
+
+case = 1
+
+if case == 0:
+    # Simple sampling
+    t = np.logspace(-1.0,2.0,10**4) # yr
+    t0 = t[0]
+    print t.size
+
+elif case == 1:
+    # Sampling increases at small separations
+    t1 = np.logspace(-2.0,3.8,num=100)
+    t2 = np.logspace(3.8,4.80619,num=100000)
+    tb = np.concatenate((t1,t2))
+    t0 = tb[0]
+    a = binary_separation_gw(tb)
+    pb = binary_orbital_period(a)
+    psi_chirp = half_opening_angle_intrinsic(a)
+
+    t = tb[-100000:-1000]
+    maxt = t.max()
+    t = maxt - t
[…]

### 2015-03-12 gw.py
Still debugging

@@ -2,7 +2,7 @@ import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
 mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
+mpl.rcParams['font.size'] = '11'
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -71,8 +71,8 @@ if debug:
         print t[i], a_gas[i]
         i += 1
 
-all = False
-if all: 
+case = 2
+if case == 0: 
         
     # Plot showing evolution of a.
 
@@ -143,7 +143,7 @@ if all:
     ax.set_xlabel('$t$ [yr]') 
     ax.set_ylabel(r'$\psi$')
 
-else:
+elif case == 1:
 
     fig = plt.figure(figsize=(7, 7), dpi=100)
     ax = fig.add_subplot(1, 1, 1)
@@ -164,7 +164,18 @@ else:
         i += 1
     
     ax.plot(mt,P[-100000:-1000])
-    
     ax.set_xlabel('$t$ [yr]') 
 
+elif case == 2: 
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    ax = fig.add_subplot(1, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    ax.plot(t,P,c='k',lw=1)
+    ax.set_xlabel('$t$ [yr]')
+    ax.set_ylabel('period [yr]') 
+    
 plt.show()

### 2015-03-12 jet3d.py

@@ -8,7 +8,7 @@ import numpy as np
 import math 
 import sys
 
-#psi = 20.0 # degrees
+psi = 20.0 # degrees
 i = 40.0 # degrees
 theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
@@ -22,7 +22,7 @@ d *= 1.0e3 # kpc
 #output_filename = 'jet_i%2d_psi%2d' % (i, psi)
 save_pdf = False
 
-#psi *= np.pi/180.0 # radians 
+psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
 theta *= np.pi/180.0 # radians
 
@@ -75,7 +75,7 @@ def Omega(time):
 
 def vel(time):
     idx = np.abs(t-time).argmin()
-    psi = psic[-idx]
+    #psi = psic[-idx]
     print time, psi 
     # We are following geometry from Gower et al. Figure 1.
     vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(time-t0)) + np.cos(psi)*np.cos(i))

### 2015-03-12 jet3d_check.py

@@ -41,8 +41,8 @@ def binary_orbital_period(a_16):
 
 def half_opening_angle_intrinsic(a_16):
     psi0 = 20.0*np.pi/180.0 # radians 
-    psi = np.arcsin(np.sin(psi0)*a0/a_16)
-    return psi#*180.0/np.pi # degrees 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
+    return angle #*180.0/np.pi # degrees 
 
 case = 1
 
@@ -62,38 +62,49 @@ elif case == 1:
     pb = binary_orbital_period(a)
     psi_chirp = half_opening_angle_intrinsic(a)
 
-    t = tb[-100000:-1000]
+    t = tb[-100000:-5]
     maxt = t.max()
     t = maxt - t
-    p = pb[-100000:-1000]
-    psic = psi_chirp[-100000:-1000]
+    p = pb[-100000:-5]
+    psic = psi_chirp[-100000:-5]
     
+def Omega(time):
+    # Angular velocity times time for jet precession 
+    idx = np.abs(t-time).argmin()
+    period = p[idx]
+    return 2.0*np.pi*time/period # radians
+
+def vel(time):
+    # We are following geometry from Gower et al. Figure 1.
+    vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(time)) + np.cos(psi)*np.cos(i))
+    vy = beta*c*np.sin(psi)*np.sin(Omega(time))
+    vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(time)))
+    return sign*vx, sign*vy, sign*vz # km/s
+
+# sign sets Forward or backward jet
+sign = 1
+
+velx, vely, velz = vel(t)
+
+y = vely*t*yrbys/kpcbykm # kpc
+z = velz*t*yrbys/kpcbykm # kpc
+
+y_obs = y/(1.0-velx/c)
+z_obs = z/(1.0-velx/c)
+
+phi_y_obs = y_obs/d * 180.0/np.pi * 3600.0 # arcsec
+phi_z_obs = z_obs/d * 180.0/np.pi * 3600.0 # arcsec
+
 fig = plt.figure(figsize=(7, 7), dpi=100)
-ax = fig.add_subplot(1, 1, 1)
[…]

### 2015-03-12 scratch.py

@@ -113,3 +113,23 @@ ax.set_ylabel('kpc',labelpad=15)
     t1 = np.logspace(-2.0,3.8,num=100)
     t2 = np.logspace(3.8,4.80619,num=100000)
     t = np.concatenate((t1,t2))
+
+    sign = -1
+
+velx, vely, velz = vel(t)
+
+yb = vely*t*yrbys/kpcbykm # kpc
+zb = velz*t*yrbys/kpcbykm # kpc
+
+y_obsb = yb/(1.0-velx/c)
+z_obsb = zb/(1.0-velx/c)
+
+phi_y_obsb = y_obsb/d * 180.0/np.pi * 3600.0 # arcsec
+phi_z_obsb = z_obsb/d * 180.0/np.pi * 3600.0 # arcsec 
+
+ax.plot(phi_z_obs,phi_y_obs,c='k',lw=1)
+ax.plot(phi_z_obsb,phi_y_obsb,c='k',lw=1)
+ax.set_xlabel('arcsec',labelpad=15)
+ax.set_ylabel('arcsec',labelpad=15)
+ax.set_xlim(40.60,40.6+1.0e-20)
+# ax.set_ylim(-1.0,1.0)

### 2015-03-12 gw.py
Implement proper psi and P evolution

@@ -44,10 +44,10 @@ def orbital_speed(a_16):
 v = orbital_speed(a) 
 
 def half_opening_angle_intrinsic(a_16):
-    psi0 = 2.0*np.pi/180.0 # radians 
     psi = np.arcsin(np.sin(psi0)*a0/a_16)
     return psi*180.0/np.pi
 
+psi0 = 2.0*np.pi/180.0 # radians 
 psi = half_opening_angle_intrinsic(a)
 
 def half_opening_angle_observed(psi):
@@ -71,7 +71,7 @@ if debug:
         print t[i], a_gas[i]
         i += 1
 
-case = 2
+case = 3
 if case == 0: 
         
     # Plot showing evolution of a.
@@ -177,5 +177,22 @@ elif case == 2:
     ax.plot(t,P,c='k',lw=1)
     ax.set_xlabel('$t$ [yr]')
     ax.set_ylabel('period [yr]') 
+
+elif case == 3:
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    ax = fig.add_subplot(1, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    #ax.set_yscale('log')
+    ax.plot(t,psi,c='k',lw=1)
+
+    psi0 = 20.0*np.pi/180.0 # radians 
+    psi = half_opening_angle_intrinsic(a)
+    ax.plot(t,psi,c='r',lw=1)
+    
+    ax.set_xlabel('$t$ [yr]')
+    ax.set_ylabel('half-opening angle [degrees]') 
     
 plt.show()

### 2015-03-12 jet3d_check2.py

@@ -1,0 +1,159 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '11'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+theta = 0.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+a0 = 1 # 1.0e16 cm
+coeff = -2.56e5
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f' % (i, beta)
+save_pdf = True
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+theta *= np.pi/180.0 # radians
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
+    return angle #*180.0/np.pi # degrees 
+
+case = int(sys.argv[1])
+if case == 0: 
+    t = np.logspace(-2.0,2.0,10000000)
+    output_filename += '_full'
+elif case == 1:
+    t = np.linspace(10.0,11.0,10000)
+    output_filename += '_zoom1'
+elif case == 2:
+    t = np.linspace(1.0,2.0,10000)
+    output_filename += '_zoom2'
+
+t0 = t[0]
+
+def t_binary(time):
+    return np.abs(time-64001.477505390176) # yr
+
[…]

### 2015-03-12 jet3d_zoom.py

@@ -20,7 +20,7 @@ coeff = -2.56e5
 d *= 1.0e3 # kpc
 
 output_filename = 'jet_i%2d_psi%2d' % (i, psi)
-save_pdf = False
+save_pdf = True
 
 psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
@@ -44,34 +44,12 @@ def half_opening_angle_intrinsic(a_16):
     angle = np.arcsin(np.sin(psi0)*a0/a_16)
     return angle #*180.0/np.pi # degrees 
 
-case = 1
+def t_binary(time):
+    return np.abs(time-64001.477505390176) # yr
 
-if case == 0:
-    # Simple sampling
-    t = np.logspace(-1.0,2.0,10**4) # yr
-    t0 = t[0]
-    print t.size
-
-elif case == 1:
-    # Sampling increases at small separations
-    t1 = np.logspace(-2.0,3.8,num=100)
-    t2 = np.logspace(3.8,4.80619,num=100000)
-    tb = np.concatenate((t1,t2))
-    t0 = tb[0]
-    a = binary_separation_gw(tb)
-    pb = binary_orbital_period(a)
-    psi_chirp = half_opening_angle_intrinsic(a)
-
-    t = tb[-100000:-5]
-    maxt = t.max()
-    t = maxt - t
-    p = pb[-100000:-5]
-    psic = psi_chirp[-100000:-5]
-    
 def Omega(time):
     # Angular velocity times time for jet precession 
-    idx = np.abs(t-time).argmin()
-    period = p[idx]
+    period = binary_orbital_period(binary_separation_gw(t_binary(time))) # yr 
     return 2.0*np.pi*time/period # radians
 
 def vel(time):
@@ -81,33 +59,59 @@ def vel(time):
     vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(time)))
     return sign*vx, sign*vy, sign*vz # km/s
 
-# sign sets Forward or backward jet
+fig = plt.figure(figsize=(7, 7), dpi=100)
[…]

### 2015-03-25 Paper/paper.tex
Add chirps to paper

@@
 
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology without accounting for opening angle
+    evolution}
+\label{fig:jet_nopsi}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology with opening angle evolution.}
+\label{fig:jet_psi}
+\end{figure*}
+
 \bibliography{refs}

### 2015-03-25 jet3d_check2.py
Change units of a

@@ -14,7 +14,9 @@ theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
-a0 = 1 # 1.0e16 cm
+a0 = 0.003241 # pc
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
 coeff = -2.56e5
 
 d *= 1.0e3 # kpc
@@ -102,6 +104,7 @@ phi_z_obsb = z_obsb/d * 180.0/np.pi * 3600.0 # arcsec
 
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
+#ax.set_ylim(-0.01,0.01)
 ax.plot(phi_z_obs,phi_y_obs,c='k',lw=1,rasterized=True)
 ax.plot(phi_z_obsb,phi_y_obsb,c='k',lw=1,rasterized=True)
 ax.set_xlabel('arcsec',labelpad=15)

### 2015-03-25 jet3d_check2.py
Add mass dependence

@@ -14,10 +14,11 @@ theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
-a0 = 0.003241 # pc
+M = 1.0e8 # Msun; total mass of the equal-mass binary 
+a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5
+coeff = -2.56e5/(M*1.0e-8)**3 
 
 d *= 1.0e3 # kpc
 
@@ -39,7 +40,7 @@ def binary_separation_gw(t):
     return a
 
 def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5) # yr 
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
     return t
 
 def half_opening_angle_intrinsic(a_16):

### 2015-03-25 scratch.py

@@ -133,3 +133,18 @@ ax.set_xlabel('arcsec',labelpad=15)
 ax.set_ylabel('arcsec',labelpad=15)
 ax.set_xlim(40.60,40.6+1.0e-20)
 # ax.set_ylim(-1.0,1.0)
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+theta = 0.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+#a0 = 0.003241 # pc
+M = 1.0e8 # Msun; total mass of the equal-mass binary 
+a0 = 1.0e-3*(M*1.0e-8)**(3./4.) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 10^16 cm
+coeff = -2.56e5/(M*1.0e-8)**3 
+
+d *= 1.0e3 # kpc

### 2015-03-25 gw.py
Add Mdot

@@ -6,17 +6,26 @@ mpl.rcParams['font.size'] = '11'
 import matplotlib.pyplot as plt
 import numpy as np
 
-M = 1.0 # 1.0e8 Msun
-coeff = -2.56e5 
+M = 1.0e8 # Msun; total mass of the equal-mass binary 
+
+pcto_10to16cm = 0.003241
+Mdot = 10.0 # Eddington units
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3 
 
 t0 = 1.0e-2 # yr
-a0 = 1 # 1.0e16 cm
 
 def binary_separation_gw(t):
     a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
     a = a**(1./4.)
     return a 
-    
+
+t_merge=coeff*a0**4/4.0
+print 'coeff=', coeff
+print 't_merge=', t_merge
+print 'log(t_merge)=', np.log10(-t_merge)
+
 # t1 = np.logspace(-2.0,4.80617,num=100)
 # a1 = binary_separation_gw(t1) 
 
@@ -26,7 +35,8 @@ def binary_separation_gw(t):
 t1 = np.logspace(-2.0,3.8,num=100)
 a1 = binary_separation_gw(t1) 
 
-t2 = np.logspace(3.8,4.80619,num=100000)
+tf = np.log10(-t_merge)
+t2 = np.logspace(3.8,tf,num=100000)
 a2 = binary_separation_gw(t2) 
 
 t = np.concatenate((t1,t2))
@@ -57,7 +67,7 @@ def half_opening_angle_observed(psi):
 psi_observed = half_opening_angle_observed(psi)
 
 def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5) # yr 
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
     return t
 
 P = binary_orbital_period(a)
@@ -71,7 +81,7 @@ if debug:
         print t[i], a_gas[i]
         i += 1
 
-case = 3
+case = 4
 if case == 0: 
         
     # Plot showing evolution of a.
@@ -194,5 +204,19 @@ elif case == 3:
     
     ax.set_xlabel('$t$ [yr]')
     ax.set_ylabel('half-opening angle [degrees]') 
[…]

### 2015-03-25 jet3d_check2.py

@@ -10,12 +10,13 @@ import sys
 
 #psi = 20.0 # degrees
 i = 40.0 # degrees
-theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
-M = 1.0e8 # Msun; total mass of the equal-mass binary 
-a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc 
+M = 1.0e10 # Msun; total mass of the equal-mass binary
+Mdot = 0.1 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
 coeff = -2.56e5/(M*1.0e-8)**3 
@@ -23,12 +24,11 @@ coeff = -2.56e5/(M*1.0e-8)**3
 d *= 1.0e3 # kpc
 
 #output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
-output_filename = 'jet_i%2d_beta%3.2f' % (i, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
 save_pdf = True
 
 #psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
-theta *= np.pi/180.0 # radians
 
 c = 3.0e5 # km/s; speed of light
 yrbys = 3.154e7
@@ -53,7 +53,9 @@ if case == 0:
     t = np.logspace(-2.0,2.0,10000000)
     output_filename += '_full'
 elif case == 1:
-    t = np.linspace(10.0,11.0,10000)
+    #t = np.linspace(10.0,14.0,10000)
+    t = np.logspace(-2.0,2.0,10000000)
+    t = t[6000000:9500000]
     output_filename += '_zoom1'
 elif case == 2:
     t = np.linspace(1.0,2.0,10000)
@@ -62,7 +64,9 @@ elif case == 2:
 t0 = t[0]
 
 def t_binary(time):
-    return np.abs(time-64001.477505390176) # yr
+    t_merge=abs(coeff)*a0**4/4.0
+    return np.abs(time-t_merge) # yr
[…]

### 2015-04-03 Paper/paper.tex
Add latest plots

@@
 
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology with opening angle evolution.}
+\label{fig:jet_psi}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology with opening angle evolution.}
+\label{fig:jet_psi}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.0_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.0_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=1$}
+\label{fig:jet_mdot1.0}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot10.00_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot10.00_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=10$}
+\label{fig:jet_10.0}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot0.10_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot0.10_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=0.1$}
+\label{fig:jet_mdot0.1}
+\end{figure*}
+
 \bibliography{refs}

### 2015-04-03 jet3d_image.py
Make jet images

@@ -1,0 +1,140 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '11'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+from matplotlib import cm
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+M = 1.0e10 # Msun; total mass of the equal-mass binary
+Mdot = 1.0 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3 
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+save_pdf = True
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
+    return angle #*180.0/np.pi # degrees 
+
+case = int(sys.argv[1])
+if case == 0: 
+    t = np.logspace(-2.0,2.0,10000000)
+    output_filename += '_full'
+elif case == 1:
[…]

### 2015-07-06 Paper/paper.tex
Made more changes suggested by Avi.

--- 2015-03-11 Paper/paper.tex
@@
-% File: paper.tex 
-%  Cre: 2013-09-12 
-%  $Id: paper.tex 4261 2015-01-27 23:49:04Z jhennawi $
-
 \documentclass[iop]{emulateapj}
@@
 
-\shorttitle{Pressure smoothing in nonlinear IGM}
+\shorttitle{Chirping Jets from Black Hole Binaries}
 \shortauthors{Kulkarni et al.}
@@
 
-\title{Flagging gravitational wave sources from radio jet morphology}
-\author{Girish Kulkarni\altaffilmark{1} and
-  Abraham Loeb\altaffilmark{2}}
-\altaffiltext{1}{Institute of Astronomy, University of Cambridge,
-  Madingley Road, Cambridge CB3 0HA, UK;
-  \email{kulkarni@ast.cam.ac.uk}}
-\altaffiltext{2}{Harvard}
+\title{Radio Crickets: Chirping Jets from Black Hole Binaries Entering
+  their Gravitational Wave Inspiral} \author{Girish
+  Kulkarni\altaffilmark{1} and Abraham Loeb\altaffilmark{2}}
+\altaffiltext{1}{Institute of Astronomy and Kavli Institute of
+  Cosmology, University of Cambridge, Madingley Road, Cambridge CB3
+  0HA, UK; \email{kulkarni@ast.cam.ac.uk}} \altaffiltext{2}{Institute
+  for Theory \& Computation, Harvard University, 60 Garden Street,
+  Cambridge, MA 02138, USA; \email{aloeb@cfa.harvard.edu}}
 
 \begin{abstract}
-Apart from a stochastic gravitational wave (GW) background, LISA will
-also be able to detect GW emission from binary black holes. The number
-of such binary sources depends on the major merger rates of gas-rich
-galaxies and subsequent evolution of their central black holes.  The
-latter of these is uncertain.  Therefore, an alternate method of
-flagging GW sources is useful.  Here we show that if one of the binary
-BHs is accreting and has a jet, then the dynamical evolution of the
-binary is imprinted on the jet morphology.  Using this dependance, it
-is possible to flag binary BHs that are entering their GW inspiral
-phase in the present epoch.  We further show that this relationship
-between the binary dynamics and the jet morphology has two other
-applications: (1) at large binary separations the Mpc-scale jet
-morphology can probe the acts as a probe of the pc-scale gas supply to
-the binary, and (2) at smaller separations the jet morphology acts as
-a test of GR.
+Apart from a stochastic gravitational wave (GW) background, eLISA will
+also be able to detect GW emission from binary black holes.  Combined
+GW and electromagnetic observations of compact binary mergers should
+enable detailed studies of astrophysical processes in the strong-field
+gravity regime.  Here we show that if one of component BHs of the
+binary has an accretion disk with an associated jet, the morphology of
+the jet can act as an electromagnetic counterpart to the GW signal
+from the binary.  For binaries with mass $10^7$--$10^{10}$ M$_\odot$
+at redshifts $z<0.5$, the jet precesses along a biconical surface due
+to the binary's orbital motion.  When the binary enters the GW phase
+of its evolution, the opening angle widens, the jet exhibits
+milliarcsecond scale wiggles and the conical surface of jet precession
+is twisted due to the superluminal motion.  These morphological
+features can be used to constrain the dyanamical parameters of the
+binary.  Monitoring of jets with these features in current and
+archival data will put a lower limit on the sources that could be
+detected by GW experiments such as eLISA and Pulsar Timing Arrays
+(PTAs).  In the future, microarcsecond interferometry with experiments
+such as SKA will increase the potential usefulness of this technique.  
 \end{abstract}
 
-\keywords{cosmology: IGM}
+\keywords{BHs}
 
@@
 
-Consider two black holes with masses $M_1$ and $M_2$ in a circular
-orbit of radius $a$ around each other \citep{2010PhRvD..81d7503L}.
-Let $M_1=M_2$ (assume equal masses for simplicity).  The orbital speed
-of each black hole is given by
+Binaries of supermassive black holes (BHs) arise naturally as a result
+of mergers of galaxies in the context of the heirarchical galaxy
+formation \citep{2000MNRAS.311..576K, 2002MNRAS.336L..61H,
+  2008ApJ...676...33D, 2012MNRAS.422.1306K}.  After a galaxy merger,
+infalling black holes lose their angular momentum in three stages
+before coalescence \citep{1980Natur.287..307B, 2005LRR.....8....8M,
+  2011ASL.....4..181C}.  In the first stage, the black holes sink to
+the center of the gravitational potential of the merger remnant due to
+dynamical friction and form a gravitationally bound binary.  The newly
+formed binary then continues to lose its angular momentum and shrink
+by scattering ambient gas and stars.  In the final stage of the
+binary's evolution, emission of gravitational waves (GWs) becomes the
+predominant mode angular momentum loss, which results in the
+coalescence of the black holes.
+
+% Any more references to add here?
+General relativisitic (GR) numerical simulations can predict the
+binary's evolution in its final stage up to coalescence
+\citep{2005PhRvL..95l1101P, 2006PhRvL..96k1102B, 2006PhRvL..96k1101C}.
+This has stimulated interest in predicting related electromagnetic
+signals of black hole coalescence.  Effects considered in the
+literature include afterglows due to infall of gas on the remnant
+black hole \citep{2005ApJ...622L..93M}, electromagnetic variability in
+the circumbinary disk due to shocks induced by the sudden mass loss of
+the binary due to GW emission at coalescence
+\citep{2007APS..APR.S1010B}, flares from shocked remnants of accretion
+disks around recoiled remnant black hole \citep{2007PhRvL..99d1103L,
+  2008ApJ...682..758S, 2009CQGra..26i4032H}, quasi-periodic
+variability in gas accretion in the early stages of black hole merger
+\citep{2008ApJ...672...83M, 2009CQGra..26i4032H, 2012MNRAS.427.2680K},
+evolution in the profiles of broad emission lines shortly before
+merger \citep{2013MNRAS.432.1468M}, characteristic evolution in
+thermal emission due to circumbinary accretion
+\citep{2015MNRAS.446L..36F}, brightening of accretion disk due to
+viscous dissipation of gravitational waves
+\citep{2008PhRvL.101d1101K}, and prompt tidal disruption of stars by
+the recoiled remnant \citep{2011MNRAS.412...75S}.
+
+% Replace `LISA' with `eLISA` or `NGO'?  Mention pulsar timing arrays?
+In addition to testing GR in the strong field limit, observations of
+electromagnetic counterparts will help localize sources of GWs for
+future detectors such as the Evolved Laser Interferometer Space
+Antenna (eLISA)\footnote{\url{https://www.elisascience.org}}, which would be
+capable of observing the peak GW luminosity for BH coalescence of
+$\sim 10^{57}$ erg$/$s up to large cosmological distances
+\citep{2003CQGra..20S..65H, 2013CQGra..30x4009S}.  Simultaneous
+detection of GW and electromagnetic signals from coalescing binaries
+will yield rich scientific pay-offs \citep{2003CQGra..20S..65H,
+  2005ApJ...629...15H}.  For example, by a redshift measurement of
+eLISA sources could constraint cosmological parameters.  Measurement
+of the black hole masses and spins from the GW signal will constrain
+accretion physics during coalescence.  But even before eLISA is
+launched, a search of electromagnetic signatures may point to the
+expected abundance of GW sources.
+
+In this Letter we study the morphology of a jet emitted by one of the
+binary members during its final stage of evolution before coalescence,
+i.e., when the binary's dynamics is dominated by GW emission.  We show
+that the helical chirping morphology of the jet can be used to infer
+the binary parameters.
+
+\section{Binary parameters}
+
+We study a BH binary formed following a gas-rich merger of two
+galaxies as a result of the processes described above.  We consider
+black holes with masses $M_1$ and $M_2$ in a circular orbit of radius
+$a$ around each other.  Let $M=M_1+M_2$.  For simplicity, we consider
+an equal-mass binary ($M_1=M_2$) on a circular orbit. We will consider
+unequal black holes masses and elliptical orbits below.  The orbital
+speed of each black hole is given by \citep{2010PhRvD..81d7503L}
 \begin{equation}
-  v_\mathrm{orb} = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2} = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
+  v = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2}\!\!\!\! = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
   \label{eqn:vorb}
 \end{equation}
-where $a_{16}\equiv a/10^{16}$ cm and $M_8\equiv M/10^8$
-M$_\odot$. Now imagine a jet coming out of one of the BHs at an angle
-$\chi$ to the direction of the orbital angular momentum.  Then because
-of the orbital motion, the jet will sweep out a cone.  This is shown
-in Figure \ref{fig:BH_orbit}.  The half opening angle of the cone is
-given by
+where $a_{16}\equiv (a/10^{16}$ cm$)$ and $M_8\equiv (M/10^8$
+M$_\odot)$ are the binary separation and mass,
+respectively.\footnote{Note that $10^{16} \mathrm{cm}\approx 3.4\times
+  10^{-3} \mathrm{pc}$.  The Schwarzchild radius of a $10^8$~M$_\odot$
+  black hole is $9.6\times 10^{-6}$~pc, or $\sim 3\times 10^{13}$~cm.}
+The orbital period is
 \begin{equation}
-  \sin\psi = \frac{v_\mathrm{orb}\cos\chi}{v_\mathrm{jet}},
+  P = 2\pi\left(\frac{GM}{a^3}\right)^{-1/2}\!\!\!\! = 1.72\, \mathrm{yr}\, a_{16}^{3/2} M_8^{-1/2}.
 \end{equation}
-where $v_\mathrm{jet}$ is the ballistic speed of the jet.  Figure
-\ref{fig:jet} shows how this looks.
-
-For observable predictions we have to project this in observer plane.
-Let the angle between the jet direction and the observer line of sight
-be $i$.  Then
-\begin{equation}
-  \psi^\mathrm{obs}\approx\frac{\psi}{\sin i}.
-\end{equation}
-
-\section{Jet structure when binary enters GW-induced inspiral}
-
-Now when the binary is at the separation at which gravitational
-wave-induced inspiral begins, it has a velocity given by Equation (5)
-of your 2009 paper:
-\begin{equation}
-  v_\mathrm{orb} = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
-  \label{eqn:v}
-\end{equation}
-where all symbols on the right hand side have the same meaning as in
-your paper.  The separation between the two black at this instant is
-given by Equation (7) of your paper:
-\begin{equation}
-  a = 8.3\times 10^{-3}\, \mathrm{pc}\, M_8^{3/4}\dot{\mathcal{M}}^{-1/4}.
-\end{equation}
-This gives
-\begin{equation}
-  \psi^\mathrm{obs}=2.42^\circ \left(\frac{v_\mathrm{jet}}{0.95 c}\right)^{-1}\cos\left(\frac{\chi}{30^{\circ}}\right)\left[\sin\left(\frac{i}{15^{\circ}}\right)\right]^{-1},
-\end{equation}
-This value in itself is probably not very interesting as it depends on
-$v_\mathrm{jet}$ and $\chi$, but it gets interesting when we consider
-the evolution of $\psi^\mathrm{obs}$ below.
+% Mention typical size of accretion disks and typical ISCO radius
+% above?
 
@@
       \fill[black] (0,0) circle (1pt);
-      %\draw [->,densely dashed] (0,0) -- (0,2.0);
-      \fill[black] (2,-0.39) circle (3pt);
-      \fill[black] (-2,0.39) circle (3pt);
-      \draw [->] (2,-0.39) -- (3.5,-0.1);
-      \draw [->,densely dashed] (2,-0.39) -- (2,1.0);
-      \draw [->] (2,-0.39) -- (2.5,0.8);
-      \draw [->,red] (2,-0.39) -- (3.5,0.8);
-      \draw (4.1,-0.1) node {${\mathbf{v}}_\mathrm{orbital}$};
-      \draw (2.7,1.0) node {${\mathbf{v}}_\mathrm{jet}$};
-      \draw (4.0,1.0) node {$\tilde{\mathbf{v}}_\mathrm{jet}$};
+      \draw [-angle 45,thick] (2,-0.39) -- (3.5,-0.1);
+      \draw [-angle 45,thick,dashed,dash pattern = on 4pt off 1pt] (2,-0.39) -- (2,1.5);
+      \draw [-angle 45,thick] (2,-0.39) -- (2.5,0.8);
+      \draw [-angle 45,thick,red] (2,-0.39) -- (3.5,0.8);
+      \fill[black] (-2,0.39) circle (2pt);
+      \fill[black] (2,-0.39) circle (2pt);
+      \draw (4.1,-0.1) node {\large ${\mathbf{v}}_\mathrm{orb}$};
+      \draw (2.7,1.0) node {\large ${\mathbf{v}}_\mathrm{jet}$};
+      \draw (4.0,1.0) node {\large $\tilde{\mathbf{v}}_\mathrm{jet}$};
+      \draw (2.0,1.8) node {\large $\mathbf{L}_\mathrm{orb}$};
     \end{tikzpicture}
   \end{center}
-  \caption{Jet and orbital configuration. The effect of
-    $\mathbf{v}_\mathrm{orbital}$ is to precess the jet as the BH
-    moves along the orbit. Note that
-    $\tilde{\mathbf{v}}_\mathrm{jet}={\mathbf{v}}_\mathrm{orbital}+{\mathbf{v}}_\mathrm{jet}$.}
+  \caption{Jet and orbital configuration. The orbital velocity
+    $\mathbf{v}_\mathrm{orb}$ causes the jet to precess as the BH
+    moves along the orbit. The net jet velocity is
+    $\tilde{\mathbf{v}}_\mathrm{jet}={\mathbf{v}}_\mathrm{orb}+{\mathbf{v}}_\mathrm{jet}$,
+    where $\mathbf{v}_\mathrm{jet}$ is its ejection velocity at the
+    source.}
   \label{fig:BH_orbit}
@@
 
-\section{Evolution of jet structure}
+In a gas-rich galaxy merger, the BH binary loses its angular
+momentum by torquing the surrounding disk through spiral arms and
+expelling gas out of a region twice as large as the binary orbit
+\citep{2005ApJ...622L..93M, 2007PASJ...59..427H, 2008ApJ...672...83M,
+  2009MNRAS.393.1423C}.  Fresh gas re-enters the region of the binary
+as a result of turbulent transport of angular momentum in the
+surrounding disk \citep{2008ApJ...672...83M, 2012ApJ...755...51N,
+  2012MNRAS.427.2680K, 2012A&A...545A.127R}.  Since the expelled gas
+carries a specific angular momentum of $\sim va$, the coalescence time
+of the binary is inversely proportional to the supply rate of fresh
+gas into the binary region.  In a steady state, this supply rate,
+$\dot M$ is proportional to the accretion rate of the circumbinary gas
+disk.  But only a fraction of gas that enters the central gap accretes
+onto the black hole and fuels quasar activity.  Therefore, it is
+appropriate to express $\dot M$ in Eddington units, $\dot{\mathcal
+  M}\equiv\dot M /\dot M_E$, where $\dot M_E$ is the accretion rate
+required to power the limiting Eddington luminosity with a radiative
+efficiency of 10\%, $\dot M_E=2.3 M_\odot\, \mathrm{yr}^{-1}M_8$.  We
+therefore have \citep{2010PhRvD..81d7503L}
+\begin{equation}
+  \frac{a}{\dot a} = t_\mathrm{gas}\approx\frac{J}{\dot Mva}=1.1\times 10^7\,
+  \mathrm{yr}\,\dot{\mathcal M}^{-1},
+\end{equation}
+where $J=\mu va$ is the binary's angular momentum and $\mu=M/2$ is the
+binary's reduced mass.  This time scale is greater by two orders of
+magnitude than the time scale for coalescence due to angular momentum
+loss by GW emission, which for circular orbits is given by
+\citep{1964PhRv..136.1224P}
+\begin{equation}
+\frac{a}{\dot a} = t_\mathrm{GW} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
+\end{equation}
+By setting $t_\mathrm{gas}=t_\mathrm{GW}$, we can get the orbital
+parameters for the binary when GWs take over.  The orbital speed at
+this stage of the binary's evolution is given by
+\begin{equation}
+  v = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
+  \label{eqn:v}
+\end{equation}
+The separation between the two black holes at this instant is given by
+\begin{equation}
+  a = 8.3\times 10^{-3}\, \mathrm{pc}\, M_8^{3/4}\dot{\mathcal{M}}^{-1/4}.
+  \label{eqn:a}
+\end{equation}
+Note that the separation is dependent on the gas accretion rate.  This
+evolution of the binary separation and orbital velocity is shown in
+Figure~\ref{fig:evolve}.
 
-Binary separation, jet velocity, and the two angles, tell us about the
-instantaneous configuration of the jet.  (Recall that we are only
-considering an equal mass binary.)  The jet evolution is governed by
-the dynamical evolution of the binary separation.  During the
-gas-dominated phase, this is given by
+\begin{figure}
+  \begin{center}
+    \begin{tikzpicture}
+      \draw [-angle 45,dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (3.0,0);
+      \draw [-angle 45,dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (0,4.0);
+      \draw [-angle 45,thick] (0,0) -- (2.0,4.0);
+      \draw [-angle 45,thick,red] (0,0) -- (5.0,4.0);
+      \draw (1.0,2.0) arc (63:90:2.2);
+      \draw (2.0,1.6) arc (38.6:63:2.56);
+      \draw (0.5,2.5) node {\large $\chi$};
+      \draw (1.90,2.3) node {\large $\psi$};
+      \draw [dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (3.0,6.0);
+      \draw (0.0,4.5) node {\large $\mathbf{L}_\mathrm{orb}$};
+      \draw (3.5,0.0) node {\large $\mathbf{v}_\mathrm{orb}$};
+      \draw (1.0,3.0) node {\large $\mathbf{v}_\mathrm{jet}$};
+      \draw (5.3,4.2) node {\large \textcolor{red}{$\mathbf{\tilde v}_\mathrm{jet}$}};
+    \end{tikzpicture}
+  \end{center}
+  \caption{The orbital velocity $\mathbf{v}_\mathrm{orbital}$
+    introduces the jet precision angle $\psi$.  Without the orbital
+    motion, $\psi=0$ and the jet traces a cylindrical surface.  With
+    the orbital motion, the net red velocity vector oscillates around
+    the solid black vector and produces a conical surface with $\psi$
+    as its half-opening angle.  Note that $v_\mathrm{orb}$ is
+    exagerated here.  In practice, $v_\mathrm{orb}\ll
+    v_\mathrm{jet}\sim c$.  This figure then shows that
+    $\sin{\psi}\sim\tan{\psi}=v_\mathrm{orb}\cos{\chi}/v_\mathrm{jet}$.}
+  \label{fig:jet}
+\end{figure}
+
+In the GW dominated phase of binary's evolution, recent simulations
+show that the BHs can continue to accrete gas from the circumbinary
+disk all the way until merger \citep{2015MNRAS.446L..36F}.  At this
+stage, we consider a ballistic jet emanating from one of binary's
+member BHs at speed $\beta=v_\mathrm{jet}/c$ at an angle $\chi$ to the
+direction of the orbital angular momentum.  This geometry is depicted
+in Figures~\ref{fig:BH_orbit} and \ref{fig:jet}.  As a result of the
+binary orbit, the jet will wiggle on a conical surface.  (In the limit
+of a small binary velocity, i.e., at large relative separations, the
+jet material will move on a cylindrical surface.)  The half-opening
+angle of the cone is given by \citep{1993ApJ...409..130R}
 \begin{equation}
-\frac{a}{\dot a} = \mu/\dot M = 1.1\times 10^7\, \mathrm{yr}\,\dot{\mathcal{M}}^{-1}.   
+  \sin\psi = \frac{v_\mathrm{orb}\cos\chi}{v_\mathrm{jet}}.
 \end{equation}
-This increases by two orders of magnitude during the GW phase to
-\begin{equation}
-\frac{a}{\dot a} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
-\end{equation}
-From Equation (\ref{eqn:vorb}), the orbital velocity of each BH evolves as
+When the binary separation is where gravitational wave-induced
+inspiral begins, it has a velocity given by equation~(\ref{eqn:v}),
+and the separation between the two black holes at this instant is
+given by equation~(\ref{eqn:a}).  Also, from equation
+(\ref{eqn:vorb}), the orbital velocity of each BH evolves as
 \begin{equation}
@@
 \end{equation}
-where $\psi_0$ and $a_0$ are the initial values. Thus as the binary
+where $\psi_0$ and $a_0$ are the initial values. Thus, as the binary
 continues to shrink the BH orbital velocity grows and the opening
-angle of the conical jet increases (cone gets wider and wider).  The
-rate at which this happens tells everything about the binary's
-dynamical evolution.
+angle of the conical jet increases (cone opens wider).  The rate at
+which this happens reflects the binary's dynamical evolution.
+Figure~\ref{fig:evolve} shows the evolution of the half-opening angle
+in the gas and GW-dominated phase.
 
-Figures \ref{fig:a_gw} show this evolution for the numbers assumed
-above.
+\begin{figure}
+  \begin{center}
+    \includegraphics*[width=\columnwidth]{evolve.pdf}
+    % gw.py (case 5) 
+  \end{center}
+  \caption{Evolution of the binary parameters.  As the binary
+    separation $a$ decreases, the orbital velocity $v_\mathrm{orb}$
+    and the opening angle $\psi$ of the precessing jet, increase.}
+  \label{fig:evolve}
+\end{figure}
 
-\section{Use jets to flag GW binaries}
+\section{Jet precession}
 
-Figure \ref{fig:a_gas} shows the evolution of the binary when it is in
-the gas accretion phase.  We have chosen steady gas accretion with
-$\dot{\mathcal{M}}=1$ here for simplicity.  Note that the starting
-separation here is about 100 pc.
-
-What we find is that in the beginning the binary orbital period is
-about $10^7$ yr. So the jet will show no signs of precision at this
-point.  However as the binary separation reduces to about 10 pc, the
-period reduces to $\sim 100$ yr and the jet will start showing
-precision.  At this point the time scale for binary merger is $\sim
-10^7$ yr.
-
-In other words \emph{since the gas accretion time scale is of the
-  order of the typical jet lifetime} the jet structure will show a
-transition from a straight jet to a precessing jet at a separation of
+We now ask if it would be possible to observe the opening of the jet
+morphology.  In order to do this, it is necessary to map the above
+calculations to the observer's frame \citep{1982ApJ...262..478G}.  At
+a large distance compared to the binary's separation, the jet geometry
+is illustrated in Figure \ref{fig:jetgeom}.  Here the source frame of
+reference is shown by $x^\prime$, $y^\prime$, $z^\prime$ coordinate
+systems.  The $x$, $y$, $z$ coordinate system represents the
+observer's frame such that the $yz$ is the plane of the sky.  Further,
+the jet is oriented such that that the $y^\prime$ axis coincides with
+the $y$ axis, the original direction of jet emission is at an angle
+$i$ relative to the line of sight ($x$-axis).  For simplicity, we also
+assume that the $z^\prime$ axis is in the $xz$ plane.  Then the jet
+will precess on a cone with a half-opening angle $\psi$ and an angular
+velocity of precession $\Omega$.  In the source frame of reference,
+the equation of the jet is then given by \citep{1981ApJ...246L.141H}
+\begin{eqnarray}
+  &v_x=&s_\mathrm{jet}\beta c\left[\sin\psi\sin i\cos\Omega t+\cos\psi\cos i\right],\\
+  &v_y=&s_\mathrm{jet}\beta c\sin\psi\sin\Omega t,~\,\mathrm{and}\\
+  &v_z=&s_\mathrm{jet}\beta c\left[\cos\psi\sin i-\sin\psi\cos i\cos\Omega t\right]. 
+\end{eqnarray}
+Here $s_\mathrm{jet}=\pm 1$ for forward and backward jets,
+respectively.  The coordinates of a jet particle in the sky is then
+given by $z=v_zt$ and $y=v_yt$.  However, because of superluminal
+motion, the observer will see this on the sky as
 \begin{equation}
-  v_\mathrm{apparant}\left(\frac{t_0}{10^7}\right).
+  z=v_zt/(1-v_x/c)
 \end{equation}
-from the binary.  \textbf{[Convert this to arcseconds on the sky.]}
-The location at which the jet morphology shows this precision and the
-opening angle evolution give a direct constraint on the epoch of
-graviational wave emission of the binary.
-
-Thus the new insight here is that the signature of the jet precision
-due to GW-inspiral is observationally accessible.
-
-\section{Jet morphology as probe of GR}
-
-Variability of lensed quasars constraints the size of accretion disks
-to about $10^{15}$ cm ($\sim 10^{-3}$ pc) \citep{2010ApJ...712.1129M}.
-This is smaller than the $a_\mathrm{GW}$ by a factor of less than 10.
-
-This means that the GW-induced precision will be visible in the jet
-precision for a very short distance near the source before the
-accretion disk is disrupted.
-
-Last plots in this note show some example jets.  In all these plots
-$\psi$ is the half-opening angle of the conical jet, and $i$ is the
-angle between the jet and the line of sight.
-
-A rough argument about observability: For our example, the period is
-0.1 yr at $t = 63967.3816055$ yr.  We can assume that the chirp begins
-at $t = 10^4$ yr when period is $\sim 2$ yr.  So the total duration of
-the chirp is $5.3\times 10^4$ yr.
-
-A 2 yr period gives $\sim 1$ milli-arcsec wiggles in the jet.  So the
-idea is that the wiggles will change from milli-arcsec to micro-arcsec
-over $\sim 10^5$ yr.  That is over a length scale of $\sim 30$ kpc.
-
-Now the question is: how big an angle would a 30 kpc jet subtend on
-the sky in arcsecs?
-
-Let us assume $z = 0.302$ which is the redshift of 1928+738 jet that
-Roos (1993) have studied.  At this redshift 30 kpc will subtend about
-6.748195445702 arcsec.
-
-So we are talking about mciro- to milli-arcssec scale wiggles in tens
-of arcsec scale jets.  Jets are routinely seen over hundreds of
-arcsec.
-
-\section{Conclusion}
-
-From Figures \ref{fig:a_gw} we can make these four points:
-
-\begin{itemize}
-\item From the jet structure, we can tell how long until the binary enters GW-induced inspiral.
-\item In particular, we can flag binaries that are entering GW-inspiral now.
-\item We could locate binaries that are now in GW inspiral phase but
-  have helical jet structures that are impossible to resolve.
-\item From the above jet structure, we get an idea of the history of gas inflow on the binary.
-\end{itemize}
-
-There will also probably be some spectral evolution along the jet.
-This spectral evolution will act as a cross check on the evolution
-deduced from the jet morphology.  An important advantage of this
-technique is that it allows us to locate binaries that may be
-impossible to locate even with VLBI.
+and
+\begin{equation}
+  y=v_yt/(1-v_x/c)
+\end{equation}
+respectively.  In angular coordinates, the motion on the sky is given
+by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular diameter
+distance of the binary.  Because of superluminal motion, the forward
+jet is stretched while the backward jet is compressed.
 
@@
     \begin{tikzpicture}
-      \draw [->,densely dashed,thick] (0,0) -- (3.0,0);
-      \draw [->,densely dashed,thick] (0,0) -- (0,4.0);
-      \draw [->,thick] (0,0) -- (2.0,4.0);
-      \draw [->,thick,red] (0,0) -- (5.0,4.0);
-      \draw (1.0,2.0) arc (63:90:2.2);
-      \draw (2.0,1.6) arc (38.6:63:2.56);
-      \draw (0.5,2.5) node {$\chi$};
-      \draw (1.90,2.3) node {$\psi$};
-      \draw [densely dashed,thick] (0,0) -- (3.0,6.0);
-      \draw (0.0,4.5) node {$\mathbf{L}_\mathrm{orbital}$};
-      \draw (4.0,0.0) node {$\mathbf{v}_\mathrm{orbital}$};
-      \draw (1.0,3.0) node {$\mathbf{v}_\mathrm{jet}$};
+      \draw [thick,color=brown] (0,0) -- (-4.0,0);
+      \draw [thick,color=brown] (0,0) -- (0,4.0);
+      \draw [thick,color=brown] (0,0) -- (2.6,-3.0);
+
+      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0.05,0.0) -- (0.05,4.0);
+      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0,0.0) -- (3.46,-2.0);
+      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0,0.0) -- (-3.87,-1.0);
+
+      \draw [thick,rotate around={-75:(-2.95,-0.76)}] (-2.95,-0.76) ellipse (0.55 and 0.2);
+      \draw [thick,rotate around={12:(3.04,0.24)}] (3.04,0.238) arc (-90:90:0.2 and 0.55);
+      \draw [thick,color=gray,rotate around={12:(3.04,0.24)},dashed](3.04,0.238) arc (270:90:0.2 and 0.55);
+
+      \draw [thick] (-2.82,-1.3) -- (2.82,1.3);
+      \draw [thick] (-3.09,-0.24) -- (3.09,0.24);
+
+      \draw (0.1,4.3) node {\large \textcolor{brown}{$y$},\textcolor{red}{$\,y^\prime$}};
+      \draw (-4.2,0) node {\textcolor{brown}{\large $z$}};
+      \draw (2.75,-3.17) node {\textcolor{brown}{\large $x$}};
+      \draw (3.1,-3.57) node {\large [line of sight]};
+      \draw (-4.1,-1.059) node {\large \textcolor{red}{$z^\prime$}};
+      \draw (3.7,-2.1) node {\textcolor{red}{\large $x^\prime$}};
+
+      \draw [thick,color=brown] (-1.0,-0.258) arc (230:283:1.8);
+      \draw (-0.5,-1.0) node {\textcolor{brown}{\large $i$}};
+
+      \draw (-2.0,-0.92) arc (204.75:194.48:2.2015);
+      \draw (-2.0,-1.3) node {\large $\psi$};
+
     \end{tikzpicture}
   \end{center}
-  \caption{The orbital velocity $\mathbf{v}_\mathrm{orbital}$
-    introduces the jet precision angle $\psi$.  Without the orbital
-    motion, $\psi=0$ and the jet traces a cylindrical surface.  With
-    the orbital motion, as the BH moves along its circular orbit, the
-    red vector oscillates around the solid black vector and produces a
-    conical surface with $\psi$ as its half-opening angle.  Note that
-    $v_\mathrm{orbital}$ is exagerated here.  In practice,
-    $v_\mathrm{orbital}\ll v_\mathrm{jet}\sim c$.  This figure then
-    shows that
-    $\sin{\psi}\sim\tan{\psi}=v_\mathrm{orbital}\cos{\chi}/v_\mathrm{jet}$.}
-  \label{fig:jet}
+  \caption{Geometry of the jet.  The observer's sky is in the $yz$
+    plane, so that the line of sight is along the $x$-axis.  The jet
+    precesses around the $z^\prime$ axis with half-opening angle
+    $\psi$. The $y$ and $y^\prime$ axes overlap, and $z$, $x$, and
+    $z^\prime$ axes are co-planer.  The angle between the $z^\prime$
+    axis and the line of sight is $i$.  This geometry is valid in the
+    limit $a\ll z^\prime\psi$.}
+  \label{fig:jetgeom}
 \end{figure}
+
+Finally, we assume that the emission from each jet particle is
+optically thin at the frequency of observation with a spectrum in the
+rest frame of the particle given by $P(\nu)\propto \nu^{-\alpha}$
+\citep{1982ApJ...262..478G}.  The observed flux density $S(\nu)$ in erg
+cm$^{-2}$ Hz$^{-1}$ is given by
+\begin{equation}
+  S(\nu)=S_r(\nu)D^{3+\alpha},
+\end{equation}
+where $D$ is the Doppler shift parameter
+$[\gamma(1-\beta\cos\phi)]^{-1}$ and $S_r$ is the rest frame flux
+density.  Further to consider the effects of jet evolution we assume
+that $S_r$ decreases with time according to a simple power law
+\begin{equation}
+  S_r\propto t^{-\delta}.
+\end{equation}
+We assume $\delta=1$ and $\alpha=1$, hereafter. 
+% Need to justify parameter values.
+% Comment on relativistic corrections
 
@@
 \begin{center}
-  \begin{tabular}{ccc}
-    \includegraphics*[width=2.2in]{a_gw.pdf} &
-    \includegraphics*[width=2.2in]{v_orb.pdf} &
-    \includegraphics*[width=2.2in]{psi.pdf}
+  \begin{tabular}{cc}
+    \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
+    \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
   \end{tabular}
 \end{center}
-\caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
-\label{fig:a_gw}
+\caption{The left panel shows forward and backward jets for an
+  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
+  assumed $i=40^\circ$, $\beta=0.9$, and $d = 100$~Mpc.  The jets are
+  twisted because of the apparant superluminal motion.  Also evident
+  is the opening of the jet and the radio jet chirp.  Right panel
+  shows normalised brightness of the inner 30~mas region of the
+  forward jet.}
+\label{fig:jet_mdot1.0}
 \end{figure*}
 
-\begin{figure*}
+\section{Flagging GW binaries and probing GR}
+
+As the binary evolves, its increasing orbital period and speed, and
+the apparant superluminal motion effect, together result in a jet
+morphology with an increasing opening angle and increasingly rapid
+periodic variation together with a twisted shape.
+Figure~\ref{fig:jet_mdot1.0} shows these three features for an
+equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  For
+concreteness, we have assumed $i=40^\circ$ and $\beta=0.9$.  The
+angular diameter distance to the binary is assumed to be $d =
+100$~Mpc, which corresponds to redshift $z\sim 0.3$.  The forward jet
+streched on the sky while the backward jet is compressed due to the
+superluminal motion.  The right panel of Figure~\ref{fig:jet_mdot1.0}
+zooms into the inner region of the forward jet and shows that the
+binary creates wiggles in the jet morphology of the order of a
+milli-arcsecond.  The duration of the binary evolution captured by
+Figure~\ref{fig:jet_mdot1.0} is 100 yr.
+
+% Add discussion about the chirp of the signal
+
+\begin{table}
 \begin{center}
-  \begin{tabular}{ccc}
-    \includegraphics*[width=2.2in]{a_gas.pdf} &
-    \includegraphics*[width=2.2in]{v_orb_gas.pdf} &
-    \includegraphics*[width=2.2in]{psi_gas.pdf}
+\begin{tabular}{c l}
+\hline 
+Parameter & Description \\ 
+\hline 
+$M$ & Total mass \\
+$\zeta$ & Mass asymmetry \\
+$\chi$ & Intrinsic jet direction \\
+$e$ & Orbital eccentricity \\
+$\alpha$ & SED slope \\
+$\delta$ & Time evolution of brightness evolution \\
+$\beta$ & Jet speed \\
+$i$ & Orbital inclination \\ 
+$\dot{\mathcal M}$ & Gas accretion rate\\
+\hline
 \end{tabular}
 \end{center}
-\caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
-\label{fig:a_gas}
-\end{figure*}
+\caption{Parameters of the problem.}
+\label{table:params}
+\end{table}
 
-%% \begin{figure}
-%%   \begin{center}
-%%     \includegraphics*[width=\columnwidth]{a_gw.pdf}
-%%   \end{center}
-%%   \caption{Evolution of the binary separation.  Time $t$ is time since
-%%     the gravitational wave induced phase began.}
-%%   \label{fig:a_gw}
-%% \end{figure}
+\section{Conclusion}
 
-%% \begin{figure}
-%%   \begin{center}
-%%     \includegraphics*[width=\columnwidth]{v_orb.pdf}
-%%   \end{center}
-%%   \caption{Evolution of the orbital velocity}
-%%   \label{fig:v_orb}
-%% \end{figure}
+Using a model of jet precession in a binary BH, we have shown that the
+jet morphology of a binary that is entering its GW-induced phase of
+inspiral shows a distinct chirping morphology. Jets in these binaries
+have a biconical morphology.  After the binary enters the GW phase,
+the opening angle of cone rapidly increases within a distance of tens
+of milli-arcseconds.  In addition, the periodicity of the jet's motion
+on the conical surface increases as the orbital period of the binary
+goes up in this stage of its evolution.  Finally, due to apparent
+superluminal motion and the widening opening angle, the jet is twisted
+into a peculiar morphology.  The jet geometry can be used to infer the
+set of dynamical and geometrical parameters, listed in
+Table~\ref{table:params}.
 
-%% \begin{figure}
-%%   \begin{center}
-%%     \includegraphics*[width=\columnwidth]{psi.pdf}
-%%   \end{center}
-%%   \caption{Evolution of the intrinsic half-opening angle of the
-%%     conical jet.  it is important to note that the jet lifetime is
-%%     about $10^7$ yr.}
-%%   \label{fig:psi}
-%% \end{figure}
+% Give redshift numbers
+These milli-arcsecond features can be resolved using long baseline
+radio interferometry \citep{2001ChJAA...1..236Q}.  At higher
+redshifts, jets are fainter and the wiggles become more difficult to
+resolve.  However, it might be possible to resolve binaries with very
+unequal mass ratios, eccentric orbits, or suitable inclinations.  The
+future Square Kilometer Array (SKA) should make microarcsecond
+resolution possible, which could increase the detection rate of these
+objects \citep{2014arXiv1412.5971P}.
 
-\pagebreak
-\pagebreak
+The jet morphology acts as an electromagnetic counterpart for chirping
+gravitational wave sources.  BH binaries are expected to be the
+brightest GW sources in the universe.  Current ground based GW
+detectors such as LIGO are sensitive to low-mass BH binaries at high
+redshift, which would be difficult to detect with this method.
+However, future detectors such as eLISA are sensitvie to GW emission
+from $10^7$--$10^9$ M$_\odot$ BHs \citep{2013CQGra..30x4009S}.  Pulsar
+Timing Arrays are sensitive to even higher binary masses.  Monitoring
+of jet morphologies with features outlined in this paper will put a
+lower bound on the expected abundance of bright GW sources for this
+experiments. Moreover, combined GW and electromagnetic observations of
+compact binary mergers should enable detailed studies of astrophysical
+processes in the strong-field gravity regime
+\citep{2005ApJ...629...15H}.
 
-\clearpage
+% Mention old observations which show similar features
+% There is a todo in a PS email from Avi
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{jet_i10_psi20.pdf}
-  \end{center}
-  \caption{$i=10^\circ$, $\psi=20^\circ$, $\beta=0.90$}
-\end{figure}
+\section*{Acknowledgments}
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{jet_i40_psi20.pdf}
-  \end{center}
-  \caption{$i=40^\circ$, $\psi=20^\circ$, $\beta=0.90$}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{jet_i80_psi20.pdf}
-  \end{center}
-  \caption{$i=80^\circ$, $\psi=20^\circ$, $\beta=0.90$}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{jet_i40_psi40.pdf}
-  \end{center}
-  \caption{$i=40^\circ$, $\psi=40^\circ$, $\beta=0.90$}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}
-\begin{center}
-  \begin{tabular}{ccc}
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40.pdf}
-  \end{tabular}
-\end{center}
-\caption{Jet structure in the source frame. $\beta=0.9, 0.8, 0.4$ from left to right.}
-\label{fig:srcframe}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \begin{tabular}{ccc}
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_obs.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_obs.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_obs.pdf}
-  \end{tabular}
-\end{center}
-\caption{Jet structure in the observer frame. $\beta=0.9, 0.8, 0.4$ from left to right.}
-\label{fig:obsframe}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \begin{tabular}{ccc}
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_ang.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_ang.pdf} &
-    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_ang.pdf}
-  \end{tabular}
-\end{center}
-\caption{Jet structure in angular coordinates. $\beta=0.9, 0.8, 0.4$ from left to right.}
-\label{fig:ang}
-\end{figure*}
+We acknowledge useful discussions with Enrico Ramirez-Ruiz and Alberto
+Rorai.  This work was supported in part by NSF grant AST-1312034 (for
+A.~L.).
 

### 2015-07-06 Paper/scratch.tex
Add new plots to paper
Made more changes suggested by Avi.

@@
   = 2.42\, \mathrm{degrees}.
+
+%% \begin{figure*}
+%% \begin{center}
+%%   \begin{tabular}{ccc}
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40.pdf}
+%%   \end{tabular}
+%% \end{center}
+%% \caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
+%% \label{fig:srcframe}
+%% \end{figure*}
+
+%% \begin{figure*}
+%% \begin{center}
+%%   \begin{tabular}{ccc}
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_obs.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_obs.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_obs.pdf}
+%%   \end{tabular}
+%% \end{center}
+%% \caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
+%% \label{fig:obsframe}
+%% \end{figure*}
+
+%% \begin{figure*}
+%% \begin{center}
+%%   \begin{tabular}{ccc}
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_ang.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_ang.pdf} &
+%%     \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_ang.pdf}
+%%   \end{tabular}
+%% \end{center}
+%% \caption{Evolution of the binary separation.  Time $t$ is time since the gravitational wave induced phase began.}
+%% \label{fig:ang}
+%% \end{figure*}
+
+are expected to arise naturally in the context of the
+heirarchical formation of galaxies in which galaxies with their black
+holes merge and create black hole binaries (cite our paper, illustris
+papers, many other papers).  As such binaries are great tools to
+understand galaxy formation.  They may also be crucial to explain the
+occurancec of massive BHs today, which are routinely invoked for AGN
+feedback etc.  Binary BHs also probe the extreme limit of GR.  They
+are thus a tool to test GR and probe physics beyond it.
+
+Binary evolution proceeds through three phases.  Describe three
+stages.  Out of this binaries spent most of their time on pc scale
+separations.  These wide binaries are expected to leave signature in
+broad lines regions (cite Shen and Loeb).  There have been many
+attempts to detect these wide binaries but none have been detected so
+far (cite Decarli).
+
+Aprt from a stochastic gravitational wave background, LISA is expected
+to detect GW from individial sources.  These GWs are produced by a
+binary BHs. Briefly describe GW chirp here?  If one of these BHs is
+accreting, it is expected to have a jet.  Due to orbital motion of the
+accreting BH, the jet will precess.  We study this jet precession in
+this paper during the last stages of the binary evolution.  Any
+peculiar morphological features canact as an electromagnetic signature
+of the binary.
+
+---------------
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology with opening angle evolution.}
+\label{fig:jet_psi}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_full.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom1.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_beta0_90_zoom2.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet morphology with opening angle evolution.}
+\label{fig:jet_psi}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.0_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.0_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=1$}
+\label{fig:jet_mdot1.0}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.00_image_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot1.00_image_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=1$}
+\label{fig:jet_mdot1.0_image}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot10.00_full}.pdf} &
+    \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot10.00_zoom1}.pdf} 
+  \end{tabular}
+\end{center}
+\caption{$a^4/M^3=10$}
+\label{fig:jet_10.0}
+\end{figure*}
+
+\begin{figure*}
+  \begin{center}
+    \begin{tabular}{cc}
+      \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot0.10_full}.pdf} &
+      \includegraphics*[width=\columnwidth]{{jet_i40_beta0.90_mdot0.10_zoom1}.pdf} 
+    \end{tabular}
+  \end{center}
+  \caption{$a^4/M^3=0.1$}
+  \label{fig:jet_mdot0.1}
+\end{figure*}
+
+and
+the binary separation shrinks
+
+For observable predictions we have to project this in observer plane.
+Let the angle between the jet direction and the observer line of sight
+be $i$.  Then
+\begin{equation}
+  \psi^\mathrm{obs}\approx\frac{\psi}{\sin i}.
+\end{equation}
+
+This gives
+\begin{equation}
+  \psi^\mathrm{obs}=2.42^\circ \left(\frac{v_\mathrm{jet}}{0.95 c}\right)^{-1}\cos\left(\frac{\chi}{30^{\circ}}\right)\left[\sin\left(\frac{i}{15^{\circ}}\right)\right]^{-1},
+\end{equation}
+This value in itself is probably not very interesting as it depends on
+$v_\mathrm{jet}$ and $\chi$, but it gets interesting when we consider
+the evolution of $\psi^\mathrm{obs}$ below.
+
+Binary separation, jet velocity, and the two angles, tell us about the
+instantaneous configuration of the jet.  (Recall that we are only
+considering an equal mass binary.)  The jet evolution is governed by
+the dynamical evolution of the binary separation.  During the
+gas-dominated phase, this is given by
+\begin{equation}
+\frac{a}{\dot a} = \mu/\dot M = 1.1\times 10^7\, \mathrm{yr}\,\dot{\mathcal{M}}^{-1}.   
+\end{equation}
+This increases by two orders of magnitude during the GW phase to
+\begin{equation}
+\frac{a}{\dot a} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
+\end{equation}
+
+Figure shows the effect of this model.
+
+Note that this section introduces a set of geometrical and
+astrophysical parameters for the system we started with in the
+previous section.  List parameters and mention our assumed values.
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{a_gw.pdf} &
+    \includegraphics*[width=2.2in]{v_orb.pdf} &
+    \includegraphics*[width=2.2in]{psi.pdf}
+  \end{tabular}
+\end{center}
+\caption{Evolution of the binary separation in the gravitational wave stage.}
+\label{fig:a_gw}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{a_gas.pdf} &
+    \includegraphics*[width=2.2in]{v_orb_gas.pdf} &
+    \includegraphics*[width=2.2in]{psi_gas.pdf}
+  \end{tabular}
+\end{center}
+\caption{Evolution of the binary separation in the gas stage.}
+\label{fig:a_gas}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{ccc}
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_90_ang.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_80_ang.pdf} &
+    \includegraphics*[width=2.2in]{jet_i40_psi20_beta_0_40_ang.pdf}
+  \end{tabular}
+\end{center}
+\caption{Jet structure in angular coordinates. $\beta=0.9, 0.8, 0.4$ from left to right.}
+\label{fig:ang}
+\end{figure*}
+
+Jet shown in figure has three features: (1) chirp, (2) beam twisting,
+and (3) cone opening.  This can be observed using VLBA.  Figure shows
+the evolution of the binary when it is in the gas accretion phase.  We
+have chosen steady gas accretion with $\dot{\mathcal{M}}=1$ here for
+simplicity.  Note that the starting separation here is about 100 pc.
+What we find is that in the beginning the binary orbital period is
+about $10^7$ yr. So the jet will show no signs of precision at this
+point.  However as the binary separation reduces to about 10 pc, the
+period reduces to $\sim 100$ yr and the jet will start showing
+precision.  At this point the time scale for binary merger is $\sim
+10^7$ yr.
+
+In other words \emph{since the gas accretion time scale is of the
+  order of the typical jet lifetime} the jet structure will show a
+transition from a straight jet to a precessing jet at a separation of
+\begin{equation}
+  v_\mathrm{apparant}\left(\frac{t_0}{10^7}\right).
+\end{equation}
+from the binary.  \textbf{[Convert this to arcseconds on the sky.]}
+The location at which the jet morphology shows this precision and the
+opening angle evolution give a direct constraint on the epoch of
+graviational wave emission of the binary.
+
+Thus the new insight here is that the signature of the jet precision
+due to GW-inspiral is observationally accessible.
+
+Variability of lensed quasars constraints the size of accretion disks
+to about $10^{15}$ cm ($\sim 10^{-3}$ pc) \citep{2010ApJ...712.1129M}.
+This is smaller than the $a_\mathrm{GW}$ by a factor of less than 10.
+
+This means that the GW-induced precision will be visible in the jet
+precision for a very short distance near the source before the
+accretion disk is disrupted.
+
+Last plots in this note show some example jets.  In all these plots
+$\psi$ is the half-opening angle of the conical jet, and $i$ is the
+angle between the jet and the line of sight.
+
+A rough argument about observability: For our example, the period is
+0.1 yr at $t = 63967.3816055$ yr.  We can assume that the chirp begins
+at $t = 10^4$ yr when period is $\sim 2$ yr.  So the total duration of
+the chirp is $5.3\times 10^4$ yr.
+
+A 2 yr period gives $\sim 1$ milli-arcsec wiggles in the jet.  So the
+idea is that the wiggles will change from milli-arcsec to micro-arcsec
+over $\sim 10^5$ yr.  That is over a length scale of $\sim 30$ kpc.
+
+Now the question is: how big an angle would a 30 kpc jet subtend on
+the sky in arcsecs?
+
+Let us assume $z = 0.302$ which is the redshift of 1928+738 jet that
+Roos (1993) have studied.  At this redshift 30 kpc will subtend about
+6.748195445702 arcsec.
+
+So we are talking about mciro- to milli-arcssec scale wiggles in tens
+of arcsec scale jets.  Jets are routinely seen over hundreds of
+arcsec.
+
+Connect with LISA
+
+From Figures \ref{fig:a_gw} we can make these four points:
+
+\begin{itemize}
+\item From the jet structure, we can tell how long until the binary enters GW-induced inspiral.
+\item In particular, we can flag binaries that are entering GW-inspiral now.
+\item We could locate binaries that are now in GW inspiral phase but
+  have helical jet structures that are impossible to resolve.
+\item From the above jet structure, we get an idea of the history of gas inflow on the binary.
+\end{itemize}
+
+There will also probably be some spectral evolution along the jet.
+This spectral evolution will act as a cross check on the evolution
+deduced from the jet morphology.  An important advantage of this
+technique is that it allows us to locate binaries that may be
+impossible to locate even with VLBI.

### 2015-07-07 Paper/paper.tex
Update abstract

@@
 \begin{abstract}
-Apart from a stochastic gravitational wave (GW) background, eLISA will
-also be able to detect GW emission from binary black holes.  Combined
-GW and electromagnetic observations of compact binary mergers should
-enable detailed studies of astrophysical processes in the strong-field
-gravity regime.  Here we show that if one of component BHs of the
-binary has an accretion disk with an associated jet, the morphology of
-the jet can act as an electromagnetic counterpart to the GW signal
-from the binary.  For binaries with mass $10^7$--$10^{10}$ M$_\odot$
-at redshifts $z<0.5$, the jet precesses along a biconical surface due
-to the binary's orbital motion.  When the binary enters the GW phase
-of its evolution, the opening angle widens, the jet exhibits
-milliarcsecond scale wiggles and the conical surface of jet precession
-is twisted due to the superluminal motion.  These morphological
-features can be used to constrain the dyanamical parameters of the
-binary.  Monitoring of jets with these features in current and
-archival data will put a lower limit on the sources that could be
-detected by GW experiments such as eLISA and Pulsar Timing Arrays
-(PTAs).  In the future, microarcsecond interferometry with experiments
-such as SKA will increase the potential usefulness of this technique.  
+We study a novel electromagnetic signature of supermassive black hole
+binaries whose inspiral stars being dominated by gravitational wave
+(GW) emission. Combined GW and electromagnetic observations of compact
+binary mergers would allow to use them as standard sirens for
+cosmology and a testing ground of general relativity.  If one of the
+binary members produces a radio jet, the helical chirping morphology
+of the jet can be used to infer the binary parameters.  For binaries
+with mass $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, the jet
+precesses along a biconical surface due to the binary's orbital
+motion.  When the binary enters the GW phase of its evolution, the
+opening angle widens, the jet exhibits milliarcsecond scale wiggles
+and the conical surface of jet precession is twisted due to the
+superluminal motion.  Monitoring these features in current and
+archival data will place a lower limit on sources that could be
+detected by eLISA and Pulsar Timing Arrays.  In the future,
+microarcsecond interferometry with the Square Kilometer Array will
+increase the potential usefulness of this technique.
 \end{abstract}

### 2015-07-08 Paper/paper.tex
Change from ApJ to MNRAS format

@@
-\documentclass[iop]{emulateapj}
-\bibliographystyle{apj}
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
+
+% MNRAS is set in Times font. If you don't have this installed (most LaTeX
+% installations will be fine) or prefer the old Computer Modern fonts, comment
+% out the following line
+%% \usepackage{newtxtext,newtxmath}
+
+% Depending on your LaTeX fonts installation, you might get better results with one of these:
+%% \usepackage{mathptmx}
+%% \usepackage{txfonts}
+
+% Use vector fonts, so it zooms properly in on-screen viewing software
+% Don't change these lines unless you know what you are doing
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
+
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+
+\usepackage{graphicx}	
 \usepackage{amsmath}	
-\usepackage{aas_macros,amssymb}
-\usepackage{natbib,graphicx,layouts}
-\usepackage{color,hyperref}
-\usepackage{bm}
-\definecolor{linkcolor}{rgb}{0,0,0.5}
-\definecolor{notecolor}{rgb}{0.8,0,0}
-\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
-  filecolor=linkcolor,urlcolor=linkcolor}
+\usepackage{amssymb}
 \usepackage{tikz}
@@
 
-\shorttitle{Chirping Jets from Black Hole Binaries}
-\shortauthors{Kulkarni et al.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+
+\title[Chirping Jets from Black Hole Binaries]{Radio Crickets:
+  Chirping Jets from Black Hole Binaries Entering their Gravitational
+  Wave Inspiral}
+
+\author[Kulkarni and Loeb]{
+Girish Kulkarni,$^{1}$\thanks{E-mail: kulkarni@ast.cam.ac.uk}
+and Abraham Loeb$^{2}$\thanks{E-mail: aloeb@cfa.harvard.edu}
+\\
+$^{1}$Institute of Astronomy and Kavli
+  Institute of Cosmology, University of Cambridge, Madingley Road,
+  Cambridge CB3 0HA, UK\\ 
+$^{2}$Institute for Theory \& Computation,
+  Harvard University, 60 Garden Street, Cambridge, MA 02138, USA\\ 
+}
+
+% These dates will be filled out by the publisher
+\date{Accepted ---. Received ---; in original form ---}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2015}
+
+% Don't change these lines
 \begin{document}
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
-\title{Radio Crickets: Chirping Jets from Black Hole Binaries Entering
-  their Gravitational Wave Inspiral} \author{Girish
-  Kulkarni\altaffilmark{1} and Abraham Loeb\altaffilmark{2}}
-\altaffiltext{1}{Institute of Astronomy and Kavli Institute of
-  Cosmology, University of Cambridge, Madingley Road, Cambridge CB3
-  0HA, UK; \email{kulkarni@ast.cam.ac.uk}} \altaffiltext{2}{Institute
-  for Theory \& Computation, Harvard University, 60 Garden Street,
-  Cambridge, MA 02138, USA; \email{aloeb@cfa.harvard.edu}}
-
+% Abstract of the paper
 \begin{abstract}
-We study a novel electromagnetic signature of supermassive black hole
-binaries whose inspiral stars being dominated by gravitational wave
-(GW) emission. Combined GW and electromagnetic observations of compact
-binary mergers would allow to use them as standard sirens for
-cosmology and a testing ground of general relativity.  If one of the
-binary members produces a radio jet, the helical chirping morphology
-of the jet can be used to infer the binary parameters.  For binaries
-with mass $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, the jet
-precesses along a biconical surface due to the binary's orbital
-motion.  When the binary enters the GW phase of its evolution, the
-opening angle widens, the jet exhibits milliarcsecond scale wiggles
-and the conical surface of jet precession is twisted due to the
-superluminal motion.  Monitoring these features in current and
-archival data will place a lower limit on sources that could be
-detected by eLISA and Pulsar Timing Arrays.  In the future,
+  We study a novel electromagnetic signature of supermassive black
+  hole binaries whose inspiral stars being dominated by gravitational
+  wave (GW) emission. Combined GW and electromagnetic observations of
+  compact binary mergers would allow to use them as standard sirens
+  for cosmology and a testing ground of general relativity.  If one of
+  the binary members produces a radio jet, the helical chirping
+  morphology of the jet can be used to infer the binary parameters.
+  For binaries with mass $10^7$--$10^{10}$ M$_\odot$ at redshifts
+  $z<0.5$, the jet precesses along a biconical surface due to the
+  binary's orbital motion.  When the binary enters the GW phase of its
+  evolution, the opening angle widens, the jet exhibits milliarcsecond
+  scale wiggles and the conical surface of jet precession is twisted
+  due to the superluminal motion.  Monitoring these features in
+  current and archival data will place a lower limit on sources that
+  could be detected by eLISA and Pulsar Timing Arrays.  In the future,
   microarcsecond interferometry with the Square Kilometer Array will
@@
 
-\keywords{BHs}
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+keyword1 -- keyword2 -- keyword3
+\end{keywords}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
 
@@
 
-We now ask if it would be possible to observe the opening of the jet
-morphology.  In order to do this, it is necessary to map the above
-calculations to the observer's frame \citep{1982ApJ...262..478G}.  At
-a large distance compared to the binary's separation, the jet geometry
-is illustrated in Figure \ref{fig:jetgeom}.  Here the source frame of
-reference is shown by $x^\prime$, $y^\prime$, $z^\prime$ coordinate
-systems.  The $x$, $y$, $z$ coordinate system represents the
-observer's frame such that the $yz$ is the plane of the sky.  Further,
-the jet is oriented such that that the $y^\prime$ axis coincides with
-the $y$ axis, the original direction of jet emission is at an angle
-$i$ relative to the line of sight ($x$-axis).  For simplicity, we also
-assume that the $z^\prime$ axis is in the $xz$ plane.  Then the jet
-will precess on a cone with a half-opening angle $\psi$ and an angular
-velocity of precession $\Omega$.  In the source frame of reference,
-the equation of the jet is then given by \citep{1981ApJ...246L.141H}
-\begin{eqnarray}
-  &v_x=&s_\mathrm{jet}\beta c\left[\sin\psi\sin i\cos\Omega t+\cos\psi\cos i\right],\\
-  &v_y=&s_\mathrm{jet}\beta c\sin\psi\sin\Omega t,~\,\mathrm{and}\\
-  &v_z=&s_\mathrm{jet}\beta c\left[\cos\psi\sin i-\sin\psi\cos i\cos\Omega t\right]. 
-\end{eqnarray}
-Here $s_\mathrm{jet}=\pm 1$ for forward and backward jets,
-respectively.  The coordinates of a jet particle in the sky is then
-given by $z=v_zt$ and $y=v_yt$.  However, because of superluminal
-motion, the observer will see this on the sky as
-\begin{equation}
-  z=v_zt/(1-v_x/c)
-\end{equation}
-and
-\begin{equation}
-  y=v_yt/(1-v_x/c)
-\end{equation}
-respectively.  In angular coordinates, the motion on the sky is given
-by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular diameter
-distance of the binary.  Because of superluminal motion, the forward
-jet is stretched while the backward jet is compressed.
-
 \begin{figure}
@@
 
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
+    \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
+  \end{tabular}
+\end{center}
+\caption{The left panel shows forward and backward jets for an
+  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
+  assumed $i=40^\circ$, $\beta=0.9$, and $d = 100$~Mpc.  The jets are
+  twisted because of the apparant superluminal motion.  Also evident
+  is the opening of the jet and the radio jet chirp.  Right panel
+  shows normalised brightness of the inner 30~mas region of the
+  forward jet.}
+\label{fig:jet_mdot1.0}
+\end{figure*}
+
+We now ask if it would be possible to observe the opening of the jet
+morphology.  In order to do this, it is necessary to map the above
+calculations to the observer's frame \citep{1982ApJ...262..478G}.  At
+a large distance compared to the binary's separation, the jet geometry
+is illustrated in Figure \ref{fig:jetgeom}.  Here the source frame of
+reference is shown by $x^\prime$, $y^\prime$, $z^\prime$ coordinate
+systems.  The $x$, $y$, $z$ coordinate system represents the
+observer's frame such that the $yz$ is the plane of the sky.  Further,
+the jet is oriented such that that the $y^\prime$ axis coincides with
+the $y$ axis, the original direction of jet emission is at an angle
+$i$ relative to the line of sight ($x$-axis).  For simplicity, we also
+assume that the $z^\prime$ axis is in the $xz$ plane.  Then the jet
+will precess on a cone with a half-opening angle $\psi$ and an angular
+velocity of precession $\Omega$.  In the source frame of reference,
+the equation of the jet is then given by \citep{1981ApJ...246L.141H}
+\begin{eqnarray}
+  &v_x=&s_\mathrm{jet}\beta c\left[\sin\psi\sin i\cos\Omega t+\cos\psi\cos i\right],\\
+  &v_y=&s_\mathrm{jet}\beta c\sin\psi\sin\Omega t,~\,\mathrm{and}\\
+  &v_z=&s_\mathrm{jet}\beta c\left[\cos\psi\sin i-\sin\psi\cos i\cos\Omega t\right]. 
+\end{eqnarray}
+Here $s_\mathrm{jet}=\pm 1$ for forward and backward jets,
+respectively.  The coordinates of a jet particle in the sky is then
+given by $z=v_zt$ and $y=v_yt$.  However, because of superluminal
+motion, the observer will see this on the sky as
+\begin{equation}
+  z=v_zt/(1-v_x/c)
+\end{equation}
+and
+\begin{equation}
+  y=v_yt/(1-v_x/c)
+\end{equation}
+respectively.  In angular coordinates, the motion on the sky is given
+by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular diameter
+distance of the binary.  Because of superluminal motion, the forward
+jet is stretched while the backward jet is compressed.
+
 Finally, we assume that the emission from each jet particle is
@@
 % Comment on relativistic corrections
-
-\begin{figure*}
-\begin{center}
-  \begin{tabular}{cc}
-    \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
-    \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
-  \end{tabular}
-\end{center}
-\caption{The left panel shows forward and backward jets for an
-  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
-  assumed $i=40^\circ$, $\beta=0.9$, and $d = 100$~Mpc.  The jets are
-  twisted because of the apparant superluminal motion.  Also evident
-  is the opening of the jet and the radio jet chirp.  Right panel
-  shows normalised brightness of the inner 30~mas region of the
-  forward jet.}
-\label{fig:jet_mdot1.0}
-\end{figure*}
 
@@
 
-\bibliography{refs}
+\bibliographystyle{mnras}
+\bibliography{refs} % if your bibtex file is called example.bib
+
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
 \end{document}

### 2015-07-08 Paper/scratch.tex
Change from ApJ to MNRAS format

@@
 impossible to locate even with VLBI.
+
+Normally the next section describes the techniques the authors used.
+It is frequently split into subsections, such as Section~\ref{sec:maths} below.
+
+\subsection{Maths}
+\label{sec:maths} % used for referring to this section from elsewhere
+
+Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
+or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
+as a numbered equation:
+
+\begin{equation}
+    x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
+	\label{eq:quadratic}
+\end{equation}
+
+Refer back to them as e.g. equation~(\ref{eq:quadratic}).
+
+\subsection{Figures and tables}
+
+Figures and tables should be placed at logical positions in the text. Don't
+worry about the exact layout, which will be handled by the publishers.
+
+Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
+e.g. Table~\ref{tab:example_table}.
+
+% Example figure
+\begin{figure}
+	% To include a figure from a file named example.*
+	% Allowable file formats are eps or ps if compiling using latex
+	% or pdf, png, jpg if compiling using pdflatex
+	\includegraphics[width=\columnwidth]{example}
+    \caption{This is an example figure. Captions appear below each figure.
+	Give enough detail for the reader to understand what they're looking at,
+	but leave detailed discussion to the main body of the text.}
+    \label{fig:example_figure}
+\end{figure}
+
+% Example table
+\begin{table}
+	\centering
+	\caption{This is an example table. Captions appear above each table.
+	Remember to define the quantities, symbols and units used.}
+	\label{tab:example_table}
+	\begin{tabular}{lccr} % four columns, alignment for each
+		\hline
+		A & B & C & D\\
+		\hline
+		1 & 2 & 3 & 4\\
+		2 & 4 & 6 & 8\\
+		3 & 5 & 7 & 9\\
+		\hline
+	\end{tabular}
+\end{table}

### 2015-07-08 jet3d_check2_ellipticity.py
Add orbital ellipticity

@@ -1,0 +1,147 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+m1 = 0.5e10 #msun
+m2 = m1 #msun
+mu = m1*m2/(m1+m2) # reduced mass
+zeta = 4.0*mu/(m1+m2) # See Loeb 2010 near eqn.1
+print 'zeta=', zeta
+
+e = 0.6 # orbital eccentricity 
+def F(e):
+    # See Merritt and Milos 2005 eqn 7 
+    return (1.0-e*e)**3.5 * (1.0 + 73.0*e*e/24.0 + 37.0*e*e*e*e/96.0)
+
+print 'e=', e
+print 'F(e)=', F(e) 
+
+M = 1.0e10 # Msun; total mass of the equal-mass binary
+Mdot = 1.0 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3
+coeff /= zeta # for unequal-mass binary; see Loeb 2010 eqn. 4
+coeff /= F(e) # eccentricity 
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
+save_pdf = False
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
[…]

### 2015-07-08 jet3d_check2_unequalmass.py
Consider unequal BH masses

@@ -1,0 +1,138 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+m1 = 0.5e10 #msun
+m2 = m1 #msun
+mu = m1*m2/(m1+m2) # reduced mass
+zeta = 4.0*mu/(m1+m2) # See Loeb 2010 near eqn.1
+print 'zeta=', zeta 
+
+M = 1.0e10 # Msun; total mass of the equal-mass binary
+Mdot = 1.0 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3
+coeff /= zeta # for unequal-mass binary; see Loeb 2010 eqn. 4
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
+save_pdf = False
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
[…]

### 2015-07-08 jet3d_image_fulljet.py
Make an image of both (fwd and bckwd) jets

@@ -1,0 +1,196 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+from matplotlib import cm
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+beta = 0.9 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+M = 1.0e8 # Msun; total mass of the equal-mass binary
+Mdot = 1.0 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3 
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+#output_filename = 'jet_check' 
+save_pdf = True
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
+    return angle #*180.0/np.pi # degrees 
+
+case = int(sys.argv[1])
+if case == 0: 
+    t = np.logspace(-1.0,1.0,10000000)
+    output_filename += '_full'
[…]

### 2015-07-08 gw.py
Make evolution plot for paper

@@ -2,14 +2,21 @@ import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
 mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
+mpl.rcParams['font.size'] = '22'
 import matplotlib.pyplot as plt
 import numpy as np
+
+mygray="#808080"
+myblue="#3D7AB8"
+mygreen="#99CC66"
+myred="#CC4D33"
+myviolet="#7171B7"
+mybeige="#CC9966"
 
 M = 1.0e8 # Msun; total mass of the equal-mass binary 
 
 pcto_10to16cm = 0.003241
-Mdot = 10.0 # Eddington units
+Mdot = 1.0 # Eddington units
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 a0 /= pcto_10to16cm # 1.0e16 cm
 coeff = -2.56e5/(M*1.0e-8)**3 
@@ -81,7 +88,7 @@ if debug:
         print t[i], a_gas[i]
         i += 1
 
-case = 4
+case = 5
 if case == 0: 
         
     # Plot showing evolution of a.
@@ -219,4 +226,45 @@ elif case == 4:
     ax.set_xlabel('$t$ [yr]',labelpad=15) 
     ax.set_ylabel('$a$ [pc]')
 
-plt.show()
+elif case == 5: 
+
+    fig = plt.figure(figsize=(7, 7), dpi=100)
+    plt.subplots_adjust(hspace=0.001)
+    
+    ax = fig.add_subplot(3, 1, 1)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    a *= 0.003241
+    ax.set_xlim(1.0e1,1.0e7)
+    ax.plot(t,a,c=myred,lw=3)
+    ax.set_ylabel('$a$ [pc]')
+    yticks = ax.yaxis.get_major_ticks()
+    yticks[1].set_visible(False)
+    ax.set_xticklabels('')
+    
+    ax = fig.add_subplot(3, 1, 2)
+    ax.tick_params('both', which='major', length=7, width=1)
+    ax.tick_params('both', which='minor', length=3, width=1)
[…]

### 2015-07-08 jet3d_check2.py
Make plots for paper

@@ -2,7 +2,7 @@ import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
 mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
+mpl.rcParams['font.size'] = '22'
 import matplotlib.pyplot as plt
 import numpy as np
 import math 
@@ -14,7 +14,7 @@ beta = 0.90 # jet speed in units of c
 d = 100.0 # Mpc; Distance between jet and observer
 
 M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 0.1 # Eddington units
+Mdot = 1.0 # Eddington units
 #a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
@@ -105,19 +105,24 @@ y_obsb = yb/(1.0-velx/c)
 z_obsb = zb/(1.0-velx/c)
 
 phi_y_obsb = y_obsb/d * 180.0/np.pi * 3600.0 # arcsec
-phi_z_obsb = z_obsb/d * 180.0/np.pi * 3600.0 # arcsec 
+phi_z_obsb = z_obsb/d * 180.0/np.pi * 3600.0 # arcsec
+
+phi_y_obsb *= 1.0e3 # mas
+phi_z_obsb *= 1.0e3 # mas 
+phi_y_obs *= 1.0e3 # mas
+phi_z_obs *= 1.0e3 # mas 
 
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
 #ax.set_ylim(-0.01,0.01)
 if case==1:
-    ax.set_ylim(-0.1,0.1)
+    ax.set_ylim(-50.0,50.0)
     #ax.set_xlim(-0.01,0.01)
 ax.plot(phi_z_obs,phi_y_obs,c='k',lw=1,rasterized=True)
 if case!=1: 
     ax.plot(phi_z_obsb,phi_y_obsb,c='k',lw=1,rasterized=True)
-ax.set_xlabel('arcsec',labelpad=15)
-ax.set_ylabel('arcsec',labelpad=15)
+ax.set_xlabel('mas',labelpad=15)
+ax.set_ylabel('mas')
 
[…]

### 2015-07-08 jet3d_image.py

@@ -2,7 +2,7 @@ import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
 mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
+mpl.rcParams['font.size'] = '22'
 import matplotlib.pyplot as plt
 import numpy as np
 import math 
@@ -26,7 +26,8 @@ coeff = -2.56e5/(M*1.0e-8)**3
 d *= 1.0e3 # kpc
 
 #output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+#output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+output_filename = 'jet_check' 
 save_pdf = True
 
 #psi *= np.pi/180.0 # radians 
@@ -111,26 +112,39 @@ intensity = np.array([0.0 if x!=x else x for x in intensity])
 max_intensity = intensity.max()
 intensity /= max_intensity
 
+intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
+
 # fig = plt.figure(figsize=(7, 7), dpi=100)
 # ax = fig.add_subplot(1, 1, 1)
 # plt.plot(t_observed, intensity)
 # ax.set_xscale('log')
 # ax.set_yscale('log')
 
+phi_y_obs *= 1.0e3 # mas
+phi_z_obs *= 1.0e3 # mas 
+
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
 if case==1:
-    ax.set_ylim(-0.1,0.1)
-    ax.set_xlim(-0.01,0.03)
+    ax.set_ylim(-50,50)
+    ax.set_xlim(-10,30)
 
-s=plt.scatter(phi_z_obs,phi_y_obs,s=1,marker='.',edgecolor='none',rasterized=True,c=intensity,cmap=cm.Reds)
-ax.set_xlabel('arcsec',labelpad=15)
-ax.set_ylabel('arcsec',labelpad=15)
+s=plt.scatter(phi_z_obs,phi_y_obs,s=4,marker='o',edgecolor='none',vmax=0.0,
[…]

### 2015-07-21 jet3d_image_fulljet.py
Change parameter values to those used in paper

@@ -11,11 +11,11 @@ from matplotlib import cm
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 #psi = 20.0 # degrees
-i = 40.0 # degrees
+i = 30.0 # degrees
 beta = 0.9 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
-M = 1.0e8 # Msun; total mass of the equal-mass binary
+M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
 #a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 

### 2015-07-21 Paper/paper.tex
Add acknowledgements

@@
 \documentclass[a4paper,fleqn,usenatbib]{mnras}
-
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-%% \usepackage{newtxtext,newtxmath}
-
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%% \usepackage{mathptmx}
-%% \usepackage{txfonts}
-
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
 \usepackage[T1]{fontenc}
 \usepackage{ae,aecompl}
-
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
 \title[Chirping Jets from Black Hole Binaries]{Radio Crickets:
@@
 \author[Kulkarni and Loeb]{
-Girish Kulkarni,$^{1}$\thanks{E-mail: kulkarni@ast.cam.ac.uk}
+Girish Kulkarni$^{1}$\thanks{E-mail: kulkarni@ast.cam.ac.uk}
 and Abraham Loeb$^{2}$\thanks{E-mail: aloeb@cfa.harvard.edu}
@@
 
-% These dates will be filled out by the publisher
 \date{Accepted ---. Received ---; in original form ---}
 
-% Enter the current year, for the copyright statements etc.
 \pubyear{2015}
 
-% Don't change these lines
 \begin{document}
@@
 
-% Abstract of the paper
 \begin{abstract}
@@
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
 \begin{keywords}
@@
 \end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
 
@@
 
-% Any more references to add here?
 General relativisitic (GR) numerical simulations can predict the
@@
 
-% Replace `LISA' with `eLISA` or `NGO'?  Mention pulsar timing arrays?
 In addition to testing GR in the strong field limit, observations of
@@
 \end{equation}
-% Mention typical size of accretion disks and typical ISCO radius
-% above?
 
@@
   \begin{tabular}{cc}
+    % jet3d_check2.py (case 0) 
     \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
+    % jet3d_image_fulljet.py (case 1) 
     \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
@@
 \end{equation}
-We assume $\delta=1$ and $\alpha=1$, hereafter. 
-% Need to justify parameter values.
-% Comment on relativistic corrections
+We assume $\delta=1$ and $\alpha=1$, hereafter
+\citep{1982ApJ...262..478G}.
 
@@
 morphology with an increasing opening angle and increasingly rapid
-periodic variation together with a twisted shape.
+periodic variation, together with a twisted shape.
 Figure~\ref{fig:jet_mdot1.0} shows these three features for an
@@
 milli-arcsecond.  The duration of the binary evolution captured by
-Figure~\ref{fig:jet_mdot1.0} is 100 yr.
+Figure~\ref{fig:jet_mdot1.0} is 100 yr.  Although the system can be
+rescaled to different masses, times, and separation, the initial
+separation in Figure~\ref{fig:jet_mdot1.0} is $\sim 10^{-3}$~pc, when
+the binary enters it GW-dominated phase.
 
-% Add discussion about the chirp of the signal
+The increasing orbital speed of the binary leads to an increase in the
+jet's precession angular velocity.  This increases the helicity of the
+jet such that the jet is wound up increasingly tightly closer to the
+source.  This gives rise to a ``chirping'' morphology as seen in
+Figure~\ref{fig:jet_mdot1.0}.  The chirp is easier to resolve in jets
+that are closer to the line of sight, where superluminal motion can
+magnify the forward jet.  A chirping black hole is an unambiguous
+signature of a rapidly evolving binary BH; the chirp is direct
+reflection of the orbital velocity evolution in
+Figure~\ref{fig:evolve}.  Together, the three features of the jet
+morphology in Figure~\ref{fig:jet_mdot1.0}, viz., conical jet geometry
+with an increasing opening angle, a radio chirp geometry, and an
+asymmetrical shape, directly probe the dynamics of the binary in its
+GW phase. Observations of this milliarcsecond geometry can thus
+constrain GR in the strong field limit.
 
@@
 
-% Give redshift numbers
 These milli-arcsecond features can be resolved using long baseline
@@
 redshift, which would be difficult to detect with this method.
-However, future detectors such as eLISA are sensitvie to GW emission
+However, future detectors such as eLISA are sensitive to GW emission
 from $10^7$--$10^9$ M$_\odot$ BHs \citep{2013CQGra..30x4009S}.  Pulsar
@@
 
-% Mention old observations which show similar features
-% There is a todo in a PS email from Avi
+We make the code implementing the model presented in this Letter
+available at \url{https://github.com/gkulkarni/JetMorphology} under
+the MIT open source software license.  This code can be used to
+explore the parameter space of the model.
 
@@
 \bibliographystyle{mnras}
-\bibliography{refs} % if your bibtex file is called example.bib
+\bibliography{refs}
 
-% Don't change these lines
-\bsp	% typesetting comment
+\bsp
 \label{lastpage}

### 2015-07-21 gas.py
Clean up

@@ -42,8 +42,6 @@ def half_opening_angle_observed(psi):
     inclination = 15.0*np.pi/180.0
     return psi/np.sin(inclination)
 
-#psi_observed = half_opening_angle_observed(psi)
-
 debug = True
 if debug: 
     data = np.vstack((t,a_gas))
@@ -94,6 +92,5 @@ ax.set_xlim(1.0e4,1.0e8)
 ax.set_xlabel('$t$ [yr]',labelpad=15) 
 ax.set_ylabel(r'$\psi_\mathrm{intrinsic}$ [degrees]')
 plt.savefig("psi_gas.pdf",bbox_inches='tight')
-
 
 plt.show()

### 2015-07-21 jet3d_image_fulljet.py

@@ -10,28 +10,24 @@ import sys
 from matplotlib import cm
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-#psi = 20.0 # degrees
 i = 30.0 # degrees
+i *= np.pi/180.0 # radians 
 beta = 0.9 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
+d = 100.0 # Mpc; Angular distance between jet and observer
+d *= 1.0e3 # kpc
+psi0 = 2.0*np.pi/180.0 # radians
 M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
 coeff = -2.56e5/(M*1.0e-8)**3 
 
-d *= 1.0e3 # kpc
-
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
 output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
-#output_filename = 'jet_check' 
 save_pdf = True
 
-#psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
+# case = 1, 2 zooms into the central region of the jet 
+case = int(sys.argv[1])
 
 c = 3.0e5 # km/s; speed of light
 yrbys = 3.154e7
@@ -47,16 +43,13 @@ def binary_orbital_period(a_16):
     return t
 
 def half_opening_angle_intrinsic(a_16):
-    psi0 = 2.0*np.pi/180.0 # radians 
     angle = np.arcsin(np.sin(psi0)*a0/a_16)
     return angle #*180.0/np.pi # degrees 
 
-case = int(sys.argv[1])
 if case == 0: 
     t = np.logspace(-1.0,1.0,10000000)
     output_filename += '_full'
 elif case == 1:
-    #t = np.linspace(10.0,14.0,10000)
     t = np.logspace(-2.0,2.0,10000000)
     t = t[6000000:9500000]
     output_filename += '_zoom1'
@@ -69,7 +62,6 @@ t0 = t[0]
 def t_binary(time):
[…]

### 2015-07-21 Paper/paper.tex
Correct typos

@@
   We study a novel electromagnetic signature of supermassive black
-  hole binaries whose inspiral stars being dominated by gravitational
+  hole binaries whose inspiral starts being dominated by gravitational
   wave (GW) emission. Combined GW and electromagnetic observations of
@@
   morphology of the jet can be used to infer the binary parameters.
-  For binaries with mass $10^7$--$10^{10}$ M$_\odot$ at redshifts
-  $z<0.5$, the jet precesses along a biconical surface due to the
-  binary's orbital motion.  When the binary enters the GW phase of its
+  The jet precesses along a biconical surface due to the binary's
+  orbital motion.  When the binary enters the GW phase of its
   evolution, the opening angle widens, the jet exhibits milliarcsecond
   scale wiggles and the conical surface of jet precession is twisted
-  due to the superluminal motion.  Monitoring these features in
-  current and archival data will place a lower limit on sources that
-  could be detected by eLISA and Pulsar Timing Arrays.  In the future,
-  microarcsecond interferometry with the Square Kilometer Array will
-  increase the potential usefulness of this technique.
+  due to the superluminal motion.  For binaries with mass
+  $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, monitoring these
+  features in current and archival data will place a lower limit on
+  sources that could be detected by eLISA and Pulsar Timing Arrays.
+  In the future, microarcsecond interferometry with the Square
+  Kilometer Array will increase the potential usefulness of this
+  technique.
 \end{abstract}
@@
 binary's evolution, emission of gravitational waves (GWs) becomes the
-predominant mode angular momentum loss, which results in the
+predominant mode of angular momentum loss, which results in the
 coalescence of the black holes.
@@
 the circumbinary disk due to shocks induced by the sudden mass loss of
-the binary due to GW emission at coalescence
+the binary following GW emission at coalescence
 \citep{2007APS..APR.S1010B}, flares from shocked remnants of accretion
@@
 future detectors such as the Evolved Laser Interferometer Space
-Antenna (eLISA)\footnote{\url{https://www.elisascience.org}}, which would be
-capable of observing the peak GW luminosity for BH coalescence of
-$\sim 10^{57}$ erg$/$s up to large cosmological distances
+Antenna (eLISA)\footnote{\url{https://www.elisascience.org}}, which
+would be capable of observing the peak GW luminosity for BH
+coalescence of $\sim 10^{57}$ erg$/$s out to cosmological distances
 \citep{2003CQGra..20S..65H, 2013CQGra..30x4009S}.  Simultaneous
@@
 will yield rich scientific pay-offs \citep{2003CQGra..20S..65H,
-  2005ApJ...629...15H}.  For example, by a redshift measurement of
-eLISA sources could constraint cosmological parameters.  Measurement
-of the black hole masses and spins from the GW signal will constrain
+  2005ApJ...629...15H}.  For example, redshift measurement of eLISA
+sources could constraint cosmological parameters.  Measurement of the
+black hole masses and spins from the GW signal could constrain
 accretion physics during coalescence.  But even before eLISA is
-launched, a search of electromagnetic signatures may point to the
-expected abundance of GW sources.
+launched, a search of electromagnetic signatures may help calibrate
+the expected abundance of GW sources.
 
 In this Letter we study the morphology of a jet emitted by one of the
-binary members during its final stage of evolution before coalescence,
+binary members during its final phase of evolution before coalescence,
 i.e., when the binary's dynamics is dominated by GW emission.  We show
@@
 $a$ around each other.  Let $M=M_1+M_2$.  For simplicity, we consider
-an equal-mass binary ($M_1=M_2$) on a circular orbit. We will consider
-unequal black holes masses and elliptical orbits below.  The orbital
+an equal-mass binary ($M_1=M_2$) on a circular orbit. The orbital
 speed of each black hole is given by \citep{2010PhRvD..81d7503L}
@@
   \caption{Jet and orbital configuration. The orbital velocity
-    $\mathbf{v}_\mathrm{orb}$ causes the jet to precess as the BH
-    moves along the orbit. The net jet velocity is
+    $\mathbf{v}_\mathrm{orb}$ causes the jet to precess as its source
+    BH moves along its binary orbit. The net jet velocity is
     $\tilde{\mathbf{v}}_\mathrm{jet}={\mathbf{v}}_\mathrm{orb}+{\mathbf{v}}_\mathrm{jet}$,
-    where $\mathbf{v}_\mathrm{jet}$ is its ejection velocity at the
-    source.}
+    where $\mathbf{v}_\mathrm{jet}$ is its ejection velocity of the
+    jet material at the source.}
   \label{fig:BH_orbit}
@@
     introduces the jet precision angle $\psi$.  Without the orbital
-    motion, $\psi=0$ and the jet traces a cylindrical surface.  With
+    motion, $\psi=0$, and the jet traces a cylindrical surface.  With
     the orbital motion, the net red velocity vector oscillates around
@@
     exagerated here.  In practice, $v_\mathrm{orb}\ll
-    v_\mathrm{jet}\sim c$.  This figure then shows that
+    v_\mathrm{jet}\sim c$.  The figure indicates that
     $\sin{\psi}\sim\tan{\psi}=v_\mathrm{orb}\cos{\chi}/v_\mathrm{jet}$.}
@@
 \end{equation}
-When the binary separation is where gravitational wave-induced
-inspiral begins, it has a velocity given by equation~(\ref{eqn:v}),
-and the separation between the two black holes at this instant is
-given by equation~(\ref{eqn:a}).  Also, from equation
-(\ref{eqn:vorb}), the orbital velocity of each BH evolves as
+When the binary separation reaches the point where gravitational
+wave-induced inspiral begins, the binary has a velocity given by
+equation~(\ref{eqn:v}), and a separation given by
+equation~(\ref{eqn:a}).  From equation (\ref{eqn:vorb}), the orbital
+velocity of each BH evolves as
 \begin{equation}
@@
   equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
-  assumed $i=40^\circ$, $\beta=0.9$, and $d = 100$~Mpc.  The jets are
-  twisted because of the apparant superluminal motion.  Also evident
-  is the opening of the jet and the radio jet chirp.  Right panel
-  shows normalised brightness of the inner 30~mas region of the
-  forward jet.}
+  assumed $i=40^\circ$, $\beta=0.9$, and angular diameter distance $d
+  = 100$~Mpc.  The jets are twisted because of the apparant
+  superluminal motion.  Also evident is the opening of the jet and the
+  radio jet chirp.  Right panel shows normalised brightness of the
+  inner 30~mas region of the forward jet.}
 \label{fig:jet_mdot1.0}
@@
 We now ask if it would be possible to observe the opening of the jet
-morphology.  In order to do this, it is necessary to map the above
+morphology.  For this purpose, it is necessary to map the above
 calculations to the observer's frame \citep{1982ApJ...262..478G}.  At
@@
 observer's frame such that the $yz$ is the plane of the sky.  Further,
-the jet is oriented such that that the $y^\prime$ axis coincides with
-the $y$ axis, the original direction of jet emission is at an angle
-$i$ relative to the line of sight ($x$-axis).  For simplicity, we also
+the jet is oriented such that the $y^\prime$ axis coincides with the
+$y$ axis, the original direction of jet emission is at an angle $i$
+relative to the line of sight ($x$-axis).  For simplicity, we also
 assume that the $z^\prime$ axis is in the $xz$ plane.  Then the jet
@@
 \end{equation}
-respectively.  In angular coordinates, the motion on the sky is given
-by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular diameter
-distance of the binary.  Because of superluminal motion, the forward
-jet is stretched while the backward jet is compressed.
+respectively.  In angular coordinates, the motion of the jet on the
+sky is given by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular
+diameter distance of the binary.  Because of superluminal motion, the
+forward jet is stretched while the backward jet is compressed.
 
-Finally, we assume that the emission from each jet particle is
-optically thin at the frequency of observation with a spectrum in the
-rest frame of the particle given by $P(\nu)\propto \nu^{-\alpha}$
-\citep{1982ApJ...262..478G}.  The observed flux density $S(\nu)$ in erg
-cm$^{-2}$ Hz$^{-1}$ is given by
+Finally, we assume that the emission from each jet is optically thin
+at the frequency of observation with a spectrum in the rest frame of
+the particle given by $P(\nu)\propto \nu^{-\alpha}$
+\citep{1982ApJ...262..478G}.  The observed flux density $S(\nu)$ in
+erg cm$^{-2}$ Hz$^{-1}$ is given by
 \begin{equation}
@@
 $[\gamma(1-\beta\cos\phi)]^{-1}$ and $S_r$ is the rest frame flux
-density.  Further to consider the effects of jet evolution we assume
+density.  Further, to consider the effects of jet evolution we assume
 that $S_r$ decreases with time according to a simple power law
@@
 
-As the binary evolves, its increasing orbital period and speed, and
-the apparant superluminal motion effect, together result in a jet
-morphology with an increasing opening angle and increasingly rapid
+As the binary evolves, the combination of increasing orbital period
+and speed, and the apparant superluminal motion effect, results in a
+jet morphology with an increasing opening angle and increasingly rapid
 periodic variation, together with a twisted shape.
@@
 100$~Mpc, which corresponds to redshift $z\sim 0.3$.  The forward jet
-streched on the sky while the backward jet is compressed due to the
+streches on the sky while the backward jet is compressed due to the
 superluminal motion.  The right panel of Figure~\ref{fig:jet_mdot1.0}
@@
 magnify the forward jet.  A chirping black hole is an unambiguous
-signature of a rapidly evolving binary BH; the chirp is direct
+signature of a rapidly evolving binary BH system; the chirp is direct
 reflection of the orbital velocity evolution in
 Figure~\ref{fig:evolve}.  Together, the three features of the jet
-morphology in Figure~\ref{fig:jet_mdot1.0}, viz., conical jet geometry
-with an increasing opening angle, a radio chirp geometry, and an
-asymmetrical shape, directly probe the dynamics of the binary in its
-GW phase. Observations of this milliarcsecond geometry can thus
+morphology in Figure~\ref{fig:jet_mdot1.0}, namely, conical jet
+geometry with an increasing opening angle, a radio chirp geometry, and
+an asymmetrical shape, directly probe the dynamics of the binary in
+its GW phase. Observations of such a milliarcsecond geometry can thus
 constrain GR in the strong field limit.
@@
 jet morphology of a binary that is entering its GW-induced phase of
-inspiral shows a distinct chirping morphology. Jets in these binaries
-have a biconical morphology.  After the binary enters the GW phase,
-the opening angle of cone rapidly increases within a distance of tens
-of milli-arcseconds.  In addition, the periodicity of the jet's motion
-on the conical surface increases as the orbital period of the binary
-goes up in this stage of its evolution.  Finally, due to apparent
-superluminal motion and the widening opening angle, the jet is twisted
-into a peculiar morphology.  The jet geometry can be used to infer the
-set of dynamical and geometrical parameters, listed in
-Table~\ref{table:params}.
+inspiral exhibits a distinct chirping morphology. Jets in these
+binaries have a biconical morphology.  After the binary enters the GW
+phase, the opening angle of cone rapidly increases within a separation
+of tens of milli-arcseconds at a distance of 100~Mpc.  In addition,
+the periodicity of the jet's motion on the conical surface increases
+as the orbital period of the binary goes up in this stage of its
+evolution.  Finally, due to the apparent superluminal motion and the
+widening opening angle, the jet is twisted into a peculiar morphology.
+The jet geometry can be used to infer the set of dynamical and
+geometrical parameters, listed in Table~\ref{table:params}.
 
-These milli-arcsecond features can be resolved using long baseline
-radio interferometry \citep{2001ChJAA...1..236Q}.  At higher
-redshifts, jets are fainter and the wiggles become more difficult to
-resolve.  However, it might be possible to resolve binaries with very
-unequal mass ratios, eccentric orbits, or suitable inclinations.  The
-future Square Kilometer Array (SKA) should make microarcsecond
-resolution possible, which could increase the detection rate of these
-objects \citep{2014arXiv1412.5971P}.
+Milli-arcsecond features can be resolved using long baseline radio
+interferometry \citep{2001ChJAA...1..236Q}.  At higher redshifts, jets
+are fainter and the wiggles become more difficult to resolve.
+However, it might be possible to resolve binaries with very unequal
+mass ratios, eccentric orbits, or suitable inclinations.  The future
+Square Kilometer Array (SKA) should make microarcsecond resolution
+possible, which could increase the detection rate of these objects
+\citep{2014arXiv1412.5971P}.
 
@@
 brightest GW sources in the universe.  Current ground based GW
-detectors such as LIGO are sensitive to low-mass BH binaries at high
-redshift, which would be difficult to detect with this method.
+detectors such as LIGO are sensitive to intermediate-mass BH binaries
+at high redshift, which would be difficult to detect with our method.
 However, future detectors such as eLISA are sensitive to GW emission
@@
 Timing Arrays are sensitive to even higher binary masses.  Monitoring
-of jet morphologies with features outlined in this paper will put a
-lower bound on the expected abundance of bright GW sources for this
-experiments. Moreover, combined GW and electromagnetic observations of
+of jet morphologies for the features outlined in this letter will put
+a lower bound on the expected abundance of bright GW
+sources. Moreover, combined GW and electromagnetic observations of
 compact binary mergers should enable detailed studies of astrophysical
-processes in the strong-field gravity regime
-\citep{2005ApJ...629...15H}.
+processes in the strong-field GR regime \citep{2005ApJ...629...15H}.
 
-We make the code implementing the model presented in this Letter
+We make the code for implementing the model presented in this Letter
 available at \url{https://github.com/gkulkarni/JetMorphology} under

### 2015-07-21 gas_shrink.py
Remove junk code

@@ -1,12 +1,0 @@
-
-import numpy as np
-from scipy.integrate import ode 
-
-def f(t, y):
-    return np.exp(t)
-
-t0, y0 = 0.0, 1.0
-r = ode(f).set_integrator('vode', method='bdf', with_jacobian=True)
-r.set_initial_value(y0, t0)
-print r.integrate(1.0)
-

### 2015-07-21 jet.py
Delete redundant code

@@ -1,226 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-case = int(sys.argv[1])
-
-if case == 0:
-
-    """ Just some example jets>"""
-
-    fig = plt.figure(figsize=(7, 7), dpi=80)
-
-    ax = fig.add_subplot(1, 1, 1) 
-    
-    x = np.arange(0, 20*np.pi, 0.1)
-    y = np.sin(x)*x*np.tan(0.1)
-    plt.plot(x,y)
-    
-    y = np.sin(x)*x*np.tan(0.4)
-    plt.plot(x,y)
-
-    plt.show()
-
-elif case == 1:
-
-    """ cone angle narrowing down"""
-
-    fig = plt.figure(figsize=(7, 7), dpi=80)
-    ax = fig.add_subplot(1, 1, 1)
-
-    x = np.arange(0, 20*np.pi, 0.1)
-
-    y = np.sin(x)*x
-    for i, s in enumerate(y):
-        rate = 0.02
-        angle = rate*x[i]
-        y[i] = s*np.tan(angle)
-    plt.plot(x,y)
-    
-    plt.show()
-
-elif case == 2:
-
-    """ cone angle narrowing down with differing rates"""
-
-    fig = plt.figure(figsize=(7, 7), dpi=80)
-    ax = fig.add_subplot(1, 1, 1)
-
-    x = np.arange(0, 20*np.pi, 0.1)
-
-    y = np.sin(x)*x
-    for i, s in enumerate(y):
-        if x[i] > 45.0: 
-            rate = 0.02
-        else:
-            rate = 0.01
-        angle = rate*x[i]
-        y[i] = s*np.tan(angle)
-            
-    plt.plot(x,y)
-    
-    plt.show()
-    
-elif case == 3:
-
-    """ different periods """ 
-
-    fig = plt.figure(figsize=(7, 7), dpi=80)
-    ax = fig.add_subplot(1, 1, 1)
-
-    x = np.arange(0, 20*np.pi, 0.1)
-    y = np.zeros(x.size)
-
-    for i, s in enumerate(y):
-        angle = 0.1
-        p = 0.5
[…]

### 2015-07-21 jet3d.py

@@ -1,175 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-psi = 20.0 # degrees
-i = 40.0 # degrees
-theta = 0.0 # degrees
-beta = 0.20 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-a0 = 1 # 1.0e16 cm
-coeff = -2.56e5
-
-d *= 1.0e3 # kpc
-
-output_filename = 'jet_i%2d_psi%2d_beta_%3.2f' % (i, psi, beta)
-save_pdf = True
-
-psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-theta *= np.pi/180.0 # radians
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    psi0 = 20.0*np.pi/180.0 # radians 
-    psi = np.arcsin(np.sin(psi0)*a0/a_16)
-    return psi#*180.0/np.pi # degrees 
-
-case = 1
-
-if case == 0: 
-    t = np.logspace(-1.0,2.0,10**4) # yr
-    t0 = t[0]
-    print t.size
-
-elif case == 1: 
-    t1 = np.logspace(-2.0,3.8,num=100)
-    t2 = np.logspace(3.8,4.80619,num=100000)
-    tb = np.concatenate((t1,t2))
-    t0 = tb[0]
-    a = binary_separation_gw(tb)
-    pb = binary_orbital_period(a)
-    psi_chirp = half_opening_angle_intrinsic(a)
-
-    t = tb[-100000:-1000]
[…]

### 2015-07-21 jet3d_check.py
Remove redundant code

@@ -1,113 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-psi = 20.0 # degrees
-i = 40.0 # degrees
-theta = 0.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-a0 = 1 # 1.0e16 cm
-coeff = -2.56e5
-
-d *= 1.0e3 # kpc
-
-output_filename = 'jet_i%2d_psi%2d' % (i, psi)
-save_pdf = False
-
-psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-theta *= np.pi/180.0 # radians
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    psi0 = 20.0*np.pi/180.0 # radians 
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
-case = 1
-
-if case == 0:
-    # Simple sampling
-    t = np.logspace(-1.0,2.0,10**4) # yr
-    t0 = t[0]
-    print t.size
-
-elif case == 1:
-    # Sampling increases at small separations
-    t1 = np.logspace(-2.0,3.8,num=100)
-    t2 = np.logspace(3.8,4.80619,num=100000)
-    tb = np.concatenate((t1,t2))
-    t0 = tb[0]
-    a = binary_separation_gw(tb)
-    pb = binary_orbital_period(a)
-    psi_chirp = half_opening_angle_intrinsic(a)
-
-    t = tb[-100000:-5]
-    maxt = t.max()
-    t = maxt - t
[…]

### 2015-07-21 jet3d_check2.py

@@ -1,178 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-#psi = 20.0 # degrees
-i = 40.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
-a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
-pcto_10to16cm = 0.003241
-a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
-
-d *= 1.0e3 # kpc
-
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
-save_pdf = True
-
-#psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    psi0 = 2.0*np.pi/180.0 # radians 
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
-case = int(sys.argv[1])
-if case == 0: 
-    t = np.logspace(-2.0,2.0,10000000)
-    output_filename += '_full'
-elif case == 1:
[…]

### 2015-07-21 jet3d_image.py

@@ -1,154 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-from matplotlib import cm
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-
-#psi = 20.0 # degrees
-i = 40.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
-a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
-pcto_10to16cm = 0.003241
-a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
-
-d *= 1.0e3 # kpc
-
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
-#output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
-output_filename = 'jet_check' 
-save_pdf = True
-
-#psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    psi0 = 2.0*np.pi/180.0 # radians 
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
-case = int(sys.argv[1])
-if case == 0: 
-    t = np.logspace(-2.0,2.0,10000000)
-    output_filename += '_full'
[…]

### 2015-07-21 jet3d_nochip.py

@@ -1,116 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-psi = 2.0 # degrees
-i = 40.0 # degrees
-theta = 0.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-d *= 1.0e3 # kpc
-
-output_filename = 'jet_i%2d_psi%2d_beta_%3.2f' % (i, psi, beta)
-save_pdf = False
-
-psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-theta *= np.pi/180.0 # radians
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-t = np.logspace(-1.0,2.0,10**6) # yr
-
-def Omega(t):
-    # Angular velocity of jet precession 
-    period = 10.0 # yr
-    return 2.0*np.pi*t/period # radians
-
-def vel(t):
-    # We are following geometry from Gower et al. Figure 1.
-    vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(t)) + np.cos(psi)*np.cos(i))
-    vy = beta*c*np.sin(psi)*np.sin(Omega(t))
-    vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(t)))
-    return sign*vx, sign*vy, sign*vz # km/s
-
-# sign sets Forward or backward jet
-sign = 1
-
-velx, vely, velz = vel(t)
-
-x = np.zeros(t.size)
-y = np.zeros(t.size)
-z = np.zeros(t.size)
-
-y = vely*t*yrbys/kpcbykm # kpc
-z = velz*t*yrbys/kpcbykm # kpc
-
-y_obs = y/(1.0-velx/c)
-z_obs = z/(1.0-velx/c)
-
[…]

### 2015-07-21 jet3d_zoom.py

@@ -1,117 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-psi = 20.0 # degrees
-i = 40.0 # degrees
-theta = 0.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-a0 = 1 # 1.0e16 cm
-coeff = -2.56e5
-
-d *= 1.0e3 # kpc
-
-output_filename = 'jet_i%2d_psi%2d' % (i, psi)
-save_pdf = True
-
-psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
-theta *= np.pi/180.0 # radians
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    psi0 = 20.0*np.pi/180.0 # radians 
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
-def t_binary(time):
-    return np.abs(time-64001.477505390176) # yr
-
-def Omega(time):
-    # Angular velocity times time for jet precession 
-    period = binary_orbital_period(binary_separation_gw(t_binary(time))) # yr 
-    return 2.0*np.pi*time/period # radians
-
-def vel(time):
-    # We are following geometry from Gower et al. Figure 1.
[…]

### 2015-07-21 scratch.py

@@ -1,150 +1,0 @@
-time_prev = t[0]
-vx_prev, vy_prev, vz_prev = vel(time_prev)
-pos_x = 0.0
-pos_y = 0.0
-pos_z = 0.0
-x[0] = pos_x
-y[0] = pos_y
-z[0] = pos_z
-t_counter = 0 
-prevtime = t[t_counter]
-
-phi_z = np.zeros(t.size) 
-phi_y = np.zeros(t.size)
-phi_z[t_counter] = 0.0
-phi_y[t_counter] = 0.0
-angpos_z = 0.0
-angpos_y = 0.0
-
-for time in t[1:]:
-    t_counter += 1 
-    vx_curr, vy_curr, vz_curr = vel(time)
-
-    pos_x += (time-prevtime)*(vx_curr+vx_prev)*0.5*yrbys/kpcbykm # kpc
-    pos_y += (time-prevtime)*(vy_curr+vy_prev)*0.5*yrbys/kpcbykm # kpc
-    pos_z += (time-prevtime)*(vz_curr+vz_prev)*0.5*yrbys/kpcbykm # kpc 
-
-    x[t_counter] = pos_x
-    y[t_counter] = pos_y
-    z[t_counter] = pos_z
-
-    prevtime = time
-    vx_prev = vx_curr
-    vy_prev = vy_curr
-    vz_prev = vz_curr
-
-t_observer = t/(1.0-velx/c) # yr
-
-
-# now calculate anglar motion on the sky 
-phi_z = np.zeros(t.size) 
-phi_y = np.zeros(t.size)
-time_prev = t[0]
-vx_prev, vy_prev, vz_prev = vel(time_prev)
-
-def binary_separation_gw(t):
-    coeff = -2.56e5 
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a 
-
-def half_opening_angle_intrinsic_old(a_16):
-    psi0 = 1.0*np.pi/180.0 # radians 
-    psi = np.arcsin(np.sin(psi0)*a0/a_16)
-    return psi*180.0/np.pi # degrees
-
-def half_opening_angle_intrinsic(t):
-    y = 0.680681*t+1.93193 # degrees
-    return y 
-
-
-t0 = t[0]
-a0 = 1 # 1.0e16 cm
-
-a = binary_separation_gw(t)
[…]

### 2015-07-21 gw.py
Clean up

@@ -1,3 +1,10 @@
+"""
+File: gw.py 
+
+Produces Figure 3.
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -32,12 +39,6 @@ t_merge=coeff*a0**4/4.0
 print 'coeff=', coeff
 print 't_merge=', t_merge
 print 'log(t_merge)=', np.log10(-t_merge)
-
-# t1 = np.logspace(-2.0,4.80617,num=100)
-# a1 = binary_separation_gw(t1) 
-
-# t2 = np.logspace(4.80617,4.80619,num=300)
-# a2 = binary_separation_gw(t2) 
 
 t1 = np.logspace(-2.0,3.8,num=100)
 a1 = binary_separation_gw(t1) 
@@ -79,7 +80,6 @@ def binary_orbital_period(a_16):
 
 P = binary_orbital_period(a)
 
-
 debug = False
 if debug: 
     data = np.vstack((t,a_gas))
@@ -89,10 +89,9 @@ if debug:
         i += 1
 
 case = 5
+
 if case == 0: 
         
-    # Plot showing evolution of a.
-
     fig = plt.figure(figsize=(7, 7), dpi=100)
     ax = fig.add_subplot(1, 1, 1)
     ax.tick_params('both', which='major', length=7, width=1)
@@ -105,9 +104,6 @@ if case == 0:
     ax.set_xlabel('$t$ [yr]',labelpad=15) 
     ax.set_ylabel('$a$ [pc]')
     plt.savefig("a_gw.pdf",bbox_inches='tight')
-    #plt.show()
-
-    # Plot showing evolution of a_gas.
 
     fig = plt.figure(figsize=(7, 7), dpi=100)
     ax = fig.add_subplot(1, 1, 1)
@@ -118,8 +114,6 @@ if case == 0:
     ax.plot(t,a_gas)
     ax.set_xlabel('$t$ [yr]') 
     ax.set_ylabel('$a$ [10^{16} cm]')
-
-    # Plot showing evolution of v_orb.
 
     fig = plt.figure(figsize=(7, 7), dpi=100)
     ax = fig.add_subplot(1, 1, 1)
@@ -133,9 +127,6 @@ if case == 0:
     ax.set_ylabel('$v_\mathrm{orbital}$ [km$/$s]')
     plt.savefig("v_orb.pdf",bbox_inches='tight')
 
[…]

### 2015-07-21 gas.py
Clean up

@@ -1,3 +1,10 @@
+"""
+File: gas.py 
+
+Evolution of a BH binary in the gas-dominated phase.
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -50,8 +57,6 @@ if debug:
         print t[i], a_gas[i], P[i]
         i += 1
 
-# Plot showing evolution of a_gas.
-    
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
 ax.tick_params('both', which='major', length=7, width=1)
@@ -65,8 +70,6 @@ ax.set_xlabel('$t$ [yr]')
 ax.set_ylabel('$a$ [pc]')
 plt.savefig("a_gas.pdf",bbox_inches='tight')
 
-# Plot showing evolution of v_orb.
-
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
 ax.tick_params('both', which='major', length=7, width=1)
@@ -78,9 +81,6 @@ ax.set_xlim(1.0e4,1.0e8)
 ax.set_xlabel('$t$ [yr]',labelpad=15) 
 ax.set_ylabel('$v_\mathrm{orbital}$ [km$/$s]')
 plt.savefig("v_orb_gas.pdf",bbox_inches='tight')
-
-# Plot showing evolution of the intrinsic half-opening angle of the
-# conical jet
 
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)

### 2015-07-21 gw.py

@@ -1,7 +1,7 @@
 """
 File: gw.py 
 
-Produces Figure 3.
+Evolution of a BH binary in the GW-dominated phase (Figure 3).
 
 """
 
@@ -252,4 +252,4 @@ elif case == 5:
     
     plt.savefig("evolve.pdf",bbox_inches='tight')
     
-#plt.show()
+plt.show()

### 2015-07-21 jet.py
Clean up

@@ -1,3 +1,12 @@
+"""
+
+File: jet.py 
+
+Jet morphology from a BH binary that is in the GW-dominated phase of
+its inspiral (Figures 4). 
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -7,35 +16,27 @@ import matplotlib.pyplot as plt
 import numpy as np
 import math 
 import sys
+from matplotlib import cm
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-#psi = 20.0 # degrees
-i = 40.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-m1 = 0.5e10 #msun
-m2 = m1 #msun
-mu = m1*m2/(m1+m2) # reduced mass
-zeta = 4.0*mu/(m1+m2) # See Loeb 2010 near eqn.1
-print 'zeta=', zeta 
-
+i = 30.0 # degrees
+i *= np.pi/180.0 # radians 
+beta = 0.9 # jet speed in units of c 
+d = 100.0 # Mpc; Angular distance between jet and observer
+d *= 1.0e3 # kpc
+psi0 = 2.0*np.pi/180.0 # radians
 M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3
-coeff /= zeta # for unequal-mass binary; see Loeb 2010 eqn. 4
+coeff = -2.56e5/(M*1.0e-8)**3 
 
-d *= 1.0e3 # kpc
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+save_pdf = True
 
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
-save_pdf = False
-
-#psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
[…]

### 2015-07-21 jet_ellipticity.py

@@ -1,3 +1,12 @@
+"""
+
+File: jet_ellipticity.py
+
+Jet morphology from a BH binary on an elliptical orbit that is in the
+GW-dominated phase of its inspiral (Figures 4).
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -8,20 +17,19 @@ import numpy as np
 import math 
 import sys
 
-#psi = 20.0 # degrees
 i = 40.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
+psi0 = 2.0*np.pi/180.0 # radians 
 
 m1 = 0.5e10 #msun
 m2 = m1 #msun
 mu = m1*m2/(m1+m2) # reduced mass
 zeta = 4.0*mu/(m1+m2) # See Loeb 2010 near eqn.1
-print 'zeta=', zeta
 
 e = 0.6 # orbital eccentricity 
 def F(e):
-    # See Merritt and Milos 2005 eqn 7 
+    # Merritt and Milosavljetvic (2005); Eqn. (7). 
     return (1.0-e*e)**3.5 * (1.0 + 73.0*e*e/24.0 + 37.0*e*e*e*e/96.0)
 
 print 'e=', e
@@ -29,7 +37,6 @@ print 'F(e)=', F(e)
 
 M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
@@ -39,11 +46,9 @@ coeff /= F(e) # eccentricity
 
 d *= 1.0e3 # kpc
 
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
 output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
 save_pdf = False
 
-#psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
 
 c = 3.0e5 # km/s; speed of light
@@ -60,16 +65,14 @@ def binary_orbital_period(a_16):
     return t
 
[…]

### 2015-07-21 jet_unqualmasses.py

@@ -1,3 +1,12 @@
+"""
+
+File: jet_unequalmasses.py
+
+Jet morphology from a BH binary that is in the GW-dominated phase of
+its inspiral (Figures 4).  BH masses are not assumed to be equal.
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -7,27 +16,32 @@ import matplotlib.pyplot as plt
 import numpy as np
 import math 
 import sys
-from matplotlib import cm
-from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-i = 30.0 # degrees
-i *= np.pi/180.0 # radians 
-beta = 0.9 # jet speed in units of c 
-d = 100.0 # Mpc; Angular distance between jet and observer
-d *= 1.0e3 # kpc
-psi0 = 2.0*np.pi/180.0 # radians
+i = 40.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+psi0 = 2.0*np.pi/180.0 # radians 
+
+m1 = 0.5e10 #msun
+m2 = m1 #msun
+mu = m1*m2/(m1+m2) # reduced mass
+zeta = 4.0*mu/(m1+m2) # See Loeb 2010 near eqn.1
+print 'zeta=', zeta 
+
 M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
+coeff = -2.56e5/(M*1.0e-8)**3
+coeff /= zeta # for unequal-mass binary; see Loeb 2010 eqn. 4
 
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
-save_pdf = True
+d *= 1.0e3 # kpc
 
-# case = 1, 2 zooms into the central region of the jet 
-case = int(sys.argv[1])
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
+save_pdf = False
+
+i *= np.pi/180.0 # radians 
 
[…]

### 2015-07-23 Paper/paper.tex
Minor modifications to text

@@
   hole binaries whose inspiral starts being dominated by gravitational
-  wave (GW) emission. Combined GW and electromagnetic observations of
-  compact binary mergers would allow to use them as standard sirens
-  for cosmology and a testing ground of general relativity.  If one of
-  the binary members produces a radio jet, the helical chirping
-  morphology of the jet can be used to infer the binary parameters.
-  The jet precesses along a biconical surface due to the binary's
-  orbital motion.  When the binary enters the GW phase of its
-  evolution, the opening angle widens, the jet exhibits milliarcsecond
-  scale wiggles and the conical surface of jet precession is twisted
-  due to the superluminal motion.  For binaries with mass
-  $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, monitoring these
-  features in current and archival data will place a lower limit on
-  sources that could be detected by eLISA and Pulsar Timing Arrays.
+  wave (GW) emission.  Recent simulations suggest that the binary's
+  member BHs can continue to accrete gas from the circumbinary
+  accretion disk in this phase of the binary's evolution, all the way
+  until coalescence.  If one of the binary members produces a radio
+  jet as a result of accretion, the jet precesses along a biconical
+  surface due to the binary's orbital motion.  When the binary enters
+  the GW phase of its evolution, the opening angle widens, the jet
+  exhibits milliarcsecond scale wiggles, and the conical surface of
+  jet precession is twisted due to apparant superluminal motion.  The
+  rapidly increasing orbital velocity of the binary gives the jet an
+  appearance of a ``chirp.''  This helical chirping morphology of the
+  jet can be used to infer the binary parameters.  For binaries with
+  mass $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, monitoring
+  these features in current and archival data will place a lower limit
+  on sources that could be detected by eLISA and Pulsar Timing Arrays.
   In the future, microarcsecond interferometry with the Square
@@
 \begin{keywords}
-keyword1 -- keyword2 -- keyword3
+  accretion, accretion discs -- black hole physics -- quasars: supermassive black holes
 \end{keywords}
@@
 Binaries of supermassive black holes (BHs) arise naturally as a result
-of mergers of galaxies in the context of the heirarchical galaxy
-formation \citep{2000MNRAS.311..576K, 2002MNRAS.336L..61H,
-  2008ApJ...676...33D, 2012MNRAS.422.1306K}.  After a galaxy merger,
-infalling black holes lose their angular momentum in three stages
-before coalescence \citep{1980Natur.287..307B, 2005LRR.....8....8M,
-  2011ASL.....4..181C}.  In the first stage, the black holes sink to
-the center of the gravitational potential of the merger remnant due to
-dynamical friction and form a gravitationally bound binary.  The newly
-formed binary then continues to lose its angular momentum and shrink
-by scattering ambient gas and stars.  In the final stage of the
-binary's evolution, emission of gravitational waves (GWs) becomes the
+of mergers of galaxies in the context of heirarchical galaxy formation
+\citep{2000MNRAS.311..576K, 2002MNRAS.336L..61H, 2008ApJ...676...33D,
+  2012MNRAS.422.1306K}.  After a galaxy merger, infalling black holes
+lose their angular momentum in three stages before coalescence
+\citep{1980Natur.287..307B, 2005LRR.....8....8M, 2011ASL.....4..181C}.
+In the first stage, the black holes sink to the center of the
+gravitational potential of the merger remnant due to dynamical
+friction and form a gravitationally bound binary.  The newly formed
+binary then continues to lose its angular momentum and shrink by
+scattering ambient gas and stars.  In the final stage of the binary's
+evolution, emission of gravitational waves (GWs) becomes the
 predominant mode of angular momentum loss, which results in the
@@
 \citep{2007APS..APR.S1010B}, flares from shocked remnants of accretion
-disks around recoiled remnant black hole \citep{2007PhRvL..99d1103L,
-  2008ApJ...682..758S, 2009CQGra..26i4032H}, quasi-periodic
-variability in gas accretion in the early stages of black hole merger
-\citep{2008ApJ...672...83M, 2009CQGra..26i4032H, 2012MNRAS.427.2680K},
-evolution in the profiles of broad emission lines shortly before
-merger \citep{2013MNRAS.432.1468M}, characteristic evolution in
-thermal emission due to circumbinary accretion
-\citep{2015MNRAS.446L..36F}, brightening of accretion disk due to
-viscous dissipation of gravitational waves
-\citep{2008PhRvL.101d1101K}, and prompt tidal disruption of stars by
-the recoiled remnant \citep{2011MNRAS.412...75S}.
+disk around the recoiled remnant black hole
+\citep{2007PhRvL..99d1103L, 2008ApJ...682..758S, 2009CQGra..26i4032H},
+quasi-periodic variability in gas accretion in the early stages of
+black hole merger \citep{2008ApJ...672...83M, 2009CQGra..26i4032H,
+  2012MNRAS.427.2680K}, evolution in the profiles of broad emission
+lines shortly before merger \citep{2013MNRAS.432.1468M},
+characteristic evolution in thermal emission due to circumbinary
+accretion \citep{2015MNRAS.446L..36F}, brightening of the circumbinary
+accretion disk due to viscous dissipation of gravitational waves
+\citep{2008PhRvL.101d1101K}, and prompt tidal disruption of ambient
+stars by the recoiled remnant \citep{2011MNRAS.412...75S}.
 
@@
 the expected abundance of GW sources.
-
-In this Letter we study the morphology of a jet emitted by one of the
-binary members during its final phase of evolution before coalescence,
-i.e., when the binary's dynamics is dominated by GW emission.  We show
-that the helical chirping morphology of the jet can be used to infer
-the binary parameters.
-
-\section{Binary parameters}
-
-We study a BH binary formed following a gas-rich merger of two
-galaxies as a result of the processes described above.  We consider
-black holes with masses $M_1$ and $M_2$ in a circular orbit of radius
-$a$ around each other.  Let $M=M_1+M_2$.  For simplicity, we consider
-an equal-mass binary ($M_1=M_2$) on a circular orbit. The orbital
-speed of each black hole is given by \citep{2010PhRvD..81d7503L}
-\begin{equation}
-  v = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2}\!\!\!\! = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
-  \label{eqn:vorb}
-\end{equation}
-where $a_{16}\equiv (a/10^{16}$ cm$)$ and $M_8\equiv (M/10^8$
-M$_\odot)$ are the binary separation and mass,
-respectively.\footnote{Note that $10^{16} \mathrm{cm}\approx 3.4\times
-  10^{-3} \mathrm{pc}$.  The Schwarzchild radius of a $10^8$~M$_\odot$
-  black hole is $9.6\times 10^{-6}$~pc, or $\sim 3\times 10^{13}$~cm.}
-The orbital period is
-\begin{equation}
-  P = 2\pi\left(\frac{GM}{a^3}\right)^{-1/2}\!\!\!\! = 1.72\, \mathrm{yr}\, a_{16}^{3/2} M_8^{-1/2}.
-\end{equation}
 
@@
 
-In a gas-rich galaxy merger, the BH binary loses its angular
-momentum by torquing the surrounding disk through spiral arms and
-expelling gas out of a region twice as large as the binary orbit
+In this Letter we study the morphology of a jet emitted by one of the
+binary members during the binary's final phase of evolution before
+coalescence, i.e., when the binary's dynamics is dominated by GW
+emission.  We show that in this case the jet has a peculiar helical
+chirping morphology that can be used to infer the binary parameters.
+
+\section{Binary parameters}
+
+We study a BH binary formed following a gas-rich merger of two
+galaxies as a result of the processes described above.  We consider
+black holes with masses $M_1$ and $M_2$ in a circular orbit of radius
+$a$ around each other.  Let $M=M_1+M_2$.  For simplicity, we consider
+an equal-mass binary ($M_1=M_2$) on a circular orbit. The orbital
+speed of each black hole is given by \citep{2010PhRvD..81d7503L}
+\begin{equation}
+  v_\mathrm{orb} = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2}\!\!\!\! = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
+  \label{eqn:vorb}
+\end{equation}
+where $a_{16}\equiv (a/10^{16}$ cm$)$ and $M_8\equiv (M/10^8$
+M$_\odot)$ are the binary separation and mass,
+respectively.\footnote{Note that $10^{16} \mathrm{cm}\approx 3.4\times
+  10^{-3} \mathrm{pc}$.  The Schwarzchild radius of a $10^8$~M$_\odot$
+  black hole is $9.6\times 10^{-6}$~pc, or $\sim 3\times 10^{13}$~cm.}
+The orbital period is
+\begin{equation}
+  P = 2\pi\left(\frac{GM}{a^3}\right)^{-1/2}\!\!\!\! = 1.72\, \mathrm{yr}\, a_{16}^{3/2} M_8^{-1/2}.
+\end{equation}
+
+In a gas-rich galaxy merger, the BH binary loses its angular momentum
+by torquing the surrounding disk through spiral arms and expelling gas
+out of a region twice as large as the binary orbit
 \citep{2005ApJ...622L..93M, 2007PASJ...59..427H, 2008ApJ...672...83M,
-  2009MNRAS.393.1423C}.  Fresh gas re-enters the region of the binary
-as a result of turbulent transport of angular momentum in the
-surrounding disk \citep{2008ApJ...672...83M, 2012ApJ...755...51N,
-  2012MNRAS.427.2680K, 2012A&A...545A.127R}.  Since the expelled gas
-carries a specific angular momentum of $\sim va$, the coalescence time
-of the binary is inversely proportional to the supply rate of fresh
-gas into the binary region.  In a steady state, this supply rate,
-$\dot M$ is proportional to the accretion rate of the circumbinary gas
-disk.  But only a fraction of gas that enters the central gap accretes
-onto the black hole and fuels quasar activity.  Therefore, it is
-appropriate to express $\dot M$ in Eddington units, $\dot{\mathcal
+  2009MNRAS.393.1423C}.  The expelled gas carries a specific angular
+momentum of $\sim va$.  Further angular momentum dissipation of the
+binary can happen due to its interaction with fresh gas that enters
+the hollowed out region as a result of tidal torques
+\citep{2008ApJ...672...83M, 2012ApJ...755...51N, 2012MNRAS.427.2680K,
+  2012A&A...545A.127R}.  Therefore, the coalescence time of the binary
+is inversely proportional to the rate at which fresh gas re-enters the
+binary region.  But only a fraction of gas that enters the central
+hollow region accretes onto the black hole and fuels quasar activity.
+Therefore, we express $\dot M$ in Eddington units, $\dot{\mathcal
   M}\equiv\dot M /\dot M_E$, where $\dot M_E$ is the accretion rate
@@
 \begin{equation}
-  v = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
+  v_\mathrm{orb} = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
   \label{eqn:v}
@@
 \end{equation}
-Note that the separation is dependent on the gas accretion rate.  This
-evolution of the binary separation and orbital velocity is shown in
-Figure~\ref{fig:evolve}.
+Note that the separation is dependent on the gas accretion rate.
 
@@
 \end{equation}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics*[width=\columnwidth]{evolve.pdf}
+    % gw.py (case 5) 
+  \end{center}
+  \caption{Evolution of the binary parameters.  As the binary
+    separation $a$ decreases, the orbital velocity $v_\mathrm{orb}$
+    and the opening angle $\psi$ of the precessing jet, increase.
+    This example describes the evolution of a binary with
+    $M=10^8$~M$_\odot$ and $a_0=8.3\times 10^{-3}$~pc, but can be
+    scaled to other masses and initial separations.}
+  \label{fig:evolve}
+\end{figure}
+
 When the binary separation reaches the point where gravitational
@@
 \begin{equation}
-  \dot\psi^\mathrm{obs}=\dot\psi = -\frac{1}{2\cot\psi}\frac{\dot a}{a}.
+ \dot\psi = -\frac{1}{2\cot\psi}\frac{\dot a}{a}.
 \end{equation}
@@
 \begin{equation}
-  \sin\psi = \sin\psi_0\left(\frac{a_0}{a}\right),
+  \sin\psi = \sin\psi_0\left(\frac{a_0}{a}\right)^{1/2},
 \end{equation}
@@
 which this happens reflects the binary's dynamical evolution.
-Figure~\ref{fig:evolve} shows the evolution of the half-opening angle
-in the gas and GW-dominated phase.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics*[width=\columnwidth]{evolve.pdf}
-    % gw.py (case 5) 
-  \end{center}
-  \caption{Evolution of the binary parameters.  As the binary
-    separation $a$ decreases, the orbital velocity $v_\mathrm{orb}$
-    and the opening angle $\psi$ of the precessing jet, increase.}
-  \label{fig:evolve}
-\end{figure}
+Figure~\ref{fig:evolve} shows the evolution of the binary separation,
+orbital velocity, and the half-opening angle for a binary in the
+GW-dominated phase.
 
@@
   \end{center}
-  \caption{Geometry of the jet.  The observer's sky is in the $yz$
-    plane, so that the line of sight is along the $x$-axis.  The jet
-    precesses around the $z^\prime$ axis with half-opening angle
-    $\psi$. The $y$ and $y^\prime$ axes overlap, and $z$, $x$, and
-    $z^\prime$ axes are co-planer.  The angle between the $z^\prime$
-    axis and the line of sight is $i$.  This geometry is valid in the
-    limit $a\ll z^\prime\psi$.}
+  \caption{Geometry of the jet \citep{1982ApJ...262..478G}.  The
+    observer's sky is in the $yz$ plane, so that the line of sight is
+    along the $x$-axis.  The jet precesses around the $z^\prime$ axis
+    with half-opening angle $\psi$. The $y$ and $y^\prime$ axes
+    overlap, and $z$, $x$, and $z^\prime$ axes are co-planer.  The
+    angle between the $z^\prime$ axis and the line of sight is $i$.
+    This geometry is valid in the limit $a\ll z^\prime\psi$.}
   \label{fig:jetgeom}
@@
 
-\section{Flagging GW binaries and probing GR}
-
-As the binary evolves, the combination of increasing orbital period
-and speed, and the apparant superluminal motion effect, results in a
-jet morphology with an increasing opening angle and increasingly rapid
-periodic variation, together with a twisted shape.
-Figure~\ref{fig:jet_mdot1.0} shows these three features for an
-equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  For
-concreteness, we have assumed $i=40^\circ$ and $\beta=0.9$.  The
-angular diameter distance to the binary is assumed to be $d =
-100$~Mpc, which corresponds to redshift $z\sim 0.3$.  The forward jet
-streches on the sky while the backward jet is compressed due to the
-superluminal motion.  The right panel of Figure~\ref{fig:jet_mdot1.0}
-zooms into the inner region of the forward jet and shows that the
-binary creates wiggles in the jet morphology of the order of a
-milli-arcsecond.  The duration of the binary evolution captured by
-Figure~\ref{fig:jet_mdot1.0} is 100 yr.  Although the system can be
-rescaled to different masses, times, and separation, the initial
-separation in Figure~\ref{fig:jet_mdot1.0} is $\sim 10^{-3}$~pc, when
-the binary enters it GW-dominated phase.
-
-The increasing orbital speed of the binary leads to an increase in the
-jet's precession angular velocity.  This increases the helicity of the
-jet such that the jet is wound up increasingly tightly closer to the
-source.  This gives rise to a ``chirping'' morphology as seen in
-Figure~\ref{fig:jet_mdot1.0}.  The chirp is easier to resolve in jets
-that are closer to the line of sight, where superluminal motion can
-magnify the forward jet.  A chirping black hole is an unambiguous
-signature of a rapidly evolving binary BH system; the chirp is direct
-reflection of the orbital velocity evolution in
-Figure~\ref{fig:evolve}.  Together, the three features of the jet
-morphology in Figure~\ref{fig:jet_mdot1.0}, namely, conical jet
-geometry with an increasing opening angle, a radio chirp geometry, and
-an asymmetrical shape, directly probe the dynamics of the binary in
-its GW phase. Observations of such a milliarcsecond geometry can thus
-constrain GR in the strong field limit.
-
 \begin{table}
@@
 
+\section{Flagging GW binaries and probing GR}
+
+As the binary evolves, the combination of increasing orbital speed and
+the apparant superluminal motion effect results in a jet morphology
+with an increasing opening angle and increasingly rapid periodic
+variation, together with a twisted shape.
+Figure~\ref{fig:jet_mdot1.0} shows these three features for an
+equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  For
+concreteness, we have assumed $i=40^\circ$ and $\beta=0.9$.  The
+angular diameter distance to the binary is assumed to be $d = 100$~Mpc
+(which corresponds to redshift $z\sim 0.025$).  The forward jet
+stretches on the sky while the backward jet is compressed due to the
+apparent superluminal motion.  The right panel of
+Figure~\ref{fig:jet_mdot1.0} zooms into the inner region of the
+forward jet and shows that the binary creates wiggles in the jet
+morphology of the order of a milli-arcsecond.  The duration of the
+binary evolution captured by Figure~\ref{fig:jet_mdot1.0} is 100 yr.
+Although the system can be rescaled to different masses, times, and
+separation, the initial separation in Figure~\ref{fig:jet_mdot1.0} is
+$\sim 10^{-3}$~pc, when the binary enters it GW-dominated phase.
+
+The increasing orbital speed of the binary leads to an increase in the
+jet's precession angular velocity.  This increases the helicity of the
+jet such that the jet is wound up increasingly tightly closer to the
+source.  This gives rise to a ``chirping'' morphology as seen in
+Figure~\ref{fig:jet_mdot1.0}.  The chirp is easier to resolve in jets
+that are closer to the line of sight, where superluminal motion can
+magnify the forward jet.  A chirping black hole is an unambiguous
+signature of a rapidly evolving binary BH system; the chirp is direct
+reflection of the orbital velocity evolution in
+Figure~\ref{fig:evolve}.  Together, the three features of the jet
+morphology in Figure~\ref{fig:jet_mdot1.0}, namely, conical jet
+geometry with an increasing opening angle, a radio chirp geometry, and
+an asymmetrical shape, directly probe the dynamics of the binary in
+its GW phase.
+
+Observations of such a milliarcsecond geometry can thus constrain GR
+in the strong field limit.  The relative separation of the jet winding
+directly constraints the ratio $a_{16}^4/M_8^3$ by constraining the
+evolution of the orbital velocity.  The jet opening angle evolution
+tracks the evolution of the binary separation.  We make the code for
+implementing the model presented in this Letter available at
+\url{https://github.com/gkulkarni/JetMorphology} under the MIT open
+source software license.  This code can be used to explore the
+dynamical, astrophysical, and geometrical parameter spaces of the
+model.
+
 \section{Conclusion}
@@
 Using a model of jet precession in a binary BH, we have shown that the
-jet morphology of a binary that is entering its GW-induced phase of
-inspiral exhibits a distinct chirping morphology. Jets in these
-binaries have a biconical morphology.  After the binary enters the GW
-phase, the opening angle of cone rapidly increases within a separation
-of tens of milli-arcseconds at a distance of 100~Mpc.  In addition,
-the periodicity of the jet's motion on the conical surface increases
-as the orbital period of the binary goes up in this stage of its
-evolution.  Finally, due to the apparent superluminal motion and the
-widening opening angle, the jet is twisted into a peculiar morphology.
-The jet geometry can be used to infer the set of dynamical and
-geometrical parameters, listed in Table~\ref{table:params}.
+jet produced by a member of a binary that is entering its GW-induced
+phase of inspiral exhibits a distinct chirping morphology.  Jets in
+these binaries have a biconical morphology.  After the binary enters
+the GW phase, the opening angle of cone rapidly increases within a
+separation of tens of milli-arcseconds at a distance of $\sim
+100$~Mpc, depending on the angle that the jet makes relative to the
+line of sight.  In addition, the periodicity of the jet's motion on
+the conical surface increases with the orbital period of the binary in
+this stage of its evolution.  Finally, due to the apparent
+superluminal motion and the widening opening angle, the jet is twisted
+into a peculiar morphology.  The jet geometry can be used to infer the
+set of dynamical and geometrical parameters listed in
+Table~\ref{table:params}.
 
@@
 
-We make the code for implementing the model presented in this Letter
-available at \url{https://github.com/gkulkarni/JetMorphology} under
-the MIT open source software license.  This code can be used to
-explore the parameter space of the model.
-
 \section*{Acknowledgments}

### 2015-07-23 Paper/scratch.tex
Add acknowledgements
Correct typos
Minor modifications to text

@@
 \end{table}
+
+ 0.0500, 0.1000, 0.2000, 0.2083, 0.2174, 0.2273, 0.2381, 0.2500, 0.2632, 0.2778, 0.2941, 0.3125, 0.3300, 0.3330
+
+We will consider
+unequal black holes masses and elliptical orbits below.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec
+volutpat felis sem, fermentum tincidunt dolor bibendum a. Sed sit amet
+justo ac tortor dignissim efficitur. Sed pellentesque, nisi sit amet
+dapibus pharetra, orci dolor dictum elit, dignissim ornare arcu mauris
+non nunc. Nulla non elementum nibh, at posuere felis. Donec a
+sollicitudin libero. Duis vulputate libero egestas, facilisis mauris
+in, interdum elit.

### 2015-07-23 jet.py
Update case 0

@@ -19,7 +19,7 @@ import sys
 from matplotlib import cm
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-i = 30.0 # degrees
+i = 40.0 # degrees
 i *= np.pi/180.0 # radians 
 beta = 0.9 # jet speed in units of c 
 d = 100.0 # Mpc; Angular distance between jet and observer
@@ -32,7 +32,7 @@ pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
 coeff = -2.56e5/(M*1.0e-8)**3 
 
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i, beta, Mdot)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i*180.0/np.pi, beta, Mdot)
 save_pdf = True
 
 # case = 1, 2 zooms into the central region of the jet 
@@ -56,7 +56,7 @@ def half_opening_angle_intrinsic(a_16):
     return angle #*180.0/np.pi # degrees 
 
 if case == 0: 
-    t = np.logspace(-1.0,1.0,10000000)
+    t = np.logspace(-2.0,2.0,10000000)
     output_filename += '_full'
 elif case == 1:
     t = np.logspace(-2.0,2.0,10000000)

### 2015-07-23 jet.py
Add jet code

@@ -1,0 +1,178 @@
+import matplotlib as mpl
+mpl.rcParams['text.usetex'] = True 
+mpl.rcParams['font.family'] = 'serif'
+mpl.rcParams['font.serif'] = 'cm'
+mpl.rcParams['font.size'] = '22'
+import matplotlib.pyplot as plt
+import numpy as np
+import math 
+import sys
+
+#psi = 20.0 # degrees
+i = 40.0 # degrees
+beta = 0.90 # jet speed in units of c 
+d = 100.0 # Mpc; Distance between jet and observer
+
+M = 1.0e10 # Msun; total mass of the equal-mass binary
+Mdot = 1.0 # Eddington units
+#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
+a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
+pcto_10to16cm = 0.003241
+a0 /= pcto_10to16cm # 1.0e16 cm
+coeff = -2.56e5/(M*1.0e-8)**3 
+
+d *= 1.0e3 # kpc
+
+#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
+output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
+save_pdf = True
+
+#psi *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # radians 
+
+c = 3.0e5 # km/s; speed of light
+yrbys = 3.154e7
+kpcbykm = 3.086e16
+
+def binary_separation_gw(t):
+    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
+    a = a**(1./4.)
+    return a
+
+def binary_orbital_period(a_16):
+    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
+    return t
+
+def half_opening_angle_intrinsic(a_16):
+    psi0 = 2.0*np.pi/180.0 # radians 
+    angle = np.arcsin(np.sin(psi0)*a0/a_16)
+    return angle #*180.0/np.pi # degrees 
+
+case = int(sys.argv[1])
+if case == 0: 
+    t = np.logspace(-2.0,2.0,10000000)
+    output_filename += '_full'
+elif case == 1:
[…]

### 2015-07-23 gw.py
Correct jet opening angle

@@ -62,7 +62,7 @@ def orbital_speed(a_16):
 v = orbital_speed(a) 
 
 def half_opening_angle_intrinsic(a_16):
-    psi = np.arcsin(np.sin(psi0)*a0/a_16)
+    psi = np.arcsin(np.sin(psi0)*np.sqrt(a0/a_16))
     return psi*180.0/np.pi
 
 psi0 = 2.0*np.pi/180.0 # radians 

### 2015-07-23 gw.py
Update documentation

@@ -1,7 +1,7 @@
 """
 File: gw.py 
 
-Evolution of a BH binary in the GW-dominated phase (Figure 3).
+Evolution of a BH binary in the GW-dominated phase (Case 5 gives Figure 3).
 
 """
 

### 2015-07-23 jet.py

@@ -1,3 +1,12 @@
+"""
+
+File: jet.py 
+
+Jet morphology from a BH binary that is in the GW-dominated phase of
+its inspiral.  Case 0 gives Figure 5 (left panel).
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'

### 2015-07-23 jet_image.py

@@ -1,9 +1,10 @@
 """
 
-File: jet.py 
+File: jet_image.py 
 
 Jet morphology from a BH binary that is in the GW-dominated phase of
-its inspiral (Figures 4). 
+its inspiral.  This code uses a jet brightness model to produce an
+image of the jet on the sky.  Case 1 gives Figure 5 (right panel).
 
 """
 

### 2015-07-23 jet.py
Remove redundant code

@@ -17,14 +17,12 @@ import numpy as np
 import math 
 import sys
 
-#psi = 20.0 # degrees
 i = 40.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
 
 M = 1.0e10 # Msun; total mass of the equal-mass binary
 Mdot = 1.0 # Eddington units
-#a0 = 3.241e-3*(M*1.0e-8)**(3./4.) # pc
 a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
 pcto_10to16cm = 0.003241
 a0 /= pcto_10to16cm # 1.0e16 cm
@@ -32,11 +30,9 @@ coeff = -2.56e5/(M*1.0e-8)**3
 
 d *= 1.0e3 # kpc
 
-#output_filename = 'jet_i%2d_psi%2d_beta%3.2f' % (i, psi, beta)
 output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f' % (i, beta, Mdot)
 save_pdf = True
 
-#psi *= np.pi/180.0 # radians 
 i *= np.pi/180.0 # radians 
 
 c = 3.0e5 # km/s; speed of light
@@ -62,7 +58,6 @@ if case == 0:
     t = np.logspace(-2.0,2.0,10000000)
     output_filename += '_full'
 elif case == 1:
-    #t = np.linspace(10.0,14.0,10000)
     t = np.logspace(-2.0,2.0,10000000)
     t = t[6000000:9500000]
     output_filename += '_zoom1'
@@ -75,7 +70,6 @@ t0 = t[0]
 def t_binary(time):
     t_merge=abs(coeff)*a0**4/4.0
     return np.abs(time-t_merge) # yr
-    #return np.abs(time-64001.477505390176) # yr
 
 def Omega(time):
     # Angular velocity times time for jet precession 
@@ -123,10 +117,8 @@ phi_z_obs *= 1.0e3 # mas
 
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
-#ax.set_ylim(-0.01,0.01)
 if case==1:
     ax.set_ylim(-50.0,50.0)
-    #ax.set_xlim(-0.01,0.01)
 ax.plot(phi_z_obs,phi_y_obs,c='k',lw=1,rasterized=True)
[…]

### 2015-12-02 Paper/paper.tex
Update documentation
Change figure filename
Resubmitted version

@@
 the binary following GW emission at coalescence
-\citep{2007APS..APR.S1010B}, flares from shocked remnants of accretion
-disk around the recoiled remnant black hole
+\citep{2007APS..APR.S1010B, 2008ApJ...676L...5L}, flares from shocked
+remnants of accretion disk around the recoiled remnant black hole
 \citep{2007PhRvL..99d1103L, 2008ApJ...682..758S, 2009CQGra..26i4032H},
@@
 black hole merger \citep{2008ApJ...672...83M, 2009CQGra..26i4032H,
-  2012MNRAS.427.2680K}, evolution in the profiles of broad emission
+  2012MNRAS.427.2680K, 2013MNRAS.436.2997D, 2014ApJ...783..134F,
+  2015ApJ...807..131S}, evolution in the profiles of broad emission
 lines shortly before merger \citep{2013MNRAS.432.1468M},
@@
 
-In this Letter we study the morphology of a jet emitted by one of the
-binary members during the binary's final phase of evolution before
+In this Letter, we study the morphology of jets from binary BHs whose
+dynamics is doiminated by GW emission.  Jets in binary BHs have been
+studied in a number of previous works \citep{1980ApJ...238L.129S,
+  1981ApJ...246L..65I, 1981ApJ...246L.141H, 1982ApJ...258L..63G,
+  1982ApJ...253L...1G, 1982ApJ...262..478G}.  It has been shown that
+the orbital motion of binary results in a precession of the jet,
+resulting in a helical or ``wiggly'' jet morphology.  Such models of
+jet precession have been used to understand the observed jet
+morphology in several AGN.  For example, recently
+\citet{2014MNRAS.445.1370K} fit a jet precession model to the observed
+jet in the quasar S5 1928$+$738 using 18 yr of very long baseline
+interferometric (VLBI) data.  This object was first studied using BH
+binary models by \citet{1993ApJ...409..130R}, but
+\citet{2014MNRAS.445.1370K} also included effects of BH spin.  Other
+objects studied in this manner are PG 1302-102
+\citep{2015MNRAS.454.1290K, 2015Natur.518...74G}, BL Lacertae
+\citep{2003MNRAS.341..405S, 2013MNRAS.428..280C}, NGC 4151
+\citep{2012ApJ...759..118B}, OJ 287 \citep{2012MNRAS.421.1861V}, S5
+1803$+$784 \citep{2008A&A...483..125R}, 3C 345
+\citep{2005A&A...431..831L}, 3C 210 \citep{2004MNRAS.349.1218C}, PKS
+420-014 \citep{2001A&A...374..784B}, 3C 273
+\citep{2000A&A...360...57R}, and Mrk 501 \citep{1999A&A...347...30V}.
+\citet{1982ApJ...262..478G} also studied several AGN using their model
+of jet precession.
+
+The shortest binary separations inferred from these studies are of the
+order of $10^{-2}$ pc.  At smaller separations of around $10^{-3}$ pc,
+the orbital velocity of the BHs typically becomes relativistic
+($v_\mathrm{orb}\sim 10^4$ km$/$s or $\beta\sim 0.01$).  It was
+thought that at these small separations, the binary evolves in a
+vacuum, because the inspiral time scale of the binary is likely
+shorter than the viscous time scale of any circumbinary gas disk.  As
+a result, gas fails to catch up with the BHs, making any accretion
+activity impossible in this stage of the binary's evolution
+\citep{2005ApJ...622L..93M, 2010PhRvD..81b4019S, 2010ApJ...714..404T}.
+However recent viscous hydrodynamical simulations do not find this
+decoupling between the BHs and the circumbinary disk.  In these
+simulations, the binary's member BHs can continue to accrete gas from
+the circumbinary accretion disk all the way until coalescence
+\citep{2012PhRvL.109v1102F, 2013MNRAS.436.2997D, 2014PhRvD..89f4060G,
+  2014PhRvD..90j4030G, 2015MNRAS.446L..36F, 2015ApJ...807..131S}.
+This opens up the possibility of jet production by one of the binary
+members during the binary's final phase of evolution before
 coalescence, i.e., when the binary's dynamics is dominated by GW
-emission.  We show that in this case the jet has a peculiar helical
-chirping morphology that can be used to infer the binary parameters.
+emission.  In this Letter we study the morphology of such jets.
+
+All of the above models of jet precession in binary BHs consider
+stable binaries, in which the time scale of evolution of the binary
+separation is much larger than the orbital period of the binary.  For
+such binaries, the helical morphology of the jet is does not change
+significantly between precession cycles.  However, we will see below
+that for a binary BH with separation $a$, the orbital period is
+proportional to $a^{1.5}$, while the time scale of evolution of the
+binary separation is proportional to $a^{4}$.  Thus, as the binary
+shrinks, the latter time scale decreases much more rapidly than the
+former, and for small enough separations, the two time scales can be
+comparable.  In this paper, we study jets in this previously ignored
+regime.  We show that at this stage of the binary's evolution, the jet
+morphology is more complicated: the jet now has a peculiar helical
+chirping morphology.  We further argue that in this case the jet
+morphology can be used to infer the binary parameters and show how
+such features can be distinguished from the usual helical jet
+morphology.
 
@@
 an equal-mass binary ($M_1=M_2$) on a circular orbit. The orbital
-speed of each black hole is given by \citep{2010PhRvD..81d7503L}
+speed of each black hole is given by 
 \begin{equation}
@@
   P = 2\pi\left(\frac{GM}{a^3}\right)^{-1/2}\!\!\!\! = 1.72\, \mathrm{yr}\, a_{16}^{3/2} M_8^{-1/2}.
+  \label{eqn:period}
 \end{equation}
@@
 out of a region twice as large as the binary orbit
-\citep{2005ApJ...622L..93M, 2007PASJ...59..427H, 2008ApJ...672...83M,
-  2009MNRAS.393.1423C}.  The expelled gas carries a specific angular
-momentum of $\sim va$.  Further angular momentum dissipation of the
-binary can happen due to its interaction with fresh gas that enters
-the hollowed out region as a result of tidal torques
+\citep{1994ApJ...421..651A, 2005ApJ...622L..93M, 2007PASJ...59..427H,
+  2008ApJ...672...83M, 2009MNRAS.393.1423C}.  The expelled gas carries
+a specific angular momentum of $\sim va$.  Further angular momentum
+dissipation of the binary can happen due to its interaction with fresh
+gas that enters the hollowed out region as a result of tidal torques
 \citep{2008ApJ...672...83M, 2012ApJ...755...51N, 2012MNRAS.427.2680K,
@@
 efficiency of 10\%, $\dot M_E=2.3 M_\odot\, \mathrm{yr}^{-1}M_8$.  We
-therefore have \citep{2010PhRvD..81d7503L}
+therefore have \citep{2010PhRvD..81d7503L, 2015MNRAS.448.3603D}
 \begin{equation}
@@
   \frac{a}{\dot a} = t_\mathrm{GW} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
+  \label{eqn:shrink}
 \end{equation}
@@
   \end{center}
-  \caption{The orbital velocity $\mathbf{v}_\mathrm{orbital}$
-    introduces the jet precision angle $\psi$.  Without the orbital
+  \caption{The orbital velocity $\mathbf{v}_\mathrm{orb}$
+    introduces the jet precession angle $\psi$.  Without the orbital
     motion, $\psi=0$, and the jet traces a cylindrical surface.  With
@@
 \end{equation}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics*[width=\columnwidth]{evolve.pdf}
-    % gw.py (case 5) 
-  \end{center}
-  \caption{Evolution of the binary parameters.  As the binary
-    separation $a$ decreases, the orbital velocity $v_\mathrm{orb}$
-    and the opening angle $\psi$ of the precessing jet, increase.
-    This example describes the evolution of a binary with
-    $M=10^8$~M$_\odot$ and $a_0=8.3\times 10^{-3}$~pc, but can be
-    scaled to other masses and initial separations.}
-  \label{fig:evolve}
-\end{figure}
+(Note that this relation is valid in the $v_\mathrm{orb}\ll
+v_\mathrm{jet}\sim c$. Also, we are ignoring any relativistic
+corrections to this equation.)
 
@@
   \dot\psi = -\frac{1}{2\cot\psi}\frac{\dot a}{a}.
+  \label{eqn:widen}
 \end{equation}
@@
 which this happens reflects the binary's dynamical evolution.
-Figure~\ref{fig:evolve} shows the evolution of the binary separation,
-orbital velocity, and the half-opening angle for a binary in the
-GW-dominated phase.
 
@@
   \begin{tabular}{cc}
-    % jet3d_check2.py (case 0) 
+    % jet.py (case 0) 
     \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
-    % jet3d_image_fulljet.py (case 1) 
-    \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
+    % jet_image.py (case 1) 
+    \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_image_zoom1}.pdf}
   \end{tabular}
@@
 \caption{The left panel shows forward and backward jets for an
-  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
-  assumed $i=40^\circ$, $\beta=0.9$, and angular diameter distance $d
-  = 100$~Mpc.  The jets are twisted because of the apparant
-  superluminal motion.  Also evident is the opening of the jet and the
-  radio jet chirp.  Right panel shows normalised brightness of the
-  inner 30~mas region of the forward jet.}
+  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  (This
+  figure shows the position of many jet particles observed at a fixed
+  time on Earth.)  We have assumed $i=40^\circ$, $\beta=0.9$, and
+  angular diameter distance $d = 100$~Mpc.  The jets are twisted
+  because of the apparant superluminal motion.  Also evident is the
+  opening of the jet and the radio jet chirp.  Right panel shows
+  normalised brightness of the inner 30~mas region of the forward
+  jet. (This panel shows the intensity of many particles observed at a
+  fixed time on Earth.)}
 \label{fig:jet_mdot1.0}
@@
 \begin{equation}
-  S_r\propto t^{-\delta}.
+  S_r\propto t_\mathrm{proper}^{-\delta},
 \end{equation}
+where $t_\mathrm{proper}$ is the proper time in the frame of the jet.
 We assume $\delta=1$ and $\alpha=1$, hereafter
@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{c l}
-\hline 
-Parameter & Description \\ 
-\hline 
-$M$ & Total mass \\
-$\zeta$ & Mass asymmetry \\
-$\chi$ & Intrinsic jet direction \\
-$e$ & Orbital eccentricity \\
-$\alpha$ & SED slope \\
-$\delta$ & Time evolution of brightness evolution \\
-$\beta$ & Jet speed \\
-$i$ & Orbital inclination \\ 
-$\dot{\mathcal M}$ & Gas accretion rate\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Parameters of the problem.}
-\label{table:params}
-\end{table}
-
-\section{Flagging GW binaries and probing GR}
+\section{Jet Morphology}
 
@@
 signature of a rapidly evolving binary BH system; the chirp is direct
-reflection of the orbital velocity evolution in
-Figure~\ref{fig:evolve}.  Together, the three features of the jet
-morphology in Figure~\ref{fig:jet_mdot1.0}, namely, conical jet
-geometry with an increasing opening angle, a radio chirp geometry, and
-an asymmetrical shape, directly probe the dynamics of the binary in
-its GW phase.
+reflection of the orbital velocity evolution of the binary. Together,
+the three features of the jet morphology in
+Figure~\ref{fig:jet_mdot1.0} are:
+\begin{itemize}
+\item Increase in the helicity of the jet towards the central source (``chirpyness''),
+\item Widening of the opening angle towards the central source, and 
+\item An apparant variation in the jet precession axis towards the central source (``twist''). 
+\end{itemize}
+These directly probe the dynamics of the binary in its GW phase.
 
-Observations of such a milliarcsecond geometry can thus constrain GR
-in the strong field limit.  The relative separation of the jet winding
-directly constraints the ratio $a_{16}^4/M_8^3$ by constraining the
-evolution of the orbital velocity.  The jet opening angle evolution
-tracks the evolution of the binary separation.  We make the code for
-implementing the model presented in this Letter available at
-\url{https://github.com/gkulkarni/JetMorphology} under the MIT open
-source software license.  This code can be used to explore the
-dynamical, astrophysical, and geometrical parameter spaces of the
-model.
+\subsection{Can chirping jets be observed?}
+
+Observations of such a milliarc second scale geometry with the three
+features described above can thus flag a BH binary nearing the end of
+its merger.  We now consider the question of whether such morphologies
+will be measurable in real observations.
+
+As discussed in Section~\ref{sec:intro}, several helical jets have
+been observed thanks to VLBI.  A resolution of few tens of microarc
+seconds can be achieved using the Very Long Baseline Array (VLBA) with
+sensitivity of up to 30 $\mu$Jy$/$beam, as demonstrated by the MOJAVE
+survey \citep{2009AJ....138.1874L}.  VLA has angular resolution of few
+tens of milliarcsecconds with a sensitivity of about 100
+$\mu$Jy$/$beam.  The European VLBI network (EVN) which can resolve up
+to 0.13 milliarc seconds and can detect surface brightness down to 19
+$\mu$Jy$/$beam.  The Very Large Array (VLA) has an angular resolution
+of few tens of milliarcsecconds with a sensitivity of about 100
+$\mu$Jy$/$beam.  The future Square Kilometer Array (SKA) on its own
+will have 2 milliarc second resolution at around 10 GHz, but SKA will
+be an order of magnitude more sensitive than current instruments with
+a sensitivity of about 50 nJy$/$beam.  Thus angular resolution of tens
+to hundreds of microarc seconds and flux sensitivity of a few
+$\mu$Jy$/$beam is achievable with current VLBI instruments.
+
+The brighter part of the jet (intensity within factor of 10 from the
+maximum) in Figure~\ref{fig:jet_mdot1.0} (right panel) has separations
+of $\sim 1$ mas.  It should therefore be possible to resolve the jet
+at an angular resolution of $\sim 2$ mas.  (In the next section, we
+discuss an algorithm to fit the jet model to observational data.)  If
+it is aligned closer to the line of sight, the forward jet is
+magnified due to apparant superluminal motion.  This will make it
+possible to resolve jets of objects that are farther away than the
+example in Figure~\ref{fig:jet_mdot1.0}.  Binaries with higher masses
+have large orbital periods that will result is a helix with wider
+seperation between its cycles.  These objects would therefore be
+easier to detect.  We have not modelled the absolute surface
+brightness of jets, only relative variation in the jet brightness (as
+shown in Figure~\ref{fig:jet_mdot1.0}), but due to the $(1+z)^{-4}$
+dimming, it will be exceedingly difficult to detect such jets at
+higher redshifts.
+
+We note here that there is at least one object in the literature that
+shows a jet morphology similar to Figure~\ref{fig:jet_mdot1.0}.  The
+quasar 4C 18.68 described by \citet{1982ApJ...253L...1G} shows a
+helical jet with an apparant opening angle that increases towards the
+central source from 11$^\circ$ to 45$^\circ$ in just two cycles
+\citet{1982ApJ...262..478G}.  The observing beamwidth of about 400 mas
+makes it difficult to resolve a chirp in this source, but it is a
+prototype of the morphology that we are discussing here.
+\citet{2015MNRAS.453.1562G} have reported that 111 candidates show
+periodic variability in brightness in the Cataline Real-time Transient
+Survey, indicating presence of binary BHs.  An analysis of these
+candidates suggests the presence of circumbinary gas at small orbital
+distances \citep{2015MNRAS.453.1562G}.  This bodes well for the
+likelihood of the presence of chirpy jets.  In low resolution wide
+field survey, detecting modulations in the lightcurves would be an
+easier diagnostic of a binary BH \citep{2011ApJ...743..136O,
+  2011ApJ...734L..37K}.
+
+\subsection{Deriving binary parameters from observations of chirping jets}
+
+From observations of chirping jets, we can infer the dynamical
+parameters of the binary BH by first fitting a purely geometric model
+to the observed jet morphology.  To create a model of this kind, we
+generalise the simple helical model considered by
+\citet{2014MNRAS.445.1370K} and \citet{2001NewA....6...61N} to write
+\begin{align}
+  z &= x^{-\alpha} \frac{B}{2\pi} u \cos{u}, \\
+  y &= x^{-\alpha} \frac{B}{2\pi} u \sin{u}, \\
+  x &= \frac{A}{2\pi} u \left(\frac{u}{2\pi}\right)^\beta.
+\end{align}
+Here $u$ is the azimuthal angle around the jet axis, $x$ and $y$ axes
+are in the plane of the sky, and the $z$ axis points to the observer.
+The $x^{-\alpha}$ term in the first two equations captures the
+widening opening angle of the jet, while the $(u/2\pi)^{\beta}$ term
+in the last equation parameterises the chirp.  This model can be fit
+to an observed jet on the sky by using two rotation matrices,
+$\mathcal{R}_\lambda$ and $\mathcal{R}_i$, which rotate the model
+in the above equtions to match with the jet precession axis and the
+jet symmetry axis.  These matrices are given by
+
+\begin{flalign}
+\mathrm{R}_{i} \!=\!
+\begin{pmatrix}
+\cos i & 0 & \sin i \\ 
+0 & 1 & 0 \\ 
+-\sin i & 0 & \cos i
+\end{pmatrix}
+\label{riota}
+\end{flalign}
+and
+\begin{flalign}
+\mathcal{R}_{\lambda}\!=\!
+\begin{pmatrix}
+\cos \lambda{_0} & -\sin \lambda{_0} & 0 \\ 
+\sin \lambda{_0} & \cos \lambda{_0} & 0 \\ 
+0 & 0 & 1%
+\end{pmatrix}.
+\label{rlambda}
+\end{flalign}
+Further, we can parameterise the ``twist'' in the jet by requiring
+that $i$ and $\lambda$ vary with $x$.  Without loss of generality,
+we take this variation in $i$ and parameterise it as
+\begin{equation}
+  i = \gamma x + i_0.
+\end{equation}
+We thus arrive at a model with 7 parameters ($\alpha$, $\beta$,
+$\gamma$, $A$, $B$, $i_0$, $\lambda_0$), which can be fit to any
+observed jet, such as that in Figure~\ref{fig:jet_mdot1.0}.  Note that
+setting $\alpha$ and $\beta$ equal to $0$ yields the usual helical jet
+morphology, without the features discussed in this paper.  Thus, a
+non-zero value of these two parameters distinguishes chirpy jets from
+non-chirpy ones.
+
+The best fit geometric parameters can then be used to derive the
+dynamical parameters of the BH binary in the following way.  The
+parameters $\alpha$ and $B$ quantify the widening of the opening angle
+$\psi$.  As each cycle of the observed helical jet corresponds to a
+period $P(t)$ of the binary, we can then write $d\psi/dP$, which
+yields a constraint on the quantity $a^4/M^3$ as a function of $P$ via
+Equations (\ref{eqn:period}), (\ref{eqn:shrink}) and
+(\ref{eqn:widen}).  The parameters $A$ and $\beta$, which quantify the
+chirp, can be used to directly infer the evolution of the binary
+separation as $da/dP$, which also contraints the quantity $a^4/M^3$
+via Equation~(\ref{eqn:shrink}). These two constraints can be used to
+determine $a(P)$ and $M$ separately.  Finally, the parameters
+$i_0$, $\lambda_0$, and $\gamma$ quantify the twist in the jet,
+which can potentially be used to constrain the quantity
+$\dot{\mathcal{M}}$ by using Equation~(\ref{eqn:v}).
+
+Further parameters can be introduced in the dynamical model by
+considering unequal-mass binaries, eccentric orbits, and
+Post-Newtonian corrections, but we not develop this further in this
+Letter.  We make the code for implementing the model presented in this
+Letter and for fitting it to observed jets using MCMC analyses
+available at \url{https://github.com/gkulkarni/JetMorphology} under
+the MIT open source software license.  This code can be used to
+explore parameter space of the model.
 
@@
 into a peculiar morphology.  The jet geometry can be used to infer the
-set of dynamical and geometrical parameters listed in
-Table~\ref{table:params}.
+set of dynamical and geometrical parameters.
 
@@
 
-We acknowledge useful discussions with Enrico Ramirez-Ruiz and Alberto
-Rorai.  This work was supported in part by NSF grant AST-1312034 (for
-A.~L.).
+We acknowledge useful discussions with Martin Haehnelt, Enrico
+Ramirez-Ruiz and Alberto Rorai.  This work was supported in part by
+NSF grant AST-1312034 (for A.~L.).  G.~K. gratefully acknowledges
+support by he FP7 ERC Advanced Grant Emergence-320596.
 

### 2015-12-02 3dplot.py
Resubmitted version

@@ -1,0 +1,19 @@
+import matplotlib as mpl
+from mpl_toolkits.mplot3d import Axes3D
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+fig = plt.figure()
+ax = fig.gca(projection='3d')
+
+# theta = np.linspace(-4 * np.pi, 4 * np.pi, 100)
+# z = np.linspace(-2, 2, 100)
+# r = z**2 + 1
+# x = r * np.sin(theta)
+# y = r * np.cos(theta)
+
+ax.plot(xv, yv, zv)
+
+plt.show()
+

### 2015-12-02 I.py

@@ -1,16 +1,12 @@
 import numpy as np 
 import matplotlib.pyplot as plt
 from matplotlib import cm
-from scipy.ndimage.filters import gaussian_filter as gf
+from scipy.ndimage.filters import gaussian_filter as gfx
 
-# a = 12.0
-# b = 0.5
-# i0 = 1.5
-# l0 = 0.0
-
-a, b, i0, l0 = 1.18611945e+01, 4.80667425e-01, 1.56853292e+00, 5.17469193e-05 
-
-#a, b, i0, l0 = 1.19996440e+01, 5.00329139e-01, 1.49997638e+00, 2.47767666e-04
+a = 12.0
+b = 0.5
+i0 = 1.5
+l0 = 0.0
 phi = 1.0
 
 def F(u):
@@ -31,7 +27,7 @@ u = np.linspace(0.0, 20.0*np.pi, 1000)
 xv = x(u)
 yv = y(u)
 
-nc = 100
+nc = 1000
 a = np.zeros((nc,nc),dtype=np.float32) 
 zl = xv.min() - 5.0
 zu = xv.max() + 5.0
@@ -54,8 +50,8 @@ for i in xrange(xv.size):
     # a[ypos, zpos] += intensity[i]
     a[ypos, zpos] += 1.0
 
-a2 = gf(a, 1.0)
-#a2.tofile('mockdata2.dat')
+a2 = gf(a, 10.0)
+a2.tofile('mockdata.dat')
 #plt.contour(a2,200,colors='k',linestyles='solid')
 
 def m(i,j):
@@ -63,6 +59,8 @@ def m(i,j):
 
 plt.imshow(a2, cmap=cm.Blues) 
 plt.show()
+
+
 
 # fig = plt.figure(figsize=(7, 7))
 # ax = fig.add_subplot(1, 1, 1)

### 2015-12-02 fitjet.py

@@ -1,0 +1,144 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import scipy.optimize as op
+import emcee
+import triangle
+
+a2 = np.fromfile('mockdata2.dat',dtype=np.float32)
+
+def I(theta, i, j):
+
+    a, b, i0, l0 = theta
+    phi = 1.0
+    u = np.linspace(0.0, 20.0*np.pi, 1000)
+
+    def F(u):
+        return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
+                (a/(2.0*np.pi)) * u * np.sin(i0))
+
+    def G(u):
+        return (b/(2.0*np.pi)) * u * np.sin(u-phi)
+
+    def x(u):
+        return F(u)*np.cos(l0) - G(u)*np.sin(l0)
+
+    def y(u):
+        return F(u)*np.sin(l0) + G(u)*np.cos(l0)
+    
+    xv = x(u)
+    yv = y(u)
+
+    nc = 100
+    a = np.zeros((nc,nc),dtype=np.float32) 
+    zl = xv.min() - 5.0
+    zu = xv.max() + 5.0
+    yl = yv.min() - 5.0
+    yu = yv.max() + 5.0 
+    lz = zu - zl 
+    ly = yu - yl 
+    dy = ly/nc
+    dz = lz/nc
+
+    def zloc(x):
+        return int((x-zl)/dz) + 1 
+
+    def yloc(x):
+        return int((x-yl)/dy) + 1 
+
+    for c in xrange(xv.size):
+        zpos = zloc(xv[c]) 
+        ypos = yloc(yv[c])
+        a[ypos, zpos] += 1.0
+
+    return a[i,j]
+
+def neglnlike(theta, i, j, intensity, intensity_err):
+    model = I(theta, i, j)  
+    inv_sigma2 = 1.0/intensity_err**2 
+    return 0.5*(np.sum((intensity-model)**2*inv_sigma2 - np.log(inv_sigma2)))
+
+a2_err = np.zeros_like(a2)
+a2_err += 0.1
+
+theta_guess = (12.0, 0.5, 1.5, 0.0)
+
+iarr = np.arange(100)
+jarr = np.arange(100) 
+
+iarr = np.repeat(iarr, 100)
+jarr = np.tile(jarr, 100)
+
+
[…]

### 2015-12-02 fitjet2.py

@@ -1,0 +1,139 @@
+""" Uses image subtraction; otherwise same as fitjet.py """ 
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import scipy.optimize as op
+import emcee
+import triangle
+
+a2 = np.fromfile('mockdata2.dat',dtype=np.float32)
+
+def I(theta):
+
+    a, b, i0, l0 = theta
+    phi = 1.0
+    u = np.linspace(0.0, 20.0*np.pi, 1000)
+
+    def F(u):
+        return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
+                (a/(2.0*np.pi)) * u * np.sin(i0))
+
+    def G(u):
+        return (b/(2.0*np.pi)) * u * np.sin(u-phi)
+
+    def x(u):
+        return F(u)*np.cos(l0) - G(u)*np.sin(l0)
+
+    def y(u):
+        return F(u)*np.sin(l0) + G(u)*np.cos(l0)
+    
+    xv = x(u)
+    yv = y(u)
+
+    nc = 100
+    a = np.zeros((nc,nc),dtype=np.float32) 
+    zl = xv.min() - 5.0
+    zu = xv.max() + 5.0
+    yl = yv.min() - 5.0
+    yu = yv.max() + 5.0 
+    lz = zu - zl 
+    ly = yu - yl 
+    dy = ly/nc
+    dz = lz/nc
+
+    def zloc(x):
+        return int((x-zl)/dz) + 1 
+
+    def yloc(x):
+        return int((x-yl)/dy) + 1 
+
+    for c in xrange(xv.size):
+        zpos = zloc(xv[c]) 
+        ypos = yloc(yv[c])
+        a[ypos, zpos] += 1.0
+
+    return a.flatten() 
+
+def neglnlike(theta, intensity, intensity_err):
+    model = I(theta)  
+    inv_sigma2 = 1.0/intensity_err**2 
+    return 0.5*(np.sum((intensity-model)**2*inv_sigma2 - np.log(inv_sigma2)))
+
+a2_err = np.zeros_like(a2)
+a2_err += 0.1
+
+theta_guess = (12.0, 0.5, 1.5, 0.0)
+result = op.minimize(neglnlike, theta_guess, args=(a2, a2_err), method='Nelder-Mead') 
+
+print result.x
+print result.success
+
[…]

### 2015-12-02 fitjet_3d.py

@@ -1,0 +1,186 @@
+""" Uses image subtraction; otherwise same as fitjet.py """ 
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import scipy.optimize as op
+import emcee
+import triangle
+import sys
+
+a2 = np.fromfile('mockdata_3d_nc100.dat',dtype=np.float32)
+
+def I(theta): 
+
+    a, b, i, l, alpha, beta, gamma = theta 
+    
+    u = np.linspace(0.0, 20.0*np.pi, 1000)
+
+    def z(u):
+        return (a/(2.0*np.pi)) * u * (u/(2.0*np.pi))**beta
+
+    zv = z(u)
+
+    def x(u):
+        return (z(u)**-alpha) * (b/(2.0*np.pi)) * u * np.cos(u)
+
+    def y(u):
+        return (z(u)**-alpha) * (b/(2.0*np.pi)) * u * np.sin(u)
+
+    xv = x(u)
+    yv = y(u)
+
+    def ri(i):
+        return np.matrix([[np.cos(i), 0.0, np.sin(i)],[0.0, 1.0, 0.0],[-np.sin(i), 0.0, np.cos(i)]])
+
+    def rl(l):
+        return np.matrix([[np.cos(l), -np.sin(l), 0.0],[np.sin(l), np.cos(l), 0.0],[0.0, 0.0, 1.0]])
+
+    zvarr = zv*gamma
+    iarr = zvarr/zvarr.max()
+    iarr *= np.pi/2.0 
+    c = np.dstack((xv, yv, zv))
+    c = np.squeeze(c)
+    d = np.zeros((1000,3))
+    lm = rl(l) 
+    for n in range(1000):
+        d[n] = c[n]*ri(iarr[n])*lm 
+
+    xv = d[:,0]
+    yv = d[:,1]
+
+    xv = xv[~np.isnan(xv)]
+    yv = yv[~np.isnan(yv)]
+
+    nc = 100                       
+    a = np.zeros((nc,nc),dtype=np.float32) 
+    zl = xv.min() - 5.0
+    zu = xv.max() + 5.0
+    yl = yv.min() - 5.0
+    yu = yv.max() + 5.0 
+    lz = zu - zl 
+    ly = yu - yl 
+    dz = lz/nc 
+    dy = -ly/nc # Because "y" coordinate increases in opposite direction to "y" array index of a (or a2).
+
+    def zloc(cood):
[…]

### 2015-12-02 jet3d.py

@@ -1,0 +1,88 @@
+import matplotlib as mpl
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import cm
+from scipy.ndimage.filters import gaussian_filter as gf
+
+a = 0.1
+b = 10.0
+
+r = np.linspace(0.0,1.0,1000) 
+
+def z(u):
+    return (a/(2.0*np.pi)) * u * (u/(2.0*np.pi))**2
+
+u = np.linspace(0.0, 20.0*np.pi, 1000)
+zv = z(u)
+
+def x(u):
+    return (u**-0.2) * (b/(2.0*np.pi)) * u * np.cos(u)
+
+def y(u):
+    return (u**-0.2) * (b/(2.0*np.pi)) * u * np.sin(u)
+
+xv = x(u)
+yv = y(u)
+
+def ri(i):
+    return np.matrix([[np.cos(i), 0.0, np.sin(i)],[0.0, 1.0, 0.0],[-np.sin(i), 0.0, np.cos(i)]])
+
+def rl(l):
+    return np.matrix([[np.cos(l), -np.sin(l), 0.0],[np.sin(l), np.cos(l), 0.0],[0.0, 0.0, 1.0]])
+
+i = 2.0
+l = 3.0
+
+zvarr = zv*0.5 
+iarr = zvarr/zvarr.max()
+iarr *= np.pi/2.0 
+
+c = np.dstack((xv, yv, zv))
+# d = c*ri(i)*rl(l)
+
+# x = np.squeeze(np.asarray(d[:,0]))
+# y = np.squeeze(np.asarray(d[:,1]))
+
+c = np.squeeze(c)
+d = np.zeros((1000,3))
+lm = rl(l) 
+for n in range(1000):
+    d[n] = c[n]*ri(iarr[n])*lm 
+
+xv = d[1:,0]
+yv = d[1:,1]
+    
+plt.plot(xv, yv)
+plt.show()
+
+sys.exit()
+
+nc = 100                       
+a = np.zeros((nc,nc),dtype=np.float32) 
+zl = xv.min() - 5.0
+zu = xv.max() + 5.0
+yl = yv.min() - 5.0
+yu = yv.max() + 5.0 
+lz = zu - zl 
+ly = yu - yl 
+dz = lz/nc 
+dy = -ly/nc # Because "y" coordinate increases in opposite direction to "y" array index of a (or a2).
+
+def zloc(cood):
[…]

### 2015-12-02 jet_image.py

@@ -91,6 +91,7 @@ sign = 1 # forward jet
 norm_sep = False
 
 velx, vely, velz = vel(t)
+v = np.sqrt(velx**2 + vely**2 + velz**2)
 
 y = vely*t*yrbys/kpcbykm # kpc
 z = velz*t*yrbys/kpcbykm # kpc
@@ -105,7 +106,8 @@ alpha = 1.0
 delta = 1.0
 t_observed = t/(1.0-velx/c)
 
-gamma = 1.0/np.sqrt(1.0-beta**2)
+# gamma = 1.0/np.sqrt(1.0-beta**2)
+gamma = 1.0/np.sqrt(1.0-v**2/c**2)
 doppler = gamma*(1.0-beta)
 
 intensity = doppler**(3.0+alpha-delta) * (t_observed**(-delta))
@@ -117,11 +119,12 @@ if norm_sep:
     intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
 
 phi_y_obs *= 1.0e3 # mas
-phi_z_obs *= 1.0e3 # mas 
+phi_z_obs *= 1.0e3 # mas
 
 sign = -1 # backward jet 
 
 velx, vely, velz = vel(t)
+v = np.sqrt(velx**2 + vely**2 + velz**2)
 
 yb = vely*t*yrbys/kpcbykm # kpc
 zb = velz*t*yrbys/kpcbykm # kpc
@@ -139,7 +142,8 @@ alpha = 1.0
 delta = 1.0
 t_observed = t/(1.0-velx/c)
 
-gamma = 1.0/np.sqrt(1.0-beta**2)
+#gamma = 1.0/np.sqrt(1.0-beta**2)
+gamma = 1.0/np.sqrt(1.0-v**2/c**2)
 doppler = gamma*(1.0-beta)
 
 intensity_b = doppler**(3.0+alpha-delta) * (t_observed**(-delta))

### 2015-12-02 jet_image_mock.py

@@ -1,3 +1,13 @@
+"""
+
+File: jet_image.py 
+
+Jet morphology from a BH binary that is in the GW-dominated phase of
+its inspiral.  This code uses a jet brightness model to produce an
+image of the jet on the sky.  Case 1 gives Figure 5 (right panel).
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -82,6 +92,7 @@ sign = 1 # forward jet
 norm_sep = False
 
 velx, vely, velz = vel(t)
+v = np.sqrt(velx**2 + vely**2 + velz**2)
 
 y = vely*t*yrbys/kpcbykm # kpc
 z = velz*t*yrbys/kpcbykm # kpc
@@ -96,7 +107,8 @@ alpha = 1.0
 delta = 1.0
 t_observed = t/(1.0-velx/c)
 
-gamma = 1.0/np.sqrt(1.0-beta**2)
+# gamma = 1.0/np.sqrt(1.0-beta**2)
+gamma = 1.0/np.sqrt(1.0-v**2/c**2)
 doppler = gamma*(1.0-beta)
 
 intensity = doppler**(3.0+alpha-delta) * (t_observed**(-delta))
@@ -108,11 +120,12 @@ if norm_sep:
     intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
 
 phi_y_obs *= 1.0e3 # mas
-phi_z_obs *= 1.0e3 # mas 
+phi_z_obs *= 1.0e3 # mas
 
 sign = -1 # backward jet 
 
 velx, vely, velz = vel(t)
+v = np.sqrt(velx**2 + vely**2 + velz**2)
 
 yb = vely*t*yrbys/kpcbykm # kpc
 zb = velz*t*yrbys/kpcbykm # kpc
@@ -130,7 +143,8 @@ alpha = 1.0
 delta = 1.0
 t_observed = t/(1.0-velx/c)
 
-gamma = 1.0/np.sqrt(1.0-beta**2)
+#gamma = 1.0/np.sqrt(1.0-beta**2)
+gamma = 1.0/np.sqrt(1.0-v**2/c**2)
 doppler = gamma*(1.0-beta)
 
 intensity_b = doppler**(3.0+alpha-delta) * (t_observed**(-delta))
@@ -149,54 +163,44 @@ max_intensity = intensity.max()
 intensity /= max_intensity
[…]

### 2015-12-02 jet_image_try.py

@@ -1,3 +1,13 @@
+"""
+
+File: jet_image.py 
+
+Jet morphology from a BH binary that is in the GW-dominated phase of
+its inspiral.  This code uses a jet brightness model to produce an
+image of the jet on the sky.  Case 1 gives Figure 5 (right panel).
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -9,7 +19,6 @@ import math
 import sys
 from matplotlib import cm
 from mpl_toolkits.axes_grid1 import make_axes_locatable
-from scipy.ndimage.filters import gaussian_filter as gf
 
 i = 40.0 # degrees
 i *= np.pi/180.0 # radians 
@@ -149,53 +158,7 @@ max_intensity = intensity.max()
 intensity /= max_intensity
 intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
 
-nc = 1000
-a = np.zeros((nc,nc),dtype=np.float32) 
-zl = phi_z.min()-5.0
-zu = phi_z.max()+5.0
-yl = phi_y.min()-5.0
-yu = phi_y.max()+5.0
-
-print zl, zu
-print yl, yu 
-
-lz = zu - zl 
-ly = yu - yl 
-dy = ly/nc
-dz = lz/nc
-print dy, dz
-def zloc(x):
-    return int((x-zl)/dz) + 1 
-
-def yloc(x):
-    return int((x-yl)/dy) + 1 
-
-for i in xrange(phi_z.size):
-    zpos = zloc(phi_z[i]) 
-    ypos = yloc(phi_y[i])
-    a[ypos, zpos] += abs(intensity[i])*100000.0
-
-fig = plt.figure(figsize=(5, 5), dpi=100)
-a2 = gf(a, 10.0)
-#a2 = a
-plt.contour(a2,100,colors='k',linestyles='solid')
-
-ylabels = range(int(-60),int(70),20)
-ylocs = [yloc(x) for x in ylabels]
-ylabels = ['$'+str(x).strip()+'$' for x in ylabels]
-
-zlabels = range(int(-20),int(50),20)
-zlocs = [zloc(x) for x in zlabels]
-zlabels = ['$'+str(x).strip()+'$' for x in zlabels]
-
[…]

### 2015-12-02 jet_nochirp.py

@@ -9,7 +9,7 @@ import math
 import sys
 
 psi = 2.0 # degrees
-i = 40.0 # degrees
+i = 20.0 # degrees
 theta = 0.0 # degrees
 beta = 0.90 # jet speed in units of c 
 d = 100.0 # Mpc; Distance between jet and observer
@@ -20,7 +20,7 @@ output_filename = 'jet_i%2d_psi%2d_beta_%3.2f' % (i, psi, beta)
 save_pdf = False
 
 psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # radians 
+i *= np.pi/180.0 # r]adians 
 theta *= np.pi/180.0 # radians
 
 c = 3.0e5 # km/s; speed of light
@@ -79,12 +79,10 @@ phi_z_obsb = z_obsb/d * 180.0/np.pi * 3600.0 # arcsec
 phi_y_obs *= 1.0e3
 phi_z_obs *= 1.0e3
 
-with open('test.out','w') as f: 
-    for i in xrange(phi_y_obs.size):
-        f.write(str(phi_y_obs[i])+'  '+str(phi_z_obs[i])+'\n')
+# with open('test.out','w') as f: 
+#     for i in xrange(phi_y_obs.size):
+#         f.write(str(phi_y_obs[i])+'  '+str(phi_z_obs[i])+'\n')
 
-sys.exit() 
-        
 fig = plt.figure(figsize=(7, 7), dpi=100)
 ax = fig.add_subplot(1, 1, 1)
 # ax.set_xlim(-0.04,0.14)

### 2015-12-02 model_100muas.py

@@ -177,7 +177,6 @@ for i in xrange(phi_z.size):
 
 fig = plt.figure(figsize=(5, 5), dpi=100)
 a2 = gf(a, 10.0)
-#a2 = a
 plt.contour(a2,100,colors='k',linestyles='solid')
 
 ylabels = range(int(-60),int(70),20)
@@ -193,7 +192,7 @@ plt.yticks(ylocs, ylabels)
 plt.xticks(zlocs, zlabels)
 plt.xlabel('mas')
 plt.ylabel('mas')
-plt.savefig('jet_2mas.pdf',bbox_inches='tight')
+plt.savefig('jet_100muas.pdf',bbox_inches='tight')
 
 sys.exit() 
 

### 2015-12-02 model_3d.py

@@ -1,0 +1,64 @@
+def I(theta): 
+
+    a = 0.1
+    b = 10.0
+    i = 2.0
+    l = 3.0
+
+    u = np.linspace(0.0, 20.0*np.pi, 1000)
+
+    def z(u):
+        return (a/(2.0*np.pi)) * u * (u/(2.0*np.pi))**2
+
+    zv = z(u)
+
+    def x(u):
+        return (z(u)**-0.2) * (b/(2.0*np.pi)) * u * np.cos(u)
+
+    def y(u):
+        return (z(u)**-0.2) * (b/(2.0*np.pi)) * u * np.sin(u)
+
+    xv = x(u)
+    yv = y(u)
+
+    def ri(i):
+        return np.matrix([[np.cos(i), 0.0, np.sin(i)],[0.0, 1.0, 0.0],[-np.sin(i), 0.0, np.cos(i)]])
+
+    def rl(l):
+        return np.matrix([[np.cos(l), -np.sin(l), 0.0],[np.sin(l), np.cos(l), 0.0],[0.0, 0.0, 1.0]])
+
+    zvarr = zv*0.5 
+    iarr = zvarr/zvarr.max()
+    iarr *= np.pi/2.0 
+    c = np.dstack((xv, yv, zv))
+    c = np.squeeze(c)
+    d = np.zeros((1000,3))
+    lm = rl(l) 
+    for n in range(1000):
+        d[n] = c[n]*ri(iarr[n])*lm 
+
+    xv = d[1:,0]
+    yv = d[1:,1]
+
+    nc = 100                       
+    a = np.zeros((nc,nc),dtype=np.float32) 
+    zl = xv.min() - 5.0
+    zu = xv.max() + 5.0
+    yl = yv.min() - 5.0
+    yu = yv.max() + 5.0 
+    lz = zu - zl 
+    ly = yu - yl 
+    dz = lz/nc 
+    dy = -ly/nc # Because "y" coordinate increases in opposite direction to "y" array index of a (or a2).
+
+    def zloc(cood):
+        return int((cood-zl)/dz) + 1 
+
+    def yloc(cood):
+        return int((cood-yl)/dy) + 1 
+
+    for i in xrange(xv.size):
+        zpos = zloc(xv[i]) 
+        ypos = yloc(yv[i])
+        a[ypos, zpos] += 1.0
+

### 2015-12-02 Paper/paper.tex
Thank the referee

@@
 
-We acknowledge useful discussions with Martin Haehnelt, Enrico
-Ramirez-Ruiz and Alberto Rorai.  This work was supported in part by
-NSF grant AST-1312034 (for A.~L.).  G.~K. gratefully acknowledges
-support by he FP7 ERC Advanced Grant Emergence-320596.
+We acknowledge helpful comments from the anonymous referee and useful
+discussions with Martin Haehnelt, Enrico Ramirez-Ruiz and Alberto
+Rorai.  This work was supported in part by NSF grant AST-1312034 (for
+A.~L.) and by the FP7 ERC Advanced Grant Emergence-320596 (for G.~K.).
 

### 2015-12-02 Paper/paper.tex
"Letter" is now "paper".

@@
 
-In this Letter, we study the morphology of jets from binary BHs whose
-dynamics is doiminated by GW emission.  Jets in binary BHs have been
+In this paper, we study the morphology of jets from binary BHs whose
+dynamics is dominated by GW emission.  Jets in binary BHs have been
 studied in a number of previous works \citep{1980ApJ...238L.129S,
@@
 coalescence, i.e., when the binary's dynamics is dominated by GW
-emission.  In this Letter we study the morphology of such jets.
+emission.  In this paper we study the morphology of such jets.
 
@@
 Post-Newtonian corrections, but we not develop this further in this
-Letter.  We make the code for implementing the model presented in this
-Letter and for fitting it to observed jets using MCMC analyses
+paper.  We make the code for implementing the model presented in this
+paper and for fitting it to observed jets using MCMC analyses
 available at \url{https://github.com/gkulkarni/JetMorphology} under
@@
 Timing Arrays are sensitive to even higher binary masses.  Monitoring
-of jet morphologies for the features outlined in this letter will put
+of jet morphologies for the features outlined in this paper will put
 a lower bound on the expected abundance of bright GW

### 2015-12-02 Paper/paper.tex
Correct typos as pointed out by referee

@@
 example in Figure~\ref{fig:jet_mdot1.0}.  Binaries with higher masses
-have large orbital periods that will result is a helix with wider
+have large orbital periods that will result in a helix with wider
 seperation between its cycles.  These objects would therefore be
@@
 widening opening angle of the jet, while the $(u/2\pi)^{\beta}$ term
-in the last equation parameterises the chirp.  This model can be fit
-to an observed jet on the sky by using two rotation matrices,
-$\mathcal{R}_\lambda$ and $\mathcal{R}_i$, which rotate the model
-in the above equtions to match with the jet precession axis and the
-jet symmetry axis.  These matrices are given by
+in the last equation parameterises the chirp.  This model can be
+fitted to an observed jet on the sky by using two rotation matrices,
+$\mathcal{R}_\lambda$ and $\mathcal{R}_i$, which rotate the model in
+the above equtions to match with the jet precession axis and the jet
+symmetry axis.  These matrices are given by
 
@@
 We thus arrive at a model with 7 parameters ($\alpha$, $\beta$,
-$\gamma$, $A$, $B$, $i_0$, $\lambda_0$), which can be fit to any
+$\gamma$, $A$, $B$, $i_0$, $\lambda_0$), which can be fitted to any
 observed jet, such as that in Figure~\ref{fig:jet_mdot1.0}.  Note that

### 2015-12-09 fitjet_3d.py
Make PDF plot for new paper revision

@@ -183,4 +183,4 @@ if plot_chain:
 
 fig = triangle.corner(samples, labels=['$A$', '$B$', '$i_0$', r'$\lambda_0$', r'$\alpha$', r'$\beta$', r'$\gamma$'],
                       truths=result.x)
-fig.savefig("triangle.png")
+fig.savefig("triangle.pdf")

### 2015-12-09 jet3d.py

@@ -55,7 +55,7 @@ yv = d[1:,1]
 plt.plot(xv, yv)
 plt.show()
 
-sys.exit()
+#sys.exit()
 
 nc = 100                       
 a = np.zeros((nc,nc),dtype=np.float32) 
@@ -64,7 +64,8 @@ zu = xv.max() + 5.0
 yl = yv.min() - 5.0
 yu = yv.max() + 5.0 
 lz = zu - zl 
-ly = yu - yl 
+ly = yu - yl
+print lz, ly
 dz = lz/nc 
 dy = -ly/nc # Because "y" coordinate increases in opposite direction to "y" array index of a (or a2).
 

### 2015-12-09 model.py
Create better plots for new paper revision

@@ -147,7 +147,7 @@ phi_y = np.concatenate([phi_y_obsb, phi_y_obs])
 intensity = np.concatenate([intensity_b, intensity])
 max_intensity = intensity.max()
 intensity /= max_intensity
-intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
+#intensity = np.array([np.log10(x) if x>0.0 else -2.0 for x in intensity])
 
 nc = 1000
 a = np.zeros((nc,nc),dtype=np.float32) 
@@ -155,6 +155,9 @@ zl = phi_z.min()-5.0
 zu = phi_z.max()+5.0
 yl = phi_y.min()-5.0
 yu = phi_y.max()+5.0
+
+zl, zu = -10.0, 10.0
+yl, yu = -5.0, 5.0
 
 print zl, zu
 print yl, yu 
@@ -165,34 +168,43 @@ dy = ly/nc
 dz = lz/nc
 print dy, dz
 def zloc(x):
-    return int((x-zl)/dz) + 1 
+    return int((x-zl)/dz)# + 1 
 
 def yloc(x):
-    return int((x-yl)/dy) + 1 
+    return int((x-yl)/dy)# + 1 
 
 for i in xrange(phi_z.size):
+    if phi_z[i] > zu or phi_z[i] < zl: continue
+    if phi_y[i] > yu or phi_y[i] < yl: continue 
     zpos = zloc(phi_z[i]) 
     ypos = yloc(phi_y[i])
-    a[ypos, zpos] += abs(intensity[i])*100000.0
+    a[ypos, zpos] += abs(intensity[i])
 
-fig = plt.figure(figsize=(5, 5), dpi=100)
+fig = plt.figure(figsize=(7,7), dpi=100)
+ax = fig.add_subplot(1, 1, 1)
 a2 = gf(a, 10.0)
-#a2 = a
-plt.contour(a2,100,colors='k',linestyles='solid')
 
-ylabels = range(int(-60),int(70),20)
+ylabels = range(int(yl),int(yu))
 ylocs = [yloc(x) for x in ylabels]
 ylabels = ['$'+str(x).strip()+'$' for x in ylabels]
 
-zlabels = range(int(-20),int(50),20)
+zlabels = range(int(zl),int(zu),3)
 zlocs = [zloc(x) for x in zlabels]
 zlabels = ['$'+str(x).strip()+'$' for x in zlabels]
 
[…]

### 2015-12-09 Paper/paper.tex
Finalise paper for resubmission

@@
   exhibits milliarcsecond scale wiggles, and the conical surface of
-  jet precession is twisted due to apparant superluminal motion.  The
+  jet precession is twisted due to apparent superluminal motion.  The
   rapidly increasing orbital velocity of the binary gives the jet an
@@
 Binaries of supermassive black holes (BHs) arise naturally as a result
-of mergers of galaxies in the context of heirarchical galaxy formation
+of mergers of galaxies in the context of hierarchical galaxy formation
 \citep{2000MNRAS.311..576K, 2002MNRAS.336L..61H, 2008ApJ...676...33D,
@@
 
-General relativisitic (GR) numerical simulations can predict the
+General relativistic (GR) numerical simulations can predict the
 binary's evolution in its final stage up to coalescence
@@
     as its half-opening angle.  Note that $v_\mathrm{orb}$ is
-    exagerated here.  In practice, $v_\mathrm{orb}\ll
+    exaggerated here.  In practice, $v_\mathrm{orb}\ll
     v_\mathrm{jet}\sim c$.  The figure indicates that
@@
   angular diameter distance $d = 100$~Mpc.  The jets are twisted
-  because of the apparant superluminal motion.  Also evident is the
+  because of the apparent superluminal motion.  Also evident is the
   opening of the jet and the radio jet chirp.  Right panel shows
@@
 As the binary evolves, the combination of increasing orbital speed and
-the apparant superluminal motion effect results in a jet morphology
+the apparent superluminal motion effect results in a jet morphology
 with an increasing opening angle and increasingly rapid periodic
@@
 forward jet and shows that the binary creates wiggles in the jet
-morphology of the order of a milli-arcsecond.  The duration of the
+morphology of the order of a milliarcsecond.  The duration of the
 binary evolution captured by Figure~\ref{fig:jet_mdot1.0} is 100 yr.
@@
 The increasing orbital speed of the binary leads to an increase in the
-jet's precession angular velocity.  This increases the helicity of the
-jet such that the jet is wound up increasingly tightly closer to the
-source.  This gives rise to a ``chirping'' morphology as seen in
-Figure~\ref{fig:jet_mdot1.0}.  The chirp is easier to resolve in jets
-that are closer to the line of sight, where superluminal motion can
-magnify the forward jet.  A chirping black hole is an unambiguous
+jet's precession angular velocity.  This increases the helical winding
+of the jet such that the jet is wound up increasingly tightly closer
+to the source.  This gives rise to a ``chirping'' morphology as seen
+in Figure~\ref{fig:jet_mdot1.0}.  The chirp is easier to resolve in
+jets that are closer to the line of sight, where superluminal motion
+can magnify the forward jet.  A chirping black hole is an unambiguous
 signature of a rapidly evolving binary BH system; the chirp is direct
@@
 \begin{itemize}
-\item Increase in the helicity of the jet towards the central source (``chirpyness''),
+\item Increase in the helical winding of the jet towards the central source (``chirpyness''),
 \item Widening of the opening angle towards the central source, and 
@@
 
-Observations of such a milliarc second scale geometry with the three
+Observations of such a milliarcsecond scale geometry with the three
 features described above can thus flag a BH binary nearing the end of
@@
 $\mu$Jy$/$beam.  The European VLBI network (EVN) which can resolve up
-to 0.13 milliarc seconds and can detect surface brightness down to 19
+to 0.13 milliarcseconds and can detect surface brightness down to 19
 $\mu$Jy$/$beam.  The Very Large Array (VLA) has an angular resolution
@@
 $\mu$Jy$/$beam.  The future Square Kilometer Array (SKA) on its own
-will have 2 milliarc second resolution at around 10 GHz, but SKA will
+will have 2 milliarcsecond resolution at around 10 GHz, but SKA will
 be an order of magnitude more sensitive than current instruments with
@@
 maximum) in Figure~\ref{fig:jet_mdot1.0} (right panel) has separations
-of $\sim 1$ mas.  It should therefore be possible to resolve the jet
-at an angular resolution of $\sim 2$ mas.  (In the next section, we
-discuss an algorithm to fit the jet model to observational data.)  If
-it is aligned closer to the line of sight, the forward jet is
-magnified due to apparant superluminal motion.  This will make it
-possible to resolve jets of objects that are farther away than the
-example in Figure~\ref{fig:jet_mdot1.0}.  Binaries with higher masses
-have large orbital periods that will result in a helix with wider
-seperation between its cycles.  These objects would therefore be
-easier to detect.  We have not modelled the absolute surface
-brightness of jets, only relative variation in the jet brightness (as
-shown in Figure~\ref{fig:jet_mdot1.0}), but due to the $(1+z)^{-4}$
-dimming, it will be exceedingly difficult to detect such jets at
-higher redshifts.
+of $\sim 1$ mas.  Figure~\ref{fig:smooth_jet} shows inner 100 square
+mas of the jet in Figure~\ref{fig:jet_mdot1.0} smoothed at 200 $\mu$as
+and 2 mas respectively.  We find that at 200 $\mu$as resolution the
+chirp is clearly resolved.  At a lower resolution of 2 mas, the chirp
+is no longer resolved but widening of the jet opening angle can
+possibly be inferred.  (In the next section, we discuss an algorithm
+to fit the jet model to observational data.)  Note that other
+geometric effects can influence these conclusions.  If it is aligned
+closer to the line of sight, the forward jet is magnified due to
+apparant superluminal motion.  This will make it possible to resolve
+jets of objects that are farther away than the example in
+Figure~\ref{fig:jet_mdot1.0}.  Binaries with higher masses have large
+orbital periods that will result in a helix with wider separation
+between its cycles.  These objects would therefore be easier to
+detect.  We have not modeled the absolute surface brightness of jets,
+only relative variation in the jet brightness (as shown in
+Figure~\ref{fig:jet_mdot1.0}), but due to the $(1+z)^{-4}$ dimming, it
+will be exceedingly difficult to detect such jets at higher redshifts.
 
@@
 central source from 11$^\circ$ to 45$^\circ$ in just two cycles
-\citet{1982ApJ...262..478G}.  The observing beamwidth of about 400 mas
+\citep{1982ApJ...262..478G}.  The observing beamwidth of about 400 mas
 makes it difficult to resolve a chirp in this source, but it is a
@@
 \citet{2015MNRAS.453.1562G} have reported that 111 candidates show
-periodic variability in brightness in the Cataline Real-time Transient
+periodic variability in brightness in the Catalina Real-time Transient
 Survey, indicating presence of binary BHs.  An analysis of these
@@
 likelihood of the presence of chirpy jets.  In low resolution wide
-field survey, detecting modulations in the lightcurves would be an
+field surveys, detecting modulations in the light curves would be an
 easier diagnostic of a binary BH \citep{2011ApJ...743..136O,
@@
 
+\begin{figure*}
+\begin{center}
+  \begin{tabular}{cc}
+    % model.py 
+    \includegraphics*[width=\columnwidth]{jet_200muas.pdf} &
+    % model.py 
+    \includegraphics*[width=\columnwidth]{jet_2mas.pdf}
+  \end{tabular}
+\end{center}
+\caption{Central 100 square mas of the jet in
+  Figure~\ref{fig:jet_mdot1.0} smoothed at 200 $\mu$as (left panel)
+  and 2 mas (right panel) respectively.  The intensity is normalised
+  in arbitrary units.}
+\label{fig:smooth_jet}
+\end{figure*}
+
 \subsection{Deriving binary parameters from observations of chirping jets}
+\label{sec:geometric_model}
 
@@
 \begin{align}
-  z &= x^{-\alpha} \frac{B}{2\pi} u \cos{u}, \\
+  z &= x^{-\alpha} \frac{B}{2\pi} u \cos{u}, \label{eqn:z}\\
   y &= x^{-\alpha} \frac{B}{2\pi} u \sin{u}, \\
@@
 $\mathcal{R}_\lambda$ and $\mathcal{R}_i$, which rotate the model in
-the above equtions to match with the jet precession axis and the jet
+the above equations to match with the jet precession axis and the jet
 symmetry axis.  These matrices are given by
@@
   i = \gamma x + i_0.
+  \label{eqn:i}
 \end{equation}
@@
 non-chirpy ones.
+
+\begin{figure*}
+\begin{center}
+  \includegraphics*[width=\textwidth]{triangle.pdf}
+\end{center}
+\caption{One and two dimensional projections of posterior probability
+  distributions of the 7-parameter geometric model described in
+  Section~\ref{sec:geometric_model} to mock data.  Wide, uniform
+  priors were chosen for all parameters.  The two-dimensional
+  projections also show that there are no parameter degeneracies in
+  the geometric model.  As described in the text, these parameter
+  values can then be used to derive various physical parameters
+  describing the black hole binary.}
+\label{fig:triangle}
+\end{figure*}
+
+Figure \ref{fig:triangle} shows how the geometric model of this
+section works in practice, by fitting it to mock data using Markov
+Chain Monte Carlo (MCMC).  We used the \texttt{emcee} package for this
+purpose \citep{2013PASP..125..306F}.  Mock data are created by using
+Equations~(\ref{eqn:z})--(\ref{eqn:i}) and then smoothing the
+resultant jet morphology by $\sim$ 2 mas.  Figure \ref{fig:triangle}
+shows posterior probability distributions of the parameters of our
+7-parameter model when it is fit to these mock data.  Wide, uniform
+priors were chosen for all parameters.  The posteriors demonstrate
+that there is no degeneracy in the geometric model parameters.  In
+fact the narrow ranges of values for which the posteriors are non-zero
+for each parameter show that the model fit the mock data very well.
 
@@
 chirp, can be used to directly infer the evolution of the binary
-separation as $da/dP$, which also contraints the quantity $a^4/M^3$
+separation as $da/dP$, which also constraints the quantity $a^4/M^3$
 via Equation~(\ref{eqn:shrink}). These two constraints can be used to
@@
 the GW phase, the opening angle of cone rapidly increases within a
-separation of tens of milli-arcseconds at a distance of $\sim
-100$~Mpc, depending on the angle that the jet makes relative to the
-line of sight.  In addition, the periodicity of the jet's motion on
-the conical surface increases with the orbital period of the binary in
+separation of tens of milliarcseconds at a distance of $\sim 100$~Mpc,
+depending on the angle that the jet makes relative to the line of
+sight.  In addition, the periodicity of the jet's motion on the
+conical surface increases with the orbital period of the binary in
 this stage of its evolution.  Finally, due to the apparent
@@
 
-Milli-arcsecond features can be resolved using long baseline radio
+Milliarcsecond features can be resolved using long baseline radio
 interferometry \citep{2001ChJAA...1..236Q}.  At higher redshifts, jets
@@
 possible, which could increase the detection rate of these objects
-\citep{2014arXiv1412.5971P}.
+\citep{2015aska.confE.143P}.
 
@@
 We acknowledge helpful comments from the anonymous referee and useful
-discussions with Martin Haehnelt, Enrico Ramirez-Ruiz and Alberto
+discussions with Martin Haehnelt, Enrico Ramirez-Ruiz, and Alberto
 Rorai.  This work was supported in part by NSF grant AST-1312034 (for

### 2015-12-09 3dplot.py
Prepare for publication

@@ -1,19 +1,0 @@
-import matplotlib as mpl
-from mpl_toolkits.mplot3d import Axes3D
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-fig = plt.figure()
-ax = fig.gca(projection='3d')
-
-# theta = np.linspace(-4 * np.pi, 4 * np.pi, 100)
-# z = np.linspace(-2, 2, 100)
-# r = z**2 + 1
-# x = r * np.sin(theta)
-# y = r * np.cos(theta)
-
-ax.plot(xv, yv, zv)
-
-plt.show()
-

### 2015-12-09 I.py

@@ -1,71 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt
-from matplotlib import cm
-from scipy.ndimage.filters import gaussian_filter as gfx
-
-a = 12.0
-b = 0.5
-i0 = 1.5
-l0 = 0.0
-phi = 1.0
-
-def F(u):
-    return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
-            (a/(2.0*np.pi)) * u * np.sin(i0))
-
-def G(u):
-    return (b/(2.0*np.pi)) * u * np.sin(u-phi)
-
-def x(u):
-    return F(u)*np.cos(l0) - G(u)*np.sin(l0)
-
-def y(u):
-    return F(u)*np.sin(l0) + G(u)*np.cos(l0)
-
-u = np.linspace(0.0, 20.0*np.pi, 1000)
-
-xv = x(u)
-yv = y(u)
-
-nc = 1000
-a = np.zeros((nc,nc),dtype=np.float32) 
-zl = xv.min() - 5.0
-zu = xv.max() + 5.0
-yl = yv.min() - 5.0
-yu = yv.max() + 5.0 
-lz = zu - zl 
-ly = yu - yl 
-dy = ly/nc
-dz = lz/nc 
-
-def zloc(x):
-    return int((x-zl)/dz) + 1 
-
-def yloc(x):
-    return int((x-yl)/dy) + 1 
-
-for i in xrange(xv.size):
-    zpos = zloc(xv[i]) 
-    ypos = yloc(yv[i])
-    # a[ypos, zpos] += intensity[i]
-    a[ypos, zpos] += 1.0
-
-a2 = gf(a, 10.0)
-a2.tofile('mockdata.dat')
-#plt.contour(a2,200,colors='k',linestyles='solid')
-
-def m(i,j):
-    return a2[i,j]
-
-plt.imshow(a2, cmap=cm.Blues) 
-plt.show()
-
-
-
-# fig = plt.figure(figsize=(7, 7))
-# ax = fig.add_subplot(1, 1, 1)
-# # plt.xlim(-2.0,2.0)
-# # plt.ylim(-2.0,2.0)
-# plt.plot(xv, yv)
-# plt.show()
-

### 2015-12-09 fitjet.py

@@ -1,144 +1,0 @@
-import numpy as np
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-from matplotlib import cm
-import scipy.optimize as op
-import emcee
-import triangle
-
-a2 = np.fromfile('mockdata2.dat',dtype=np.float32)
-
-def I(theta, i, j):
-
-    a, b, i0, l0 = theta
-    phi = 1.0
-    u = np.linspace(0.0, 20.0*np.pi, 1000)
-
-    def F(u):
-        return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
-                (a/(2.0*np.pi)) * u * np.sin(i0))
-
-    def G(u):
-        return (b/(2.0*np.pi)) * u * np.sin(u-phi)
-
-    def x(u):
-        return F(u)*np.cos(l0) - G(u)*np.sin(l0)
-
-    def y(u):
-        return F(u)*np.sin(l0) + G(u)*np.cos(l0)
-    
-    xv = x(u)
-    yv = y(u)
-
-    nc = 100
-    a = np.zeros((nc,nc),dtype=np.float32) 
-    zl = xv.min() - 5.0
-    zu = xv.max() + 5.0
-    yl = yv.min() - 5.0
-    yu = yv.max() + 5.0 
-    lz = zu - zl 
-    ly = yu - yl 
-    dy = ly/nc
-    dz = lz/nc
-
-    def zloc(x):
-        return int((x-zl)/dz) + 1 
-
-    def yloc(x):
-        return int((x-yl)/dy) + 1 
-
-    for c in xrange(xv.size):
-        zpos = zloc(xv[c]) 
-        ypos = yloc(yv[c])
-        a[ypos, zpos] += 1.0
-
-    return a[i,j]
-
-def neglnlike(theta, i, j, intensity, intensity_err):
-    model = I(theta, i, j)  
-    inv_sigma2 = 1.0/intensity_err**2 
-    return 0.5*(np.sum((intensity-model)**2*inv_sigma2 - np.log(inv_sigma2)))
-
-a2_err = np.zeros_like(a2)
-a2_err += 0.1
-
-theta_guess = (12.0, 0.5, 1.5, 0.0)
-
-iarr = np.arange(100)
-jarr = np.arange(100) 
-
-iarr = np.repeat(iarr, 100)
-jarr = np.tile(jarr, 100)
-
-
[…]

### 2015-12-09 fitjet2.py

@@ -1,139 +1,0 @@
-""" Uses image subtraction; otherwise same as fitjet.py """ 
-
-import numpy as np
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-from matplotlib import cm
-import scipy.optimize as op
-import emcee
-import triangle
-
-a2 = np.fromfile('mockdata2.dat',dtype=np.float32)
-
-def I(theta):
-
-    a, b, i0, l0 = theta
-    phi = 1.0
-    u = np.linspace(0.0, 20.0*np.pi, 1000)
-
-    def F(u):
-        return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
-                (a/(2.0*np.pi)) * u * np.sin(i0))
-
-    def G(u):
-        return (b/(2.0*np.pi)) * u * np.sin(u-phi)
-
-    def x(u):
-        return F(u)*np.cos(l0) - G(u)*np.sin(l0)
-
-    def y(u):
-        return F(u)*np.sin(l0) + G(u)*np.cos(l0)
-    
-    xv = x(u)
-    yv = y(u)
-
-    nc = 100
-    a = np.zeros((nc,nc),dtype=np.float32) 
-    zl = xv.min() - 5.0
-    zu = xv.max() + 5.0
-    yl = yv.min() - 5.0
-    yu = yv.max() + 5.0 
-    lz = zu - zl 
-    ly = yu - yl 
-    dy = ly/nc
-    dz = lz/nc
-
-    def zloc(x):
-        return int((x-zl)/dz) + 1 
-
-    def yloc(x):
-        return int((x-yl)/dy) + 1 
-
-    for c in xrange(xv.size):
-        zpos = zloc(xv[c]) 
-        ypos = yloc(yv[c])
-        a[ypos, zpos] += 1.0
-
-    return a.flatten() 
-
-def neglnlike(theta, intensity, intensity_err):
-    model = I(theta)  
-    inv_sigma2 = 1.0/intensity_err**2 
-    return 0.5*(np.sum((intensity-model)**2*inv_sigma2 - np.log(inv_sigma2)))
-
-a2_err = np.zeros_like(a2)
-a2_err += 0.1
-
-theta_guess = (12.0, 0.5, 1.5, 0.0)
-result = op.minimize(neglnlike, theta_guess, args=(a2, a2_err), method='Nelder-Mead') 
-
-print result.x
-print result.success
-
[…]

### 2015-12-09 fitjet_3d.py

@@ -1,4 +1,11 @@
-""" Uses image subtraction; otherwise same as fitjet.py """ 
+"""
+
+File: fitjet_3d.py
+
+Fits a geometric model to mock jet data. Uses image subtraction;
+otherwise same as fitjet.py
+
+""" 
 
 import numpy as np
 import matplotlib as mpl
@@ -9,6 +16,7 @@ import emcee
 import triangle
 import sys
 
+# These mock data are produced by jet3d.py.
 a2 = np.fromfile('mockdata_3d_nc100.dat',dtype=np.float32)
 
 def I(theta): 

### 2015-12-09 geom.py

@@ -1,73 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt
-from matplotlib import cm
-from scipy.ndimage.filters import gaussian_filter as gf
-
-# a = 12.0
-# b = 0.5
-# i0 = 1.5
-# l0 = 0.0
-
-a, b, i0, l0 = 1.18611945e+01, 4.80667425e-01, 1.56853292e+00, 5.17469193e-05 
-
-#a, b, i0, l0 = 1.19996440e+01, 5.00329139e-01, 1.49997638e+00, 2.47767666e-04
-phi = 1.0
-
-def F(u):
-    return ((b/(2.0*np.pi)) * u * np.cos(u-phi) * np.cos(i0) +
-            (a/(2.0*np.pi)) * u * np.sin(i0))
-
-def G(u):
-    return (b/(2.0*np.pi)) * u * np.sin(u-phi)
-
-def x(u):
-    return F(u)*np.cos(l0) - G(u)*np.sin(l0)
-
-def y(u):
-    return F(u)*np.sin(l0) + G(u)*np.cos(l0)
-
-u = np.linspace(0.0, 20.0*np.pi, 1000)
-
-xv = x(u)
-yv = y(u)
-
-nc = 100
-a = np.zeros((nc,nc),dtype=np.float32) 
-zl = xv.min() - 5.0
-zu = xv.max() + 5.0
-yl = yv.min() - 5.0
-yu = yv.max() + 5.0 
-lz = zu - zl 
-ly = yu - yl 
-dy = ly/nc
-dz = lz/nc 
-
-def zloc(x):
-    return int((x-zl)/dz) + 1 
-
-def yloc(x):
-    return int((x-yl)/dy) + 1 
-
-for i in xrange(xv.size):
-    zpos = zloc(xv[i]) 
-    ypos = yloc(yv[i])
-    # a[ypos, zpos] += intensity[i]
-    a[ypos, zpos] += 1.0
-
-a2 = gf(a, 1.0)
-#a2.tofile('mockdata2.dat')
-#plt.contour(a2,200,colors='k',linestyles='solid')
-
-def m(i,j):
-    return a2[i,j]
-
-plt.imshow(a2, cmap=cm.Blues) 
-plt.show()
-
-# fig = plt.figure(figsize=(7, 7))
-# ax = fig.add_subplot(1, 1, 1)
-# # plt.xlim(-2.0,2.0)
[…]

### 2015-12-09 jet3d.py

@@ -1,3 +1,12 @@
+"""
+
+File: jet3d.py 
+
+Creates mock jet data for fitjet_3d.py.
+
+
+"""
+
 import matplotlib as mpl
 import numpy as np
 import matplotlib.pyplot as plt
@@ -38,11 +47,6 @@ iarr = zvarr/zvarr.max()
 iarr *= np.pi/2.0 
 
 c = np.dstack((xv, yv, zv))
-# d = c*ri(i)*rl(l)
-
-# x = np.squeeze(np.asarray(d[:,0]))
-# y = np.squeeze(np.asarray(d[:,1]))
-
 c = np.squeeze(c)
 d = np.zeros((1000,3))
 lm = rl(l) 
@@ -82,8 +86,9 @@ for i in xrange(xv.size):
 
 a2 = gf(a, 1.0)
 
-#a2.tofile('mockdata_3d_nc100.dat')
-#plt.contour(a2,200,colors='k',linestyles='solid')
+save_data = False 
+if save_data: 
+    a2.tofile('mockdata_3d_nc100.dat') # Save for fitjet_3d.py 
 
 plt.imshow(a2, cmap=cm.Blues) 
 plt.show()

### 2015-12-09 jet_image_mock.py

@@ -1,227 +1,0 @@
-"""
-
-File: jet_image.py 
-
-Jet morphology from a BH binary that is in the GW-dominated phase of
-its inspiral.  This code uses a jet brightness model to produce an
-image of the jet on the sky.  Case 1 gives Figure 5 (right panel).
-
-"""
-
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-from matplotlib import cm
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-from scipy.ndimage.filters import gaussian_filter as gf
-
-i = 40.0 # degrees
-i *= np.pi/180.0 # radians 
-beta = 0.9 # jet speed in units of c 
-d = 100.0 # Mpc; Angular distance between jet and observer
-d *= 1.0e3 # kpc
-psi0 = 2.0*np.pi/180.0 # radians
-M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 1.0 # Eddington units
-a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
-pcto_10to16cm = 0.003241
-a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
-
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i*180.0/np.pi, beta, Mdot)
-save_pdf = True
-
-# case = 1, 2 zooms into the central region of the jet 
-case = int(sys.argv[1])
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
[…]

### 2015-12-09 jet_image_try.py

@@ -1,185 +1,0 @@
-"""
-
-File: jet_image.py 
-
-Jet morphology from a BH binary that is in the GW-dominated phase of
-its inspiral.  This code uses a jet brightness model to produce an
-image of the jet on the sky.  Case 1 gives Figure 5 (right panel).
-
-"""
-
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-from matplotlib import cm
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-
-i = 40.0 # degrees
-i *= np.pi/180.0 # radians 
-beta = 0.9 # jet speed in units of c 
-d = 100.0 # Mpc; Angular distance between jet and observer
-d *= 1.0e3 # kpc
-psi0 = 2.0*np.pi/180.0 # radians
-M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 1.0 # Eddington units
-a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
-pcto_10to16cm = 0.003241
-a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
-
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i*180.0/np.pi, beta, Mdot)
-save_pdf = True
-
-# case = 1, 2 zooms into the central region of the jet 
-case = int(sys.argv[1])
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
[…]

### 2015-12-09 jet_nochirp.py

@@ -1,99 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '11'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-
-psi = 2.0 # degrees
-i = 20.0 # degrees
-theta = 0.0 # degrees
-beta = 0.90 # jet speed in units of c 
-d = 100.0 # Mpc; Distance between jet and observer
-
-d *= 1.0e3 # kpc
-
-output_filename = 'jet_i%2d_psi%2d_beta_%3.2f' % (i, psi, beta)
-save_pdf = False
-
-psi *= np.pi/180.0 # radians 
-i *= np.pi/180.0 # r]adians 
-theta *= np.pi/180.0 # radians
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-t = np.logspace(-1.0,2.0,10**6) # yr
-
-def Omega(t):
-    # Angular velocity of jet precession 
-    period = 10.0 # yr
-    return 2.0*np.pi*t/period # radians
-
-def vel(t):
-    # We are following geometry from Gower et al. Figure 1.
-    vx = beta*c*(np.sin(psi)*np.sin(i)*np.cos(Omega(t)) + np.cos(psi)*np.cos(i))
-    vy = beta*c*np.sin(psi)*np.sin(Omega(t))
-    vz = beta*c*(np.cos(psi)*np.sin(i)-np.sin(psi)*np.cos(i)*np.cos(Omega(t)))
-    return sign*vx, sign*vy, sign*vz # km/s
-
-# sign sets Forward or backward jet
-sign = 1
-
-velx, vely, velz = vel(t)
-
-x = np.zeros(t.size)
-y = np.zeros(t.size)
-z = np.zeros(t.size)
-
-y = vely*t*yrbys/kpcbykm # kpc
-z = velz*t*yrbys/kpcbykm # kpc
-
-y_obs = y/(1.0-velx/c)
-z_obs = z/(1.0-velx/c)
-
[…]

### 2015-12-09 model.py

@@ -1,3 +1,11 @@
+"""
+
+File: model.py 
+
+Creates a Gaussian smoothed image of a jet. 
+
+"""
+
 import matplotlib as mpl
 mpl.rcParams['text.usetex'] = True 
 mpl.rcParams['font.family'] = 'serif'
@@ -206,30 +214,4 @@ cb.set_label('intensity', labelpad=20)
 cb.solids.set_edgecolor("face")
 
 plt.savefig('jet_2mas.pdf',bbox_inches='tight')
-
-sys.exit() 
-
-fig = plt.figure(figsize=(7, 7), dpi=100)
-ax = fig.add_subplot(1, 1, 1)
-
-if case==1:
-    ax.set_ylim(-50,50)
-    ax.set_xlim(-10,30)
-
-s=plt.scatter(phi_z,phi_y,s=4,marker='o',edgecolor='none',vmax=0.0,
-              vmin=-2.0,rasterized=True,c=intensity,cmap=cm.Blues)
-
-ax.set_xlabel('mas',labelpad=15)
-ax.set_ylabel('mas',labelpad=15)
-
-divider = make_axes_locatable(ax)
-cax = divider.append_axes("right", "5%", pad="3%")
-cb = plt.colorbar(s, cax=cax)
-cb.set_label('logarithmic intensity', labelpad=20)
-cb.solids.set_edgecolor("face")
-
-if save_pdf: 
-    plt.savefig(output_filename+'.pdf',bbox_inches='tight')
-
-
 

### 2015-12-09 model_100muas.py

@@ -1,222 +1,0 @@
-import matplotlib as mpl
-mpl.rcParams['text.usetex'] = True 
-mpl.rcParams['font.family'] = 'serif'
-mpl.rcParams['font.serif'] = 'cm'
-mpl.rcParams['font.size'] = '22'
-import matplotlib.pyplot as plt
-import numpy as np
-import math 
-import sys
-from matplotlib import cm
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-from scipy.ndimage.filters import gaussian_filter as gf
-
-i = 40.0 # degrees
-i *= np.pi/180.0 # radians 
-beta = 0.9 # jet speed in units of c 
-d = 100.0 # Mpc; Angular distance between jet and observer
-d *= 1.0e3 # kpc
-psi0 = 2.0*np.pi/180.0 # radians
-M = 1.0e10 # Msun; total mass of the equal-mass binary
-Mdot = 1.0 # Eddington units
-a0 = 8.3e-3*(M*1.0e-8)**(3./4.)*(Mdot**(-0.25)) # pc 
-pcto_10to16cm = 0.003241
-a0 /= pcto_10to16cm # 1.0e16 cm
-coeff = -2.56e5/(M*1.0e-8)**3 
-
-output_filename = 'jet_i%2d_beta%3.2f_mdot%3.2f_image' % (i*180.0/np.pi, beta, Mdot)
-save_pdf = True
-
-# case = 1, 2 zooms into the central region of the jet 
-case = int(sys.argv[1])
-
-c = 3.0e5 # km/s; speed of light
-yrbys = 3.154e7
-kpcbykm = 3.086e16
-
-def binary_separation_gw(t):
-    a = (4.0/coeff) * (t - t0 + coeff*a0**4/4.0)
-    a = a**(1./4.)
-    return a
-
-def binary_orbital_period(a_16):
-    t = 1.72*(a_16**1.5)/np.sqrt(M*1.0e-8) # yr 
-    return t
-
-def half_opening_angle_intrinsic(a_16):
-    angle = np.arcsin(np.sin(psi0)*a0/a_16)
-    return angle #*180.0/np.pi # degrees 
-
-if case == 0: 
-    t = np.logspace(-2.0,2.0,10000000)
-    output_filename += '_full'
-elif case == 1:
-    t = np.logspace(-2.0,2.0,10000000)
[…]

### 2015-12-09 model_3d.py

@@ -1,64 +1,0 @@
-def I(theta): 
-
-    a = 0.1
-    b = 10.0
-    i = 2.0
-    l = 3.0
-
-    u = np.linspace(0.0, 20.0*np.pi, 1000)
-
-    def z(u):
-        return (a/(2.0*np.pi)) * u * (u/(2.0*np.pi))**2
-
-    zv = z(u)
-
-    def x(u):
-        return (z(u)**-0.2) * (b/(2.0*np.pi)) * u * np.cos(u)
-
-    def y(u):
-        return (z(u)**-0.2) * (b/(2.0*np.pi)) * u * np.sin(u)
-
-    xv = x(u)
-    yv = y(u)
-
-    def ri(i):
-        return np.matrix([[np.cos(i), 0.0, np.sin(i)],[0.0, 1.0, 0.0],[-np.sin(i), 0.0, np.cos(i)]])
-
-    def rl(l):
-        return np.matrix([[np.cos(l), -np.sin(l), 0.0],[np.sin(l), np.cos(l), 0.0],[0.0, 0.0, 1.0]])
-
-    zvarr = zv*0.5 
-    iarr = zvarr/zvarr.max()
-    iarr *= np.pi/2.0 
-    c = np.dstack((xv, yv, zv))
-    c = np.squeeze(c)
-    d = np.zeros((1000,3))
-    lm = rl(l) 
-    for n in range(1000):
-        d[n] = c[n]*ri(iarr[n])*lm 
-
-    xv = d[1:,0]
-    yv = d[1:,1]
-
-    nc = 100                       
-    a = np.zeros((nc,nc),dtype=np.float32) 
-    zl = xv.min() - 5.0
-    zu = xv.max() + 5.0
-    yl = yv.min() - 5.0
-    yu = yv.max() + 5.0 
-    lz = zu - zl 
-    ly = yu - yl 
-    dz = lz/nc 
-    dy = -ly/nc # Because "y" coordinate increases in opposite direction to "y" array index of a (or a2).
-
-    def zloc(cood):
-        return int((cood-zl)/dz) + 1 
-
-    def yloc(cood):
-        return int((cood-yl)/dy) + 1 
-
-    for i in xrange(xv.size):
-        zpos = zloc(xv[i]) 
-        ypos = yloc(yv[i])
-        a[ypos, zpos] += 1.0
-
